### 2025-03-29 main.tex
Initial commit

%% 
%% Copyright 2007-2024 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.3 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.3 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%% $Id: elsarticle-template-num.tex 249 2024-04-06 10:51:24Z rishi $
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsmath package provides various useful equation environments.
\usepackage{amsmath}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

\journal{Journal of Computer Languages}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

\title{GeckoGraph: A Visual Language for Polymorphic Types}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author{} %% Author name

%% Author affiliation
\affiliation{organization={},%Department and Organization
            addressline={}, 
            city={},
            postcode={}, 
            state={},
            country={}}

%% Abstract
\begin{abstract}
Polymorphic types are an important feature in most strongly typed programming languages. They allow functions to be written in a way that can be used with different data types, while still enforcing the relationship and constraints between the values. However, programmers often find polymorphic types difficult to use and understand and tend to reason using concrete types. We propose GeckoGraph, a graphical notation for types. GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier. We conducted a large-scale human study using GeckoGraph compared to text-based type notation. To our knowledge, this is the largest controlled user study on functional programming ever conducted. The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.  
\end{abstract}

%% Keywords
\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% Add \usepackage{lineno} before \begin{document} and uncomment 
%% following line to enable line numbers
\linenumbers

\section{Introduction} \label{sec:intro}
In programming languages, a polymorphic type \cite{Cardelli1987-fp} can represent values of different types while providing a interface to describe common behaviors for those values. Polymorphic types are central to the succinctness of contemporary statically-typed functional languages, enabling a considerable degree of type-safe abstraction and hence reusability. Parametric polymorphism, a common technique to denote polymorphic type, is available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}. Polymorphism allows programs to be written in a way that is more generic and adaptable to different data types, enabling greater flexibility and code reuse. Polymorphic types are ideal for modeling abstractions, such as properties of mathematical objects and laws that hold on these objects.

Although polymorphic typing promises robustness and a high degree of code reusability, studies~\cite{Jun2000-ec, Jun2000-yu} show that using polymorphism in practice poses challenges, especially for novice users. These studies have shown that humans tend to focus on concrete types and only rely on polymorphic type checking as a last resort. In practice, polymorphic types often pose usability problems for programmers. New polymorphic type variables can be created during type checking. These intermediate type variables are often kept behind the curtain unless a type error is encountered. This often results in programmers resolving type errors with type variables not authored by the programmers themselves (Fig. \ref{fig:example-foldable}).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Foldable}
  \caption{\label{fig:example-foldable} An example type error where a programmer mistakenly provided a \texttt{Char} instead of a \texttt{String} literal. \textbf{Left --} The compiler shows an error message comparing the provided Char type to a confusingly named type \texttt{t0 a0}. \textbf{Right --} GeckoGraph shows the exact message with the two types in graphic notation, highlighting the structural difference rather than identifier names.}
\end{figure}

While type polymorphism is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}, little research focuses on the \textit{usability} of polymorphic types. Hage argues that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}. We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI techniques. To achieve this, we propose GeckoGraph, a graphical notation for types. GeckoGraph aims to complement traditional text-based type notation and make reading, understanding, and comparing types easier. GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts. Our study evaluating 714 participants' ability to solve type adaptation challenges in GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept. The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the type challenges, especially for novice programmers.

\section{GeckoGraph}

GeckoGraph is a visual notation for type annotations in statically typed programming languages. It is intended to work tangibly with text-based annotations, but uses colors, shapes, and symbols to make structures of types easy to identify at a glance. In this section, we describe the design of GeckoGraph and highlight some unique benefits of programming with GeckoGraph.

\subsection{Design of GeckoGraph}
The design of GeckoGraph focuses on visualizing types in functional languages (e.g., Haskell, ML). In this paper, we use Haskell as an example. As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, and some advanced type-level features. However, GeckoGraph could also be used in imperative and multiparadigm languages such as Rust. We provide a GeckoGraph construction library for Haskell.

We identified three main design goals for GeckoGraph based on the challenges of using polymorphic types~\cite{Jun2000-ec, Jun2000-yu} and how programmers tend to use type annotations~\cite{Justin_Lubin2021-yy}, as follows.

\paragraph{\textbf{(D1) Low barrier to learn}}\label{goal1} GeckoGraph should take little to no effort to learn. The rules to translate a text-based type notation to GeckoGraph should be minimal. Where possible, GeckoGraph should be intuitive to programmers who are familiar with text-based type notation.

\paragraph{\textbf{(D2) Easy to parse for humans}}  \label{goal2} GeckoGraph should make the task of reading and understanding type notation easy. It should emphasize the less obvious properties of a type signature. GeckoGraph should eliminate the need for mental backtracking, such as counting opening and closing parentheses and remembering which type classes are required on which variables.

\paragraph{\textbf{(D3) Easy to compare and search}} \label{goal3} GeckoGraph should aim to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss. This also includes the task of choosing an ideal function from a list of potential functions. For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity, one of the argument types, or type class it must fulfill).

\paragraph{Simple Types} 
Simple types, such as type variables and concrete types, are displayed in a cell \includegraphics[height=1em]{figures/SimpleType}: a solid-colored rectangular box with an angled corner on the top left. Each type identifier encodes a distinct color hue of the cell.  Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and strong secondary encoding. The angled corner in the top left provides visual separation between two cells, even when the same color cells are next to each other, allowing GeckoGraph to be zoomed out to extremely small sizes (Section \ref{subsec:space}) without suffering readability (design goal \ref{goal2}).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Design}
  \caption{
        \label{fig:design}
        Examples of various types as represented in GeckoGraph, including type variables (A) and concrete types (B). Data types (C, D, E), function types (F, G, H), and type classes (I, J).
  }
\end{figure}

\paragraph{Data types}
GeckoGraph displays an algebraic data type as a larger cell, where the type constructor half encloses its arguments: \includegraphics[height=1em]{figures/DataType.png}. The arguments are aligned in the bottom right of the cell. Two distinct visual dimensions are used to provide additional visual clarity (design goal \ref{goal2}). Data types containing more arguments (e.g., \texttt{ a b c} or \texttt{(a b) c}) will expand horizontally \includegraphics[height=1em]{figures/DataTypeWide.png}. Data types that are nested (e.g. \texttt{ a (b c)}) will grow taller \includegraphics[height=1.2em]{figures/DataTypeNested.png}. This distinction accommodates our design goal \ref{goal3}. Note that the height of GeckoGraph grows only upwards, but not downwards. Not only does this allow GeckoGraph to be more efficient in its space usage, but it also allows the legend text to be correctly aligned at the bottom and can be read similarly as regular type notation (design goal \ref{goal1}).

\paragraph{Function Types}
Functions are the fundamental building blocks of functional programming languages, and function types are ubiquitous and the most important in type-level programming. In Haskell, \texttt{(->)} is defined as an infix type operator with the right associativity to provide succinct type annotation. GeckoGraph preserves this syntax feature to make the notation more intuitive (design goal \ref{goal1}): the 2 arguments of a function type in GeckoGraph are placed on both sides of the cell \includegraphics[height=1em]{figures/Function.png}. A special function indicator (\texttt{$>>>$}) is displayed at the top of the cell.

Curried functions (e.g., \texttt{ a -> b -> c}) display as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}; the second function overlaps on top of the first, indicating that the second function is the return type of the first. Regular high-order functions (e.g., \texttt{(a -> b) -> c} ) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}. The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and high-order functions (design goal \ref{goal3}). It is easy to tell high-order functions from the vertical position of its function indicator. Similarly, it is easy to count the arity of a function by counting the number of horizontally connected function indicators (Fig. \ref{fig:indicator}).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Indicator}
  \caption{
        \label{fig:indicator}
        An example of using the function indicator. The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators. For high-order functions where functions are arguments of other functions, it is very easy to see the ``order" of functions and how they are arranged. 
  }
\end{figure}

\paragraph{Type Classes} 
Type classes are an intrinsic part of Haskell \cite{Hudak2007-kn}, and many other functional languages. In GeckoGraph, the type classes (e.g., \texttt{ (A a, B a) => a}) are indicated in the extended area below one or more GeckoGraph cells \includegraphics[height=1.2em]{figures/TypeClass.png}. Each type class required on a type variable displays as a square indicator aligned on the right of the extended area. In GeckoGraph, type-class constraints are associated with every instance of the type variable that requires them. This means when displaying the type \texttt{(==) :: Eq a => a -> a -> Bool} in GeckoGraph, the constraint \texttt{Eq} appears in both occurrences of \texttt{a}.

The GeckoGraph type class's design promotes the type class placements rather than the type class names. Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or pop-up window) to identify the name of the type class. We believe that this decision is well justified. For example, when reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable. GeckoGraph helps minimize the effort to associate each type variable with all its type classes (design goal \ref{goal2}).

\subsection{Benefits of Using GeckoGraph}\label{sec:benefits}
\paragraph{Generalization patterns in type classes}
A frequently cited confusion among novices is the overlap between list type \texttt{[a]} and its more generic counterpart \texttt{Foldable} instances. The subtle differences are often not important for beginners. However, when encountering type errors in working with lists, Haskell often explains the error with the \texttt{Foldable} instance. For the example in Fig. \ref{fig:example-foldable} where a programmer mistakenly provided a \texttt{Char} instead of a \texttt{String} literal, the compiler shows an error message comparing the provided \texttt{Char} type with a confusingly named type \texttt{t0 a0}. Although any \texttt{Foldable} instance is perfectly suited for the list \cite{Waldmann2018-hu}, this generalization may reduce programmers' confidence in their understanding of the language and the ability to navigate out of a type error. In GeckoGraph, a list type and a type with a Foldable instance have the same shape. This allows the generic type \texttt{t0 a0} in the error message to assume the same shape as \texttt{[a]}. This generalization of concrete types and abstraction of type-class instances aims to allow for teaching fundamental functional programming concepts without hiding high-level abstractions. The same benefits apply to polymorphic numbers and strings.

\paragraph{Consistent color scheme}\label{par:color-scheme}
A common task in programming is to scan for a desired function from a sea of potentially useful functions, such as library documentation. During scanning, programmers often have partial knowledge of the desired function, e.g., the arity, one of the argument types, or the type class it must fulfill.  A typical example is conversion: using a known \texttt{String} type to produce a desired \texttt{Data.Text}  type. Another example is the `lookup' function: using a known \texttt{Data.Map a b} to produce a desired \texttt{b} type. GeckoGraph supports this task by using consistent colors for the same type identifier. Programmers can rely on the color grouping to scan for the desired type in their project or in third-party library documentation.

\paragraph{Advanced Type Feature Visualization}
The design of GeckoGraph enables the visualization of many advanced type-level features. \textbf{Kind visualization}: if the kind of type variables can be inferred, the kind information is consistently displayed in GeckoGraph. For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} needs at least the kind \texttt{* -> *} because of its use on the right-hand side. GeckoGraph respects this kind information and displays it as a constructor type over an empty structure, indicated using a dotted outline.  \textbf{Qualified constraints}: GeckoGraph's type class notation naturally extends to support qualified constraints. In the type \texttt{forall b. A (a b) => a b}, GeckoGraph shows the scope type class requirement on \texttt{a b} (Fig. \ref{fig:advanced} B).
\textbf{Multiple Parameter Type Class}:  GeckoGraph supports multiple parameter type classes by using multiple shapes with the same color hue to indicate the different parameters of the same type class.  For example, for the type \texttt{A a b => a b},  GeckoGraph shows that the variables a and b both need an A class, but they are the different parameters of A (Fig. \ref{fig:advanced} C).

\paragraph{Precise Interactivity}
Modern programming environments often allow programmers to mouse over part of the source code to query detailed information, such as definition, references or documentation. However, with text-based source code, it is often hard to distinguish whether programmers want the most specific fragment under the cursor or larger blocks. Because of its graphical layout, GeckoGraph allows programmers to precisely select which part of a type signature they intend to query, that is, in Fig. \ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail.

\begin{figure}[h]
\label{fig:advanced}
  \includegraphics[width=\linewidth]{figures/Advanced}
  \caption{
  \label{fig:advanced}
  Advanced features of GeckoGraph. (A) GeckoGraph supports Kind Visualization if the inferred kind is greater than \texttt{*}. (B) GeckoGraph supports qualified constraints by extending the extended area across multiple type variables. (C) GeckoGraph supports Multiple Parameter Type Classes, using different shapes of the same color to indicate that multiple variables must satisfy certain type classes collectively. (D) GeckoGraph supports the precise selection of its sub-structures. }
\end{figure}

\paragraph{Language Agnostic}
GeckoGraph can be implemented in any language that uses static typing. In programming projects, GeckoGraph supports polyglot programming projects. Typical circumstances include projects using foreign function interfaces or multiple languages for client- and server-side programming. GeckoGraph provides a common notation to describe the functionality and features of systems. In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages.  For example, when describing algebraic data types, different language communities use various names: tuple, enum, struct, etc. It is important to realize that these are the same concepts and ignore the minute linguistic barriers.

\subsection{Previous iterations of GeckoGraph}
GeckoGraph was designed through many different iterations. Many research methods were used to verify ideas, including prototyping, cognitive walk-throughs, and formative studies.  We list some notable design elements and their major feedback.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/PreviousVersions}
  \caption{\label{fig:previous}Previous Versions of GeckoGraph. Different encodings represent named types, type variables, type constructors, and high-order functions.}
\end{figure}

\paragraph{\textbf{Encoding the direction of function types}}
This can be found in the design v1 (Fig. \ref{fig:previous}). This design shares many similarities with a prior project \cite{Jung2000-oc} in type visualization. Although many students and experts agree on this, its use of space on both width and height scales is proportional to the types in question, causing too much inconvenience. In addition, this design does not produce a canonical form for a curried function. For example, for composition (.), it is unclear whether it takes two functions as input and returns a binary function or two functions and a single structure as input and returns a different structure.

\paragraph{Encoding the depth and size} Grayscales, gradience, size, and simulated 3-dimensional elevation are promising visual representations for depicting numeric dimensions such as the depth and width of the parse tree. However, gradient and elevation were dismissed because of their requirements for a more demanding rendering process, making GeckoGraph harder to implement in more restricted user interfaces, such as the command line. In addition, all of these visual dimensions reduce readability when scaled down to a very small size.

\paragraph{Encoding symbolic names} In some variations, we tested using different geometric shapes to indicate the symbolic names of type variables and concrete types. We decided against using icons due to the limited number of different shapes until they were indistinguishable. The color provides more encoding spaces, and the letters provide familiarity with the original type annotations. This was shown to help reduce friction in the adoption of GeckoGraph.

\section{Evaluation}
To evaluate the usefulness of GeckoGraph, we designed a controlled experiment in the form of a game called "Zero to Hero". The game contains 10 levels of varying difficulty. At each level, participants are asked to implement a function called "\texttt{zeroToHero}" using only a list of available functions. These available functions are different at each level, and the target types of Zero and Hero vary at each level. The details of each level are provided in the Appendix (Appendix \ref{levels}).

The experiment aims to study how polymorphic types are used and reasoned about during programming tasks. In particular, we studied how programmers scan and select potentially useful functions from a library and compare intended types and actual types during type errors.

\subsection{An example level}
We illustrate the task of the user study using level 4 of the game. At this level (Fig. \ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a b -> Hero b b}. The available functions are \texttt{f1::Zero a b -> Hero b a}, \texttt{f2::Zero a a -> Hero a a}, \texttt{f3::Zero a b -> Hero b a}, and \texttt{f4::Zero a b -> Hero b b}. Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be won without the use of generic functions. The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).

\begin{figure}[h]
  \includegraphics[width=\linewidth]{figures/GamePlay}
  \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero. On this level -- level 4 (Shown at F) -- the players need to implement the function \texttt{ zeroToHero :: Zero a b -> Hero b b} (B). They write their own definitions in the code editor (D) using a set of provided functions (G). The inferred type of their current definition is shown in (C). When ready, they can test their solution by clicking on the \textit{Attempt} button (A). They can also skip a level by clicking on the \textit{Bypass} button next to it. The output from the compiler, if there is any, is shown in a window below (E). The GeckoGraph in the screenshot uses a different color scheme that is optimized for computer screens.}
\end{figure}

At each level of the game, the programmer must select the right functions to achieve the target result. In particular, for this level, participants must discover that only \texttt{f4} and \texttt{f2} are necessary to produce the desired results. An implement that satisfies the target type is \texttt{zeroTohero z = f2  (f4  z)}.

\subsection{Recruitment}
Participants were recruited online through the Haskell community on Reddit and Discord. Participation is fully anonymized; detailed ethical implications of these experiments were reviewed and approved by the IRB of the authors' institution.

\subsection{Group Assignments}

The experiment uses a between-subject design. However, all participants receive both treatments (with and without GeckoGraph) during their runs.
Participants are assigned to one of two groups. Both groups receive the same tasks in the same order. Group one participants are assisted by GeckoGraph on even levels and only text-based type annotation on odd levels. Group two participants are the same but with the order flipped. Both groups have access to the text-based type signature for all tasks. The number of participants in the two groups is counterbalanced.

\subsection{Hypothesis}
In programming tasks that involve reading and understanding polymorphic types, graphic notation using visual elements that provide high grouping strength (colors, shapes, sizes, and symbols) can improve the performance of such tasks compared to traditional text-based type notation. Our null hypothesis is that "Using graphic notation has no effect compared to traditional text-based type notation." This hypothesis and the task design were registered at the Open Science Foundation prior to data collection.

\subsection{Task Design} \label{subsection:task}
Participants in both groups receive the same 10 tasks. The tasks start off easy but gradually increase in difficulty.  In each task, a target type signature of the function \texttt{zeroToHero} is given to the participants. Participants are provided with a list of available functions to implement the target function. This is to simulate the tasks of selecting useful functions from a library. In addition, participants are not allowed to use any other functions or variables outside the provided functions; even the Haskell prelude is not available. This ensures that everyone has the same knowledge and minimizes the effect of familiarity.

Participants can skip a level during the game if they are stuck. We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their qualitative input of the tool. We limit the number of skips that a participant can use during the game to four times so that submitting qualitative feedback without completing at least some levels is impossible.

\subsection{Measurements}
During the study, the time spent by participants on each task is recorded. We also record the resulting status of each level, whether it is a success or failure. Before each run, participants nominate their level of Haskell experience on a four-level scale: beginner, familiar,  knowledgeable, and expert.  If a participant has completed all 10 levels (with the help of skipping), we invite the participant to complete a post-study survey. In it, we ask for their opinion on how intuitive the GeckoGraph design is, how distracting they find GeckoGraph, and how helpful GeckoGraph is during the game, using a seven-point scale. In the end, we ask a few open-ended questions, inviting participants to provide their experience using GeckoGraph and their expectations about the potential applications of GeckoGraph.

Data collection from the human study was stopped after the planned cut-off period of 14 days. After the cut-off date, the ZeroToHero game is open source and available for free evaluation \cite{Anonymous_undated-ne} and repeating our experiment, but no further data was collected.

\section{Results}

During the data collection period, a total of 714 users participated in the study. Among them, 245 are novice users, 216 are familiar with Haskell, 216 are knowledgeable users, and 88 are expert users.

\subsection{Time to complete levels}

The 10 levels are designed to gradually increase difficulty. From the results of the experiment, most of the tasks align with this trend. However, three tasks stand out in Fig. \ref{fig:level-time}.  Level 7 (mean = 334 seconds) is the hardest task in the game in terms of time, followed by level 8 (mean = 228 seconds) and level 5 (mean = 224 seconds). To complete an average level, the beginner group uses an average of 100 seconds, the familiar group uses 90 seconds, the knowledgeable group uses 80 seconds, and the expert group uses 70 seconds. This roughly aligns with self-reported expertise. We show that the task time on each level follows normal distributions using a Shapiro-Wilk test \cite{Shaphiro1965-dx} (p-value  $ \leq 1.018 \times 10^-16$, for an alpha value of 0.05, p less than 0.05 is considered normal distribution).

Level 5, 7 and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience. Level 3 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple. Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 the \texttt{fmap} function of the \texttt{Functor} class. The authors speculate that the more experienced participants are much more familiar with these functions, hence the strong contrast on these three levels.

However, when comparing the task time between the two treatments, we were unable to reject the null hypothesis. In a two-sample T-test, we could not find any significant difference between the two groups overall (p-value = 0.457), nor does there differ between the two groups in any of the four levels of experience (beginner: p-value = 0.845, familiar: p-value = 0.524, Knowledgeable: p-value = 0.712, expert p-value = 0.771).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/LevelTime}
  \caption{\label{fig:level-time} Time spent on each level, with 95\% confidence interval. We show that the difficulty steadily increases across the game, but levels 5, 7, and 8 are significantly harder than the authors intended. The overall task time of each group roughly matches experience level.}
\end{figure}

\subsection{Success rate}
We saw that, overall, GeckoGraph provides a higher success rate (96.88\%) than text-based type notation (94.62\%). This trend can be seen in every experienced group: beginner group (95. 12\% vs. 92. 68\%), familiar group (97. 39\% vs. 93. 34\%), knowledgeable group (96. 82\% vs. 96. 06\%) and expert group (98. 2\% vs. 96. 40\%). We saw the significance decrease as the user's experience increased. When performing a proportion test on each group, we see that the effect is most significant with the beginner group and reject the null hypothesis (z score = 2.0228, p-value = 0.0431), followed by the familiar group (z score = 1.7495, p-value = 0.0802). The knowledgeable group (z score = 1.0295, p-value = 0.3032) and the expert group (z score = 0.8660, p-value = 0.3756) show less significant differences between treatments.

When breaking down the result in each task (Fig. \ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group and task 10 of the familiar group \ref{fig:success-rate}. We will address this correlation in Section~\ref{sec:discussion}.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/SuccessfulRate}
  \caption{\label{fig:success-rate}Success rate of each task with and without GeckoGraph, grouped by experience level. The figure is cropped from 70\% to 100\% for readability. In most tasks, GeckoGraph provides a small edge. However, significant differences were found in task 10 of the beginner group and task 10 of the familiar group. }
\end{figure}

\subsection{Qualitative Feedback}
In their responses to the post-study survey, most programmers believe that the design of GeckoGraph is intuitive and that its appearance in the interface does not cause distraction.
For the question ``Do you find the GeckoGraph distracting", most of the participants rated a negative score, with an average of 2.88 (Fig. \ref{fig:qualitative} left). For the question ``How intuitive do you find the GeckoGraph?", we saw a reverse correlation of experience (Fig. \ref{fig:qualitative} Middle): experts find the GeckoGraph most intuitive (5.07), followed by the knowledgeable group (4.87), and the familiar group (4.80). The beginner group found it to be the least intuitive but still rated a positive score of (4.71). 
 When answering the question ``How helpful do you find GeckoGraph in finding the solution during the game?", the answer is more divided into different experience groups (Fig. \ref{fig:qualitative} right). It is slightly positive for beginners (4.25) and slightly negative for the other groups, the familiar group (3.86) and the knowledgeable group (3.32). The expert group finds that GeckoGraph is relatively unhelpful, with an average score of 2.95.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Qualitative}
  \caption{\label{fig:qualitative} The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph. Overall, programmers consider GeckoGraph to be intuitive and not distracting. However, opinions are split on its helpfulness. }
\end{figure}

\subsection{Threats to validity}

\paragraph{Task design}
In our human study, most of the provided functions are very abstract. These functions are created by the authors solely for the gamified study. They are designed to be different from well-known Haskell functions to minimize the familiarity variable. They are also designed with an interest of being puzzling and fun.  These functions may not be the most representative of real-world Haskell programming.

\paragraph{The use of skips}
Although we justified the use of skips in Section \ref{subsection:task}, the availability of skipping does allow users to adopt more utilitarian strategies, often involving skipping a level without giving it a fair try. This happened more often in the later levels when users realized they had enough skip opportunities left to ``complete the game". These strategies may result in lower recorded success rates than if no skips were allowed.

\section{Discussion} \label{sec:discussion}
\subsection{Strengths}
From the results of our experiment, we see that using GeckoGraph has a significant effect on the success rate of our participants, especially on less experienced programmers. We also see that with the data we collected, we did not find a significant time difference between programming with and without GeckoGraph. To extrapolate the observed expressiveness, we speculate on the practical benefits of programming with GeckoGraph.

\subsubsection{Identify the Most Important Features}
One trend that we saw from the qualitative feedback is that programmers find GeckoGraph helpful for finding patterns and important features of the types. Programmers are very positive about GeckoGraph's ability to reveal the most helpful features of a type in distinctive visual elements such as color, length, and height.

\textbf{The colors of GeckoGraph} help programmers to see the permutation of type variables in the input and output of a function. A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers. One example of GeckoGraph's effective use of color is the ``rotation" function in the user study (Fig. \ref{fig:rotate}). With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions. For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game is less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}. To quote a participant, ``GeckoGraph is quite intuitive to see the permutations of the arguments. Also, to see how to produce and consume arguments."

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/rotate}
  \caption{\label{fig:rotate} The `rotate' function in level 8 of the user study. The name given in the game is `f2'. It shuffles the type arguments of a Zero type}
\end{figure}

\textbf{The horizontal axis of GeckoGraph} often becomes intuitive when identifying differences in function arities. For example, in Fig. \ref{fig:add3}, the programmer intended to implement a function that sums 3 integers. In the implementation, the programmer missed a \texttt{(+)} function at the end; the resulting function type is largely different in length. It is also clear that the function needs to apply to one more binary function to satisfy the length requirement.

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/Add3}
  \caption{\label{fig:add3} An implementation of function \texttt{add3} but the author missed an (+) from the correct implementation (.) ((+) .) (+). GeckGraph highlights the difference in arity, and reveals that a binary function is needed on the right-hand side for the arity to match. }
\end{figure}

\textbf{The vertical axis of GeckoGraph} often sheds light on the most complex structure of this type. This can often be very useful when inspecting mismatching type errors where data types are nested. Common examples include when programmers forget to apply the value to ``return" in a monadic block or to use \texttt{liftIO} to cast an \texttt{IO} effect. For example, in Fig. \ref{fig:maybe}, the uses of \texttt{return} are excessive. It can be easily identified by examining the difference in the vertical layers of the two types. In text-based type notation, this is distinguished by different pairs of parenthesis. However, parenthesis is an overloaded syntax in type notation. In Haskell, parentheses are used to enclose tuples \texttt{(a, b)}, specify the fixity \texttt{ (a -> b) -> c}, or have no effect \texttt{a -> (b -> c)}.

To quote some feedback from participants: ``Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature. " ``It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is."

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/Maybe}
  \caption{\label{fig:maybe} The function \texttt{f} is planned to have the type \texttt{Maybe a -> a -> Maybe a}. The programmer mistakenly applied the result to the \texttt{return} function, making the result inside a Monad instance.  GeckoGraph reveals the difference in the ``layers" of types. }
\end{figure}

\subsubsection{Low barrier to learn and understand}
GeckoGraph has some key similarities to traditional text-based type notation. GeckoGraph respects the left-to-right reading order. GeckoGraph uses the familiar symbolic name as the secondary encoding. GeckoGraph simulates the prefix notation in type constructors and the infix notation in type operators. With these considerations, we ensured that programmers were able to use GeckoGraph fluently with a minimal amount of training.

One important class of feedback from the open section is that many programmers mentioned that they did not have any prior knowledge of Haskell but were able to solve the puzzles with the help of GeckoGraph.
``It is similar enough to traditional types that it is intuitive." ``This was how I parse the textual representation of types" was pointed out by multiple participants.

\subsection{Weaknesses}
\subsubsection{Space Usage}\label{subsec:space}
GeckoGraph uses horizontal space in proportion to the size of the type signature syntax tree, and GeckoGraph uses vertical space in proportion to the depth of the syntax tree. Compared to the traditional text-based language, GeckoGraph has the limitation of requiring vertical space. We have identified some approaches to minimize space usage while retaining most of the advantages of using GeckoGraph, such as displaying only the color blocks without the secondary encoding of identifier names.

\subsubsection{Color Encoding}
GeckoGraph highly relies on color hue as a main encoding. It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify subtle patterns in types, such as the order and placement of substructures. However, the perception of color is different from person to person. This becomes an even bigger issue for color-blind or visually impaired programmers. Although GeckoGraph uses color-blind friendly schemes, it is only a method to avoid indistinguishable types and is not a strong guarantee of effectiveness. For this, we are exploring different encodings, such as patterns and shapes, to maximize the accessibility of GeckoGraph.

\subsection{Gamified Human Study}

It is important to recognize that the human study is designed to be a series of puzzles. The tasks are meant to contain entertaining values. We practiced multiple gamification techniques: levels, story/theme, and goals/rewards. \cite{Hamari2014-mc} This not only allowed us to have confidence that participants are motivated to complete the tasks, it also lent us popularity in the Haskell community and led to a historically high participation rate. Gamification has been shown to improve engagement and motivation. This has been harnessed by many research projects to improve participation in human studies \cite{He2014-vp}. We identify that studies on functional programming are often technical and intimidating; our use of gamification not only attracted historically high participation, but also attracted a wide distribution of experience levels.

\subsection{Potential Applications}
\subsubsection{Programming Assistance}
We envision many ways GeckoGraph can be integrated into programming tools. GeckoGraph can visualize and inspect types in tooltips and pop-ups. It can be used to discover the mismatching parts of two conflicting types in type errors. It can be used to generate type expressions and edit type expressions structurally. In our post-study survey, the potential integration of text editors and programming assistance were the most requested use cases proposed by the participants.

\subsubsection{Documentation Assistance}
From what we have learned from our human study, GeckoGraph is well suited to support the documentation of the programming library and the API documentation. It works in tandem with the traditional text-based language and can be generated mechanically, making it possible to standardize with minimal effort. For documentation sites that allow searching by name (e.g., Hoogle \cite{Mitchell_undated-fb}), programmers often need to sieve through a list of identically named functions. For example, a simple Hoogle search for the name \texttt{make} shows a list of functions with vastly different usage and purpose. GeckoGraph can help speed up the selection process by providing a visual notation for each type, and programmers can use a visual grouping of colors, sizes, and positions to home in on the correct documentation page.

\subsubsection{Pedagogical Applications}
We believe that GeckoGraph can be a valuable tool in teaching techniques and theories in programming languages that are difficult to convey in plain language. In fact, many participants in our study reported that they had no prior knowledge of Haskell programming and that they could understand the programming concepts in the game and complete all the puzzles with the help of GeckoGraph.

Furthermore, the advanced features of GeckoGraph (Section~\ref{sec:benefits}) are also suitable for teaching and learning high-level functional programming concepts. Consider the \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension (Fig. \ref{fig:assoc}). Although the type signature is short, it is very difficult to trace the semantics mentally due to the number of variables, and their kinds are not obvious from the text-based notation. GeckoGraph makes understanding the type easier by visualizing the ``hidden" higher-kinded types, revealing all the partially applied data types in play.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/assoc}
  \caption{\label{fig:assoc} The \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension. Even for people who are not familiar with the exact definitions, it is easy to see that the variables \texttt{f},  \texttt{g}, and \texttt{h} are all high-kinded types.}
\end{figure}

% GeckoGraph, as depicted in this paper, uses a single color for each type variable and constant. In fact, to create enough visual contrast, the range of the hue spectrum may not provide enough values. In practice, GeckoGraph can encode type names using two or more colors, which will provide enough combinations.

\section{Related work}
\subsection{Visualizing polymorphic types}
A similar half-enclosing structure was proposed in the visualization of types by Jung \cite{Jung2000-oc}. In Jung's notation, the type constructor half encloses its arguments, but the figure for the type constructor is placed on the bottom right (Fig. \ref{fig:jung}).  In contrast, GeckoGraph follows the natural reading order, allowing larger structures in a type signature to take precedence over smaller ones.

Compared to functions in Jung's notation,  GeckoGraph shows two major benefits. First, GeckoGraph naturally translates the general notion of a curried function. Partially, the application of a function can be read as removing the first one of its arguments. This is not the case with Jung's notation. Second, the shape of a function type remains consistent with the shape of normal data types. In Jung's design, a function \texttt{a -> b} looks sufficiently distinct from a data type \texttt{f a b}. This is important because, in functional languages, it is very common for abstraction to be drawn from function and normal data types. For example,  functions and lists both have a functor instance, and their inner values can be altered using a \texttt{fmap} function. The consistent shape of GeckoGraph makes this generalization easier to see visually.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Jung}
  \caption{
        \label{fig:jung}
        Comparing the map function (\texttt{(a -> b) -> [a] -> [b]}) in text notation, GeckoGraph, and Jung's notation.
  }
\end{figure}

%Many studies \cite{Jun2000-ec, Jun2000-yu} explored the challenges of how human experts use polymorphic types. The authors then show a history-preserving type inference algorithm and explanation generator that can explain the types in human language. Similarly, \cite{Beaven1993-ay} showed a system that explains the infeasibility of types using human-like languages, explaining the type inference process using a series of "Why" questions and "How" questions. Both studies use natural language to reduce the challenge of understanding polymorphic types. While these studies are promising, explaining polymorphic types with natural language has limitations. They are generally verbose and have to use another visualization assist (text decoration, icons, or indentation) to clarify the naturally hierarchical information. A very similar graphical type representation \cite{Jung2000-oc} that utilizes blocks and colors was proposed. In the paper, the visualization uses 2D blocks to represent types and enclosures to represent type constructors and type arguments. We have discussed thoroughly the difference in design between Jung's notation and GeckoGraph. In addition, the evaluation of Jung's notation uses fixed questions and answers, while GeckoGraph is evaluated using real programming tasks.

\subsection{Visualization in programming}
Using visualization techniques to improve the comprehension of polymorphic types is not new. This is often practiced to represent document properties, runtime information, and static analysis results.  FluidEdt \cite{Ou2015-vr} displays heap graphs in the left margin. I3 \cite{Beck2015-my} offers search similarity and change history in compact block-based diagrams. Almeida et al. ~\cite{Almeida2022-bv} introduced a novel visualization to aid in understanding ownership and borrowing in the Rust language. While the field of graphical type representation is relatively narrow, it has been studied to some extent. Clerici et al. ~\cite{Clerici2013-ru} proposed a graph-based type inference system that shows the visual representations of unification states. GeckoGraph positions itself similarly to these projects, using color, shapes, symbols, and icons to provide easy-to-glance information. However, GeckoGraph focuses on type-level information, which no other research projects do. In addition, GeckoGraph is evaluated in a much larger study than the other projects, and GeckoGraph is evaluated with a wider range of experience levels.

\subsection{Visual vs Textual representation}

Many studies have compared the effectiveness of a visual-based programming environment with a textual-based one. Studies \cite{Noone2018-wl, Da_Silva_Ribeiro2014-tm, Cliburn2008-jo, Daly2011-is} found that compared to a purely textual programming language with similar positioning, students who were taught a visual programming language show greater confidence, better retention, and enjoyment in programming courses. While showing a similar trend, GeckoGraph experiments in the context of accompanying text-based notation rather than replacing it.

Many have studied the effect of visual augmentation, providing a visual representation of programming objects without removing the text-based notation. Greenfoot \cite{Montero2010-uh} allows visual and textual representations of programming concepts to be accessible to the learner. PILeT \cite{Alshaigy2015-wy}, providing a programming environment that is an adaptive presentation based on the user's preference. Both tools show positive results in the use of visual augmentation. Although similar to GeckoGraph in combining visual language and text-based programming environment, both studies evaluated the effect based on imperative languages (Java and Python), while our evaluation focused on the effect on a functional language (Haskell).

\section{Conclusion and Future Work}
In this paper, we propose GeckoGraph, a graphical notation for type annotations in functional programming languages. GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier. We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of. The results of the study show that GeckoGraph helps improve programmers' ability to succeed in programming tasks we designed, especially for novice programmers.

Our work in this area opens many new directions for future research.  In particular:

\noindent\textbf{In-the-wild Studies}
Although our experiment scenarios are drawn from real-world programming tasks, a certain level of variable control is still applied to remove the effect of familiarity with the tools and libraries. However, it is necessary to assess the usefulness of tools such as GeckoGraph by their real-life usage. To study this, different research methods should be used to study realistic usage and human experience. This may include field deployments or case studies.

\noindent\textbf{Alternative Type Visualization}
We strongly believe that visualization is an underutilized technique in this effort. GeckoGraph focuses on a faithful view of the tree structure of programming types. However, many more areas and types demand a human-focused approach. For instance, visualizing the ordinal relationship of subsumption or visualizing the numeric changes in dependently typed ``type programs".

\bibliographystyle{elsarticle-num} 
\bibliography{paperpile}

\appendix
\section{Game levels} \label{levels}
We provide all the level settings we used in our user study. The online game is still open source and available for evaluation \cite{Anonymous_undated-ne}. However, this can be attempted locally with a Haskell interpreter or even with a pen and paper. The target type is the desired type signature for the function \texttt{zeroToHero}. The available functions show a list of functions that are allowed to be used in the implementation. It is not required to use all the available functions, and it is not forbidden to use any other functions or variables outside the provided functions; even the Haskell prelude is not available.

\subsection{Level 1: Trial run}

\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f :: Zero a -> Hero a}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f z}
\end{itemize}

\subsection{Level 2: Assemble required}

\paragraph{Target Type} 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{runZero :: Zero a -> a}
    \item \texttt{mkHero :: a -> Hero a}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = mkHero (runZero z)}
\end{itemize}

\subsection{Level 3: Which path?}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero (a, a)}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a -> Hero a}
    \item \texttt{f2 :: Zero a -> (a, a)}
    \item \texttt{f3 :: Hero a -> Hero (a, a)}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f1 \$ z}
\end{itemize}

\subsection{Level 4: A repeating pattern}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b -> Hero b b}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b -> Hero b a}
    \item \texttt{f2 :: Zero a a -> Hero a a}
    \item \texttt{f3 :: Zero a b -> Zero b a}
    \item \texttt{f4 :: Zero a b -> Zero b b}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f2 . f4 \$ z}
\end{itemize}

\subsection{Level 5: A perfect pair}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b -> Hero b b}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{fst :: (a, b) -> a}
    \item \texttt{snd :: (a, b) -> b}
    \item \texttt{f1 :: Zero a b -> Hero b a}
    \item \texttt{f2 :: Zero a a -> Hero a a}
    \item \texttt{f3 :: Zero a b -> Zero b a}
    \item \texttt{f4 :: Zero a b -> Zero b b}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = snd .f3 . f1 \$ z}
\end{itemize}

\subsection{Level 6: Monty Hall}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c -> Hero c a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c-> Zero c b a}
    \item \texttt{f2 :: Zero a b c -> Zero a c c}
    \item \texttt{f3 :: Zero a b c -> Hero b c}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f1 . f2 \$ z}
\end{itemize}

\subsection{Level 7: TIE fighter}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c -> Hero c}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Hero (a -> b)}
    \item \texttt{f2 :: Zero a b c -> Hero (b -> c)}
    \item \texttt{f3 :: Zero a b c -> Hero a}
    \item \texttt{(<\$>) :: (a -> b) -> Hero a -> Hero b}
    \item \texttt{(<*>) :: Hero (a -> c) -> Hero a -> Hero c}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f2 z <*> (f1 z <*> f3 z)}
\end{itemize}

\subsection{Level 8: The middle man}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: (a -> d) -> (b -> d) -> (c -> d) -> Zero a b c ->  Hero a d c}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Zero c a b}
    \item \texttt{f2 :: Zero a b c -> Hero a b c}
    \item \texttt{fmap :: (c -> d) -> Zero a b c -> Zero a b d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero ad bd cd z = f2  . f1  . f1  . fmap bd  . f1 \$ z}
\end{itemize}

\subsection{Level 9: Split the difference}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c d ->  Hero d d d d}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Zero c a b}
    \item \texttt{f2 :: Zero a b c -> Hero a b c}
    \item \texttt{fmap :: (c -> d) -> Zero a b c -> Zero a b d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero ad bd cd z = f2 \$ f1 \$ f1 \$ f3 \$ z}
\end{itemize}

\subsection{Level 10: The roller coaster}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero (a -> b -> c -> d) a b c  -> Hero d}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero (a -> b) a c d -> Zero () b c d}
    \item \texttt{f2 :: Zero a b c d -> Zero b c d a}
    \item \texttt{f3 :: Zero a b c d -> Hero d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f2 . f2 . f1 . f2 . f1 . f2 . f1 \$ z}
\end{itemize}

\end{document}
\endinput

### 2025-03-29 main.tex
Text changes

@@
 %% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
+ \author{Shuai Fu}
+
+ \author{John Grundy}
+
+ \author{Peter J. Stuckey}
+
+ \author{Tim Dwyer}
+
+ \affiliation{organization={Monash University}}
+
 %% \ead{email address}
@@
 %%             country={}}
-
-\author{} %% Author name
-
-%% Author affiliation
-\affiliation{organization={},%Department and Organization
-            addressline={}, 
-            city={},
-            postcode={}, 
-            state={},
-            country={}}
 
@@
 
-\begin{figure}[]
+\begin{figure}[!h]
   \includegraphics[width=\linewidth]{figures/Design}
@@
 
-\paragraph{Data types}
+\paragraph{Complex types}
 GeckoGraph displays an algebraic data type as a larger cell, where the type constructor half encloses its arguments: \includegraphics[height=1em]{figures/DataType.png}.
@@
 Note that the height of GeckoGraph grows only upwards, but not downwards.
-Not only does this allow GeckoGraph to be more efficient in its space usage, but it also allows the legend text to be correctly aligned at the bottom and can be read similarly as regular type notation (design goal \ref{goal1}).
+Not only does this allow GeckoGraph to be more efficient in its space usage, but it also allows the legend text to be consistently aligned at the bottom and can be read as though it is a text-based type notation (design goal \ref{goal1}).
 
@@
 
-Curried functions (e.g., \texttt{ a -> b -> c}) display as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}; the second function overlaps on top of the first, indicating that the second function is the return type of the first.
+Curried functions (e.g., \texttt{ a -> b -> c}) display as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}.
+It displays as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
 Regular high-order functions (e.g., \texttt{(a -> b) -> c} ) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
@@
 Although the type signature is short, it is very difficult to trace the semantics mentally due to the number of variables, and their kinds are not obvious from the text-based notation.
-GeckoGraph makes understanding the type easier by visualizing the ``hidden" higher-kinded types, revealing all the partially applied data types in play.
+GeckoGraph makes understanding the type easier by exposing the ``hidden" higher-kinded types, revealing all the partially applied data types in play.
 
@@
   \caption{\label{fig:assoc} The \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension.
-Even for people who are not familiar with the exact definitions, it is easy to see that the variables \texttt{f},  \texttt{g}, and \texttt{h} are all high-kinded types.}
+Even for people who are not familiar with the exact definitions, it is easy to see that the variables \texttt{f},  \texttt{g}, and \texttt{h} are all high-kinded types, which is hard to tell from the text-based type annotation.}
 \end{figure}

### 2025-04-01 main.tex
Eval

@@
 %% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.3 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.3 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%% 
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-%% $Id: elsarticle-template-num.tex 249 2024-04-06 10:51:24Z rishi $
 %% 
@@
 
-%% The amssymb package provides various useful mathematical symbols
 \usepackage{amssymb}
-%% The amsmath package provides various useful equation environments.
 \usepackage{amsmath}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
 \usepackage{lineno}
-
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
 \journal{Journal of Computer Languages}
@@
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \affiliation for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
  \author{Shuai Fu}
@@
 
- \affiliation{organization={Monash University}}
-
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \affiliation{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%% \fntext[label3]{}
+ \affiliation{organization={Monash University}, country={Australia}}
 
 \title{GeckoGraph: A Visual Language for Polymorphic Types}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \affiliation[label1]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
-%%
-%% \affiliation[label2]{organization={},
-%%             addressline={},
-%%             city={},
-%%             postcode={},
-%%             state={},
-%%             country={}}
 
@@
 
-%% Add \usepackage{lineno} before \begin{document} and uncomment 
-%% following line to enable line numbers
 \linenumbers
@@
 GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
-Our study evaluating 714 participants' ability to solve type adaptation challenges in GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
+Our study evaluating 714 participants' ability to solve a series of programming challenges using GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
 The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the type challenges, especially for novice programmers.
@@
 In this paper, we use Haskell as an example.
-As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, and some advanced type-level features.
-However, GeckoGraph could also be used in imperative and multiparadigm languages such as Rust.
-We provide a GeckoGraph construction library for Haskell.
+As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, and many advanced type-level features.
+However, GeckoGraph could also be used in imperative and multiparadigm languages such as TypeScript and Rust.
+We provide a GeckoGraph visualization tool avaiable online for multiple languages \todo{cite GeckoGraph website}.
 
-We identified three main design goals for GeckoGraph based on the challenges of using polymorphic types~\cite{Jun2000-ec, Jun2000-yu} and how programmers tend to use type annotations~\cite{Justin_Lubin2021-yy}, as follows.
+We identified three main design goals for GeckoGraph based on the challenges of using polymorphic types~\cite{Jun2000-ec, Jun2000-yu} and how programmers tend to use type annotations to support their programming tasks~\cite{Justin_Lubin2021-yy}, as follows.
 
-\paragraph{\textbf{(D1) Low barrier to learn}}\label{goal1} GeckoGraph should take little to no effort to learn.
+\paragraph{\textbf{(D1) Easy to learn}}\label{goal1} GeckoGraph should take little to no effort to learn.
 The rules to translate a text-based type notation to GeckoGraph should be minimal.
@@
 
-\paragraph{\textbf{(D3) Easy to compare and search}} \label{goal3} GeckoGraph should aim to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss.
+\paragraph{\textbf{(D3) Optimized for comparing and searching}} \label{goal3} GeckoGraph should aim to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss.
 This also includes the task of choosing an ideal function from a list of potential functions.
-For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity, one of the argument types, or type class it must fulfill).
+For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity of a function, containing a known type, or type classes requirements).
 
@@
 Each type identifier encodes a distinct color hue of the cell.
-Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and strong secondary encoding.
+Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and a strong secondary encoding.
 The angled corner in the top left provides visual separation between two cells, even when the same color cells are next to each other, allowing GeckoGraph to be zoomed out to extremely small sizes (Section \ref{subsec:space}) without suffering readability (design goal \ref{goal2}).
@@
         \label{fig:indicator}
-        An example of using the function indicator.
+        An example of visual hirachy of the function indicator.
 The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators.
-For high-order functions where functions are arguments of other functions, it is very easy to see the ``order" of functions and how they are arranged.
+For higher-order functions where functions are arguments or return values of other functions, it is very easy to see the ``level" of functions.
   }
@@
 Each type class required on a type variable displays as a square indicator aligned on the right of the extended area.
-In GeckoGraph, type-class constraints are associated with every instance of the type variable that requires them.
-This means when displaying the type \texttt{(==) :: Eq a => a -> a -> Bool} in GeckoGraph, the constraint \texttt{Eq} appears in both occurrences of \texttt{a}.
+In GeckoGraph, a type-class gliph is embeded in every type variable that requires this  class.
+This means when displaying the type \texttt{(==) :: Eq a =>
+a -> a -> Bool} in GeckoGraph, the constraint \texttt{Eq} appears in twice on both occurrences of the type variable \texttt{a}.
 
@@
 Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or pop-up window) to identify the name of the type class.
-We believe that this decision is well justified.
-For example, when reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable.
-GeckoGraph helps minimize the effort to associate each type variable with all its type classes (design goal \ref{goal2}).
+We believe that this is a well justified trade-off.
+The strength of this appraoch lies in situations when type class context is complex.
+When reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable.
+GeckoGraph uses approximity to help minimize the effort to associate each type variable with all its type classes (design goal \ref{goal2}).
 
-\subsection{Benefits of Using GeckoGraph}\label{sec:benefits}
+\subsection{Strengths of GeckoGraph}\label{sec:benefits}
+
 \paragraph{Generalization patterns in type classes}
-A frequently cited confusion among novices is the overlap between list type \texttt{[a]} and its more generic counterpart \texttt{Foldable} instances.
-The subtle differences are often not important for beginners.
-However, when encountering type errors in working with lists, Haskell often explains the error with the \texttt{Foldable} instance.
-For the example in Fig.
-\ref{fig:example-foldable} where a programmer mistakenly provided a \texttt{Char} instead of a \texttt{String} literal, the compiler shows an error message comparing the provided \texttt{Char} type with a confusingly named type \texttt{t0 a0}.
-Although any \texttt{Foldable} instance is perfectly suited for the list \cite{Waldmann2018-hu}, this generalization may reduce programmers' confidence in their understanding of the language and the ability to navigate out of a type error.
-In GeckoGraph, a list type and a type with a Foldable instance have the same shape.
-This allows the generic type \texttt{t0 a0} in the error message to assume the same shape as \texttt{[a]}.
-This generalization of concrete types and abstraction of type-class instances aims to allow for teaching fundamental functional programming concepts without hiding high-level abstractions.
-The same benefits apply to polymorphic numbers and strings.
+A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generalized types and those that work with concrete types \cite{Tirronen2015-nr}.
+For instance, the type constraints {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, as well as {\tt Foldable} instances versus {\tt []} can be particularly challenging.
+Many Haskell learners may find the differences between concrete types and their generic counterparts irrelevant to their programming tasks, thereby complicating their quest for solutions.
+In the example shown in Fig.
+\ref{fig:example-foldable}, a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String} literal.
+The compiler responds with an error message that compares the provided \texttt{Char} type to a confusingly named type \texttt{t0 a0}.
+In this case, the information has been generalized too much and the programmer can hardly find their bearing to the original problem.
+This level of generalization can undermine a programmer's confidence in understanding the language and navigating type errors.
+In GeckoGraph, both list types and types with a Foldable instance share a similar structure.
+Therefore, the type \texttt{t0 a0} in the error message matches the shape of \texttt{[a]}.
+Bring programmer closer to the real issue, trying to match a complex type with a concrete simple type.
+This feature lowers the entry barrier for languages like Haskell, where even basic arithmetic functions {\tt +} and {\tt -} are defined on generic types.
+This approach lowers the barrier to introduce foundational language concepts without sacrificing generality.
 
@@
 The design of GeckoGraph enables the visualization of many advanced type-level features.
-\textbf{Kind visualization}: if the kind of type variables can be inferred, the kind information is consistently displayed in GeckoGraph.
-For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} needs at least the kind \texttt{* -> *} because of its use on the right-hand side.
-GeckoGraph respects this kind information and displays it as a constructor type over an empty structure, indicated using a dotted outline.
+\textbf{Kind visualization}: Kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is type constructor (e.g.
+[], Maybe) or a type constant (e.g.
+Bool, Char), because of this, kind is often described as the type of types.
+Goanna infer the kind of type variables,  consistently display this information in GeckoGraph.
+For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} is infered to have the kind \texttt{* -> *} because of its use on the right-hand side.
+GeckoGraph respects this kind information and displays the same kind level on the first occurance of \texttt{a}, but over an empty structure, indicated using a dotted outline.
 \textbf{Qualified constraints}: GeckoGraph's type class notation naturally extends to support qualified constraints.
@@
 Because of its graphical layout, GeckoGraph allows programmers to precisely select which part of a type signature they intend to query, that is, in Fig.
-\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail.
+\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a sepecific query that is hard to hard to execute with point and click over text based type annotation.
 
@@
 \paragraph{Language Agnostic}
-GeckoGraph can be implemented in any language that uses static typing.
-In programming projects, GeckoGraph supports polyglot programming projects.
-Typical circumstances include projects using foreign function interfaces or multiple languages for client- and server-side programming.
-GeckoGraph provides a common notation to describe the functionality and features of systems.
+GeckoGraph can be implemented in any language that exercises static typing.
+We identified applications of this feature in programming and education.
+In programming projects often use more than one programming languages, situations include projects using foreign function interfaces or multiple languages for client-side and server-side programming.
+In these cases, GeckoGraph provides a shared notation to describe the functionality and features of systems, dispide their text-based notations are different.
 In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages.
-For example, when describing algebraic data types, different language communities use various names: tuple, enum, struct, etc.
-It is important to realize that these are the same concepts and ignore the minute linguistic barriers.
+For example, when describing the product type in algebraic data types, different language communities use various names: record, named tuple, struct, etc.
+GechoGraph helps to unify the same programming concept and ignore the minute nomenclature differences .
 
@@
 \section{Evaluation}
-To evaluate the usefulness of GeckoGraph, we designed a controlled experiment in the form of a game called "Zero to Hero".
+To evaluate the usefulness of GeckoGraph, we designed and conducted a controlled experiment in the form of a game called "Zero to Hero".
 The game contains 10 levels of varying difficulty.
-At each level, participants are asked to implement a function called "\texttt{zeroToHero}" using only a list of available functions.
-These available functions are different at each level, and the target types of Zero and Hero vary at each level.
+At each level, participants are asked to implement a function named "\texttt{zeroToHero}" with the help of a limited list of provided functions.
+These functions are different at each level, and the target types of Zero and Hero vary at each level as well.
 The details of each level are provided in the Appendix (Appendix \ref{levels}).
 
-The experiment aims to study how polymorphic types are used and reasoned about during programming tasks.
+The experiment aims to study how programmers use and reason about polymorphic type when performing programming tasks.
 In particular, we studied how programmers scan and select potentially useful functions from a library and compare intended types and actual types during type errors.
@@
 The available functions are \texttt{f1::Zero a b -> Hero b a}, \texttt{f2::Zero a a -> Hero a a}, \texttt{f3::Zero a b -> Hero b a}, and \texttt{f4::Zero a b -> Hero b b}.
-Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be won without the use of generic functions.
+Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions.
 The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
@@
 
-At each level of the game, the programmer must select the right functions to achieve the target result.
+At each level of the game, the programmer must select the right set of functions, and compose them correctly to achieve the target result.
+Not all provided functions are useful.
 In particular, for this level, participants must discover that only \texttt{f4} and \texttt{f2} are necessary to produce the desired results.
-An implement that satisfies the target type is \texttt{zeroTohero z = f2  (f4  z)}.
+One implementation that satisfies the target type is \texttt{zeroTohero z = f2 (f4  z)}.
 
@@
 
-The experiment uses a between-subject design. However, all participants receive both treatments (with and without GeckoGraph) during their runs.
+The experiment uses a between-subject design.
+However, all participants receive both treatments (with and without GeckoGraph) during their runs.
 Participants are assigned to one of two groups.
@@
 \subsection{Hypothesis}
-In programming tasks that involve reading and understanding polymorphic types, graphic notation using visual elements that provide high grouping strength (colors, shapes, sizes, and symbols) can improve the performance of such tasks compared to traditional text-based type notation.
-Our null hypothesis is that "Using graphic notation has no effect compared to traditional text-based type notation." This hypothesis and the task design were registered at the Open Science Foundation prior to data collection.
+In programming tasks that involve reading and understanding polymorphic types, graphic notation using visual elements that provide higher grouping strength (colors, shapes, sizes, and symbols) can improve the performance of such tasks compared to traditional text-based type notation.
+Our null hypothesis is: {\it Using graphic notation has no effect compared to traditional text-based type notation.} This hypothesis and the task design were registered at the Open Science Foundation prior to data collection \todo{cite this}.
 
@@
 Participants can skip a level during the game if they are stuck.
-We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their qualitative input of the tool.
-We limit the number of skips that a participant can use during the game to four times so that submitting qualitative feedback without completing at least some levels is impossible.
+We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their feedback of the tool.
+We limit the number of skips that a participant can use during the game to four so that submitting qualitative feedback without completing at least some levels is impossible.
 
@@
 
+The level 10 is the most cognatively taxing: the provided functions are very long and complex, and the solution is very challenging and deceptive, programmers have to utilize all the provided functions, and compese them eight time to reach the final position.
+Because level 10 require programmers to switch back and forth to remember what each function does.
+We speculate that this level require the mental agility of a Haskell expert to internalize these functions, and GeckoGraph help less experienced programmers achieve the similar level efficacy.
+
 \begin{figure}[]
   \includegraphics[width=\linewidth]{figures/SuccessfulRate}
-  \caption{\label{fig:success-rate}Success rate of each task with and without GeckoGraph, grouped by experience level.
-The figure is cropped from 70\% to 100\% for readability.
+  \caption{\label{fig:success-rate}{\bf Success rate of each task with and without GeckoGraph, grouped by experience level.} The figure is cropped from 70\% to 100\% for readability.
 In most tasks, GeckoGraph provides a small edge.
@@
 
+It is interesting to note that programmers find GeckoGraph intuitive, and the more experienced they are, the more likely they find GeckoGraph easy to use.
+We believe this aligns with our design decision to make GeckoGraph resemble text-based type annotation.
+Those familiar with text-based type annotation will find GeckoGraph second nature to read.
+However, the correlation reverses when it comes to perceived helpfulness.
+The more experienced the users are, the less likely they are to find GeckoGraph helpful.
+We speculate this pattern occurs because users who are experts in text-based type annotations naturally prefer reading text-based notation and tend to ignore GeckoGraph, diminishing its usefulness.
+Conversely, novice users are drawn to GeckoGraph and find its visualization offers more clarity than text-based notations.
+
 \begin{figure}[]
   \includegraphics[width=\linewidth]{figures/Qualitative}
-  \caption{\label{fig:qualitative} The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph.
-Overall, programmers consider GeckoGraph to be intuitive and not distracting.
+  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph.} Overall, programmers consider GeckoGraph to be intuitive and not distracting.
 However, opinions are split on its helpfulness.
@@
 GeckoGraph simulates the prefix notation in type constructors and the infix notation in type operators.
-With these considerations, we ensured that programmers were able to use GeckoGraph fluently with a minimal amount of training.
+With these considerations, we ensured that programmers who were familiar with the Haskell language had no problem understanding and using GeckoGraph fluently with a minimal amount of training.
 
@@
 GeckoGraph highly relies on color hue as a main encoding.
-It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify subtle patterns in types, such as the order and placement of substructures.
+It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify patterns in types, such as the order and placement of substructures.
 However, the perception of color is different from person to person.
-This becomes an even bigger issue for color-blind or visually impaired programmers.
+This becomes an even bigger issue for  visually impaired programmers.
 Although GeckoGraph uses color-blind friendly schemes, it is only a method to avoid indistinguishable types and is not a strong guarantee of effectiveness.
-For this, we are exploring different encodings, such as patterns and shapes, to maximize the accessibility of GeckoGraph.
+For this, we are exploring different visual encoding mehods, such as patterns and shapes, to maximize the accessibility of GeckoGraph.
 
@@
 \subsubsection{Pedagogical Applications}
+
 We believe that GeckoGraph can be a valuable tool in teaching techniques and theories in programming languages that are difficult to convey in plain language.
@@
 \end{figure}
-
-% GeckoGraph, as depicted in this paper, uses a single color for each type variable and constant.
-In fact, to create enough visual contrast, the range of the hue spectrum may not provide enough values.
-In practice, GeckoGraph can encode type names using two or more colors, which will provide enough combinations.
 

### 2025-04-04 main.tex
Some changes in eval

@@
 
-The level 10 is the most cognatively taxing: the provided functions are very long and complex, and the solution is very challenging and deceptive, programmers have to utilize all the provided functions, and compese them eight time to reach the final position.
-Because level 10 require programmers to switch back and forth to remember what each function does.
-We speculate that this level require the mental agility of a Haskell expert to internalize these functions, and GeckoGraph help less experienced programmers achieve the similar level efficacy.
+Level 10 is the most cognitively demanding: the provided functions are the lengthiest, and the solution is challenging and deceptive.
+Programmers must use the provided rotate function {\it f2} to adjust the positions of the four components of the {\it Zero} value four time to get to the correct position.
+Completing the level requires applying the functions {\it f1}, {\it f2}, and {\it f3} in a nested sequence that is 8 layers deep.
+Consequently, programmers need to frequently switch contexts to recall the purpose of each function.
+We speculate that this level demands the mental flexibility typical of a Haskell expert to quickly internalize these functions.
+GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding structual clarity.
 
@@
 
-It is interesting to note that programmers find GeckoGraph intuitive, and the more experienced they are, the more likely they find GeckoGraph easy to use.
+Two interesting trends emerge.
+First, the more experienced the users are, the more likely they find GeckoGraph intuitive and easy to use.
 We believe this aligns with our design decision to make GeckoGraph resemble text-based type annotation.
-Those familiar with text-based type annotation will find GeckoGraph second nature to read.
-However, the correlation reverses when it comes to perceived helpfulness.
+Those familiar with text-based type annotation find GeckoGraph second nature to read.
+Second, the correlation reverses when it comes to perceived helpfulness.
 The more experienced the users are, the less likely they are to find GeckoGraph helpful.
 We speculate this pattern occurs because users who are experts in text-based type annotations naturally prefer reading text-based notation and tend to ignore GeckoGraph, diminishing its usefulness.
-Conversely, novice users are drawn to GeckoGraph and find its visualization offers more clarity than text-based notations.
+Conversely, novice users are more likely to be drawn to GeckoGraph and find its visualization offers more clarity and useful insights.
 

### 2025-04-06 main.tex
Citations

@@
 
+ \author{Peter J. Stuckey}
+
  \author{John Grundy}
-
- \author{Peter J. Stuckey}
 
@@
 
-Although polymorphic typing promises robustness and a high degree of code reusability, studies~\cite{Jun2000-ec, Jun2000-yu} show that using polymorphism in practice poses challenges, especially for novice users.
-These studies have shown that humans tend to focus on concrete types and only rely on polymorphic type checking as a last resort.
-In practice, polymorphic types often pose usability problems for programmers.
-New polymorphic type variables can be created during type checking.
-These intermediate type variables are often kept behind the curtain unless a type error is encountered.
-This often results in programmers resolving type errors with type variables not authored by the programmers themselves (Fig.
+While promoting robustness and code reusability, polymorphic types present challenges in comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}.
+Research indicates that humans tend to focus on concrete types and rely on polymorphic type checking only as a last resort.
+Moreover, polymorphic types often complicate type error usability.
+During type checking, compilers frequently generate new polymorphic type variables temporarily.
+These variables are typically discarded after type checking unless a type error occurs, in which case they are referenced in error messages.
+Consequently, programmers often face the task of resolving type errors involving type variables they did not explicitly author (Fig.
 \ref{fig:example-foldable}).
@@
 While type polymorphism is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}, little research focuses on the \textit{usability} of polymorphic types.
-Hage argues that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}.
+It is often argued that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}.
 We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI techniques.
-To achieve this, we propose GeckoGraph, a graphical notation for types.
+In this paper, we propose GeckoGraph, a graphical notation for types.
 GeckoGraph aims to complement traditional text-based type notation and make reading, understanding, and comparing types easier.
@@
 However, GeckoGraph could also be used in imperative and multiparadigm languages such as TypeScript and Rust.
-We provide a GeckoGraph visualization tool avaiable online for multiple languages \todo{cite GeckoGraph website}.
+We provide a GeckoGraph visualization tool avaiable online for multiple languages \cite{Fu2025-ka}.
 
@@
 The experiment uses a between-subject design.
-However, all participants receive both treatments (with and without GeckoGraph) during their runs.
 Participants are assigned to one of two groups.
-Both groups receive the same tasks in the same order.
-Group one participants are assisted by GeckoGraph on even levels and only text-based type annotation on odd levels.
+The tasks asiigned to the two groups are the same order.
+Participants receive both treatments (with and without GeckoGraph) during their runs but in different order.
+Group one is assisted by GeckoGraph on even levels and text-only type annotation on odd levels.
 Group two participants are the same but with the order flipped.
-Both groups have access to the text-based type signature for all tasks.
 The number of participants in the two groups is counterbalanced.
@@
 In programming tasks that involve reading and understanding polymorphic types, graphic notation using visual elements that provide higher grouping strength (colors, shapes, sizes, and symbols) can improve the performance of such tasks compared to traditional text-based type notation.
-Our null hypothesis is: {\it Using graphic notation has no effect compared to traditional text-based type notation.} This hypothesis and the task design were registered at the Open Science Foundation prior to data collection \todo{cite this}.
+Our null hypothesis is: {\it Using graphic notation has no effect compared to traditional text-based type notation.} This hypothesis and the task design were registered at the Open Science Foundation prior to data collection \cite{Stuckey2023-vl}.
 
 \subsection{Task Design} \label{subsection:task}
-Participants in both groups receive the same 10 tasks.
+During the study, participants receive 10 tasks, represented by the 10 levels of the game {\it zeroToHero}.
 The tasks start off easy but gradually increase in difficulty.
-In each task, a target type signature of the function \texttt{zeroToHero} is given to the participants.
-Participants are provided with a list of available functions to implement the target function.
+In each task, a target type of the function \texttt{zeroToHero} is shown to the participants.
+Participants are provided with a list of available functions to implement the target function, and to ensure it assume the target type.
 This is to simulate the tasks of selecting useful functions from a library.
-In addition, participants are not allowed to use any other functions or variables outside the provided functions; even the Haskell prelude is not available.
+Participants are not allowed to use any other functions or variables outside the provided functions; even the Haskell prelude is not available.
 This ensures that everyone has the same knowledge and minimizes the effect of familiarity.
@@
 
-Data collection from the human study was stopped after the planned cut-off period of 14 days.
+Data collection from the study was stopped after the planned cut-off period of 14 days.
 After the cut-off date, the ZeroToHero game is open source and available for free evaluation \cite{Anonymous_undated-ne} and repeating our experiment, but no further data was collected.

### 2025-04-08 main.tex
keywords
changes from john

@@
 
+ \author{Tim Dwyer}
+
  \author{Peter J. Stuckey}
@@
  \author{John Grundy}
-
- \author{Tim Dwyer}
 
@@
 %% or \MSC[2008] code \sep code (2000 is the default)
-
+Polymorphic Types \sep Functional Programming \sep Visual Languages \sep Visualization
 \end{keyword}
@@
 \section{Introduction} \label{sec:intro}
-In programming languages, a polymorphic type \cite{Cardelli1987-fp} can represent values of different types while providing a interface to describe common behaviors for those values.
-Polymorphic types are central to the succinctness of contemporary statically-typed functional languages, enabling a considerable degree of type-safe abstraction and hence reusability.
-Parametric polymorphism, a common technique to denote polymorphic type, is available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}.
+In programming languages, a polymorphic type \cite{Cardelli1987-fp} can represent values of different types while providing an interface to describe common behaviors for those values.
+Polymorphic types are central to the succinctness of contemporary statically-typed languages, enabling a considerable degree of type-safe abstraction and reusability.
+Parametric types are available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}.
 Polymorphism allows programs to be written in a way that is more generic and adaptable to different data types, enabling greater flexibility and code reuse.
@@
 
-While type polymorphism is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}, little research focuses on the \textit{usability} of polymorphic types.
+Polymorphic type is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}.
+However, little research focuses on the \textit{usability} of polymorphic types.
 It is often argued that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}.
@@
 Our study evaluating 714 participants' ability to solve a series of programming challenges using GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
-The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the type challenges, especially for novice programmers.
+The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the programming challenges, especially for novice programmers.
 
@@
 GeckoGraph is a visual notation for type annotations in statically typed programming languages.
-It is intended to work tangibly with text-based annotations, but uses colors, shapes, and symbols to make structures of types easy to identify at a glance.
+It is intended to work tangibly with text-based annotations but uses colors, shapes, and symbols to make structures of types easy to identify at a glance.
 In this section, we describe the design of GeckoGraph and highlight some unique benefits of programming with GeckoGraph.
@@
 
-\paragraph{\textbf{(D2) Easy to parse for humans}}  \label{goal2} GeckoGraph should make the task of reading and understanding type notation easy.
-It should emphasize the less obvious properties of a type signature.
-GeckoGraph should eliminate the need for mental backtracking, such as counting opening and closing parentheses and remembering which type classes are required on which variables.
-
-\paragraph{\textbf{(D3) Optimized for comparing and searching}} \label{goal3} GeckoGraph should aim to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss.
-This also includes the task of choosing an ideal function from a list of potential functions.
-For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity of a function, containing a known type, or type classes requirements).
-
-\paragraph{Simple Types} 
-Simple types, such as type variables and concrete types, are displayed in a cell \includegraphics[height=1em]{figures/SimpleType}: a solid-colored rectangular box with an angled corner on the top left.
-Each type identifier encodes a distinct color hue of the cell.
-Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and a strong secondary encoding.
-The angled corner in the top left provides visual separation between two cells, even when the same color cells are next to each other, allowing GeckoGraph to be zoomed out to extremely small sizes (Section \ref{subsec:space}) without suffering readability (design goal \ref{goal2}).
-
 \begin{figure}[!h]
@@
 \end{figure}
+
+\paragraph{\textbf{(D2) Easy to parse for humans}}  \label{goal2} GeckoGraph should make the task of reading and understanding type notation easy.
+It should emphasize the less obvious properties of a type signature.
+GeckoGraph should eliminate the need for mental backtracking, such as counting opening and closing parentheses and remembering which type classes are required on which variables.
+
+\paragraph{\textbf{(D3) Optimized for comparing and searching}} \label{goal3} GeckoGraph aims to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss.
+This also includes the task of choosing an ideal function from a list of potential functions.
+For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity of a function, containing a known type, or type class requirements).
+
+\paragraph{Simple Types} 
+Simple types, such as type variables and concrete types, are displayed in a cell \includegraphics[height=1em]{figures/SimpleType}: a solid-colored rectangular box with an angled corner on the top left.
+Each type identifier encodes a distinct color hue of the cell.
+Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and a strong secondary encoding.
+The angled corner in the top left provides visual separation between two cells, even when the same color cells are next to each other, allowing GeckoGraph to be zoomed out to extremely small sizes (Section \ref{subsec:space}) without suffering readability (design goal \ref{goal2}).
 
@@
 
-Curried functions (e.g., \texttt{ a -> b -> c}) display as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}.
-It displays as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
+Curried functions (e.g., \texttt{ a -> b -> c}) are displayed as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}.
+It appears as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
 Regular high-order functions (e.g., \texttt{(a -> b) -> c} ) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
@@
         \label{fig:indicator}
-        An example of visual hirachy of the function indicator.
+        An example of visual hierarchy of the function indicator.
 The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators.
@@
 Each type class required on a type variable displays as a square indicator aligned on the right of the extended area.
-In GeckoGraph, a type-class gliph is embeded in every type variable that requires this  class.
+In GeckoGraph, a type-class gliph is embedded in every type variable that requires this class.
 This means when displaying the type \texttt{(==) :: Eq a =>
@@
 Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or pop-up window) to identify the name of the type class.
-We believe that this is a well justified trade-off.
-The strength of this appraoch lies in situations when type class context is complex.
+We believe that this is a well-justified trade-off.
+The strength of this approach lies in situations when type class context is complex.
 When reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable.
@@
 
-\subsection{Strengths of GeckoGraph}\label{sec:benefits}
+\subsection{Key Design Criteria of GeckoGraph}\label{sec:benefits}
 
-\paragraph{Generalization patterns in type classes}
-A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generalized types and those that work with concrete types \cite{Tirronen2015-nr}.
-For instance, the type constraints {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, as well as {\tt Foldable} instances versus {\tt []} can be particularly challenging.
-Many Haskell learners may find the differences between concrete types and their generic counterparts irrelevant to their programming tasks, thereby complicating their quest for solutions.
-In the example shown in Fig.
-\ref{fig:example-foldable}, a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String} literal.
-The compiler responds with an error message that compares the provided \texttt{Char} type to a confusingly named type \texttt{t0 a0}.
-In this case, the information has been generalized too much and the programmer can hardly find their bearing to the original problem.
-This level of generalization can undermine a programmer's confidence in understanding the language and navigating type errors.
-In GeckoGraph, both list types and types with a Foldable instance share a similar structure.
-Therefore, the type \texttt{t0 a0} in the error message matches the shape of \texttt{[a]}.
-Bring programmer closer to the real issue, trying to match a complex type with a concrete simple type.
-This feature lowers the entry barrier for languages like Haskell, where even basic arithmetic functions {\tt +} and {\tt -} are defined on generic types.
-This approach lowers the barrier to introduce foundational language concepts without sacrificing generality.
+GeckoGraph was designed through many different iterations.
+Many research methods were used to verify ideas, including prototyping, cognitive walk-throughs, and formative studies.
+We discuss some notable visual dimensions in GeckoGraph and their justifications.
 
-\paragraph{Consistent color scheme}\label{par:color-scheme}
-A common task in programming is to scan for a desired function from a sea of potentially useful functions, such as library documentation.
-During scanning, programmers often have partial knowledge of the desired function, e.g., the arity, one of the argument types, or the type class it must fulfill.
-A typical example is conversion: using a known \texttt{String} type to produce a desired \texttt{Data.Text}  type.
-Another example is the `lookup' function: using a known \texttt{Data.Map a b} to produce a desired \texttt{b} type.
-GeckoGraph supports this task by using consistent colors for the same type identifier.
-Programmers can rely on the color grouping to scan for the desired type in their project or in third-party library documentation.
+\begin{figure}[]
+  \includegraphics[width=\linewidth]{figures/PreviousVersions}
+  \caption{\label{fig:previous}Previous Versions of GeckoGraph. Different encodings represent named types, type variables, type constructors, and high-order functions.}
+\end{figure}
 
-\paragraph{Advanced Type Feature Visualization}
+\paragraph{Visual Dimensions of GeckoGraph}
+\textbf{The colors of GeckoGraph} help programmers to see the permutation of type variables in the input and output of a function.
+A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
+One example of GeckoGraph's use of color is the ``rotation" function  (Fig.
+\ref{fig:rotate}) in the user study (Section \ref{sec:evaluation}).
+With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions.
+For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game is less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.
+
+\begin{figure}[]
+  \includegraphics[width=0.6\linewidth]{figures/rotate}
+  \caption{\label{fig:rotate} The `rotate' function in level 8 of the user study. The name given in the game is `f2'. It shuffles the type arguments of a Zero type}
+\end{figure}
+
+\textbf{The horizontal axis of GeckoGraph} often are helpful in identifying differences in function arities.
+For example, in Fig.
+\ref{fig:add3}, the programmer intended to implement a function that sums 3 integers.
+In the implementation, the programmer missed a \texttt{(+)} function at the end; the resulting function type is largely different in length.
+It is also clear that the function needs to apply to one more binary function to satisfy the length requirement.
+
+\begin{figure}[]
+  \includegraphics[width=0.6\linewidth]{figures/Add3}
+  \caption{\label{fig:add3} An implementation of function \texttt{add3} but the author missed an (+) from the correct implementation (.) ((+) .) (+).
+GeckGraph highlights the difference in arity and reveals that a binary function is needed on the right-hand side for the arity to match.
+}
+\end{figure}
+
+\textbf{The vertical axis of GeckoGraph} often helps identify the number of levels in a nested complex type.
+This can often be very useful when inspecting mismatching type errors where data types are nested in different levels.
+Common examples include when programmers forget to apply the value to ``return" in a monadic block or to use \texttt{liftIO} to cast an \texttt{IO} effect.
+In Fig.
+\ref{fig:maybe}, the uses of \texttt{return} are excessive.
+It can be easily identified by examining the difference in the vertical layers of the two types.
+In text-based type notation, this is distinguished by different pairs of parentheses.
+However, parenthesis is an overloaded syntax in type notation.
+In Haskell, parentheses are used to enclose tuples \texttt{(a, b)}, specify the fixity \texttt{ (a -> b) -> c}, or have no effect \texttt{a -> (b -> c)}.
+
+\begin{figure}[]
+  \includegraphics[width=0.6\linewidth]{figures/Maybe}
+  \caption{\label{fig:maybe} The function \texttt{f} is planned to have the type \texttt{Maybe a -> a -> Maybe a}.
+The programmer mistakenly applied the result to the \texttt{return} function, making the result inside a Monad instance.
+GeckoGraph reveals the difference in the ``layers" of types.
+}
+\end{figure}
+
+\paragraph{Support Advanced Type Features}
 The design of GeckoGraph enables the visualization of many advanced type-level features.
@@
 Bool, Char), because of this, kind is often described as the type of types.
-Goanna infer the kind of type variables,  consistently display this information in GeckoGraph.
-For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} is infered to have the kind \texttt{* -> *} because of its use on the right-hand side.
-GeckoGraph respects this kind information and displays the same kind level on the first occurance of \texttt{a}, but over an empty structure, indicated using a dotted outline.
+GeckoGraph infers the {\it kind} of type variables and consistently displays this information in GeckoGraph.
+For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} is inferred to have the kind \texttt{* -> *} because of its use on the right-hand side.
+GeckoGraph respects this {\it kind} information and displays the same kind level on the first occurrence of \texttt{a}, but over an empty structure, indicated using a dotted outline.
 \textbf{Qualified constraints}: GeckoGraph's type class notation naturally extends to support qualified constraints.
@@
 
-\paragraph{Precise Interactivity}
-Modern programming environments often allow programmers to mouse over part of the source code to query detailed information, such as definition, references or documentation.
+\paragraph{Provide Precise Interactivity}
+Modern programming environments often allow programmers to mouse over part of the source code to query detailed information, such as definition, references, or documentation.
 However, with text-based source code, it is often hard to distinguish whether programmers want the most specific fragment under the cursor or larger blocks.
@@
 \end{figure}
-
-\paragraph{Language Agnostic}
-GeckoGraph can be implemented in any language that exercises static typing.
-We identified applications of this feature in programming and education.
-In programming projects often use more than one programming languages, situations include projects using foreign function interfaces or multiple languages for client-side and server-side programming.
-In these cases, GeckoGraph provides a shared notation to describe the functionality and features of systems, dispide their text-based notations are different.
-In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages.
-For example, when describing the product type in algebraic data types, different language communities use various names: record, named tuple, struct, etc.
-GechoGraph helps to unify the same programming concept and ignore the minute nomenclature differences .
-
-\subsection{Previous iterations of GeckoGraph}
-GeckoGraph was designed through many different iterations.
-Many research methods were used to verify ideas, including prototyping, cognitive walk-throughs, and formative studies.
-We list some notable design elements and their major feedback.
-
-\begin{figure}[]
-  \includegraphics[width=\linewidth]{figures/PreviousVersions}
-  \caption{\label{fig:previous}Previous Versions of GeckoGraph. Different encodings represent named types, type variables, type constructors, and high-order functions.}
-\end{figure}
-
-\paragraph{\textbf{Encoding the direction of function types}}
-This can be found in the design v1 (Fig.
-\ref{fig:previous}).
-This design shares many similarities with a prior project \cite{Jung2000-oc} in type visualization.
-Although many students and experts agree on this, its use of space on both width and height scales is proportional to the types in question, causing too much inconvenience.
-In addition, this design does not produce a canonical form for a curried function.
-For example, for composition (.), it is unclear whether it takes two functions as input and returns a binary function or two functions and a single structure as input and returns a different structure.
-
-\paragraph{Encoding the depth and size} Grayscales, gradience, size, and simulated 3-dimensional elevation are promising visual representations for depicting numeric dimensions such as the depth and width of the parse tree.
-However, gradient and elevation were dismissed because of their requirements for a more demanding rendering process, making GeckoGraph harder to implement in more restricted user interfaces, such as the command line.
-In addition, all of these visual dimensions reduce readability when scaled down to a very small size.
-
-\paragraph{Encoding symbolic names} In some variations, we tested using different geometric shapes to indicate the symbolic names of type variables and concrete types.
-We decided against using icons due to the limited number of different shapes until they were indistinguishable.
-The color provides more encoding spaces, and the letters provide familiarity with the original type annotations.
-This was shown to help reduce friction in the adoption of GeckoGraph.
 
@@
 
-At each level of the game, the programmer must select the right set of functions, and compose them correctly to achieve the target result.
+At each level of the game, the programmer must select the right set of functions and compose them correctly to achieve the target result.
 Not all provided functions are useful.
@@
 Participants are assigned to one of two groups.
-The tasks asiigned to the two groups are the same order.
-Participants receive both treatments (with and without GeckoGraph) during their runs but in different order.
+The tasks assigned to the two groups are in the same order.
+Participants receive both treatments (with and without GeckoGraph) during their runs but in a different order.
 Group one is assisted by GeckoGraph on even levels and text-only type annotation on odd levels.
@@
 In each task, a target type of the function \texttt{zeroToHero} is shown to the participants.
-Participants are provided with a list of available functions to implement the target function, and to ensure it assume the target type.
+Participants are provided with a list of available functions to implement the target function and to ensure it assumes the target type.
 This is to simulate the tasks of selecting useful functions from a library.
@@
 
-The 10 levels are designed to gradually increase difficulty.
+The 10 levels are designed to increase difficulty gradually.
 From the results of the experiment, most of the tasks align with this trend.
@@
 
-Level 5, 7 and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience.
+Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience.
 Level 3 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple.
-Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 the \texttt{fmap} function of the \texttt{Functor} class.
+Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class.
 The authors speculate that the more experienced participants are much more familiar with these functions, hence the strong contrast on these three levels.
@@
 96.
-06\%) and expert group (98.
+06\%), and expert group (98.
 2\% vs.
@@
 Level 10 is the most cognitively demanding: the provided functions are the lengthiest, and the solution is challenging and deceptive.
-Programmers must use the provided rotate function {\it f2} to adjust the positions of the four components of the {\it Zero} value four time to get to the correct position.
+Programmers must use the provided rotate function {\it f2} to adjust the positions of the four components of the {\it Zero} value four times to get to the correct position.
 Completing the level requires applying the functions {\it f1}, {\it f2}, and {\it f3} in a nested sequence that is 8 layers deep.
 Consequently, programmers need to frequently switch contexts to recall the purpose of each function.
-We speculate that this level demands the mental flexibility typical of a Haskell expert to quickly internalize these functions.
-GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding structual clarity.
+We speculate that this level demands the mental flexibility typical of a Haskell expert to internalize these functions quickly.
+GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding structural clarity.
 
@@
 \ref{fig:qualitative} Middle): experts find the GeckoGraph most intuitive (5.07), followed by the knowledgeable group (4.87), and the familiar group (4.80).
-The beginner group found it to be the least intuitive but still rated a positive score of (4.71).
+The beginner group found it to be the least intuitive but still rated a positive score of 4.71.
 When answering the question ``How helpful do you find GeckoGraph in finding the solution during the game?", the answer is more divided into different experience groups (Fig.
@@
 Two interesting trends emerge.
-First, the more experienced the users are, the more likely they find GeckoGraph intuitive and easy to use.
-We believe this aligns with our design decision to make GeckoGraph resemble text-based type annotation.
+First, the more experienced the users are, the more likely they are to find GeckoGraph intuitive and easy to use.
+This aligns with our design decision to make GeckoGraph resemble text-based type annotation.
 Those familiar with text-based type annotation find GeckoGraph second nature to read.
@@
 We speculate this pattern occurs because users who are experts in text-based type annotations naturally prefer reading text-based notation and tend to ignore GeckoGraph, diminishing its usefulness.
-Conversely, novice users are more likely to be drawn to GeckoGraph and find its visualization offers more clarity and useful insights.
+Conversely, novice users are more likely to be drawn to GeckoGraph and find that its visualization offers more clarity and useful insights.
+
+% To quote some feedback from participants: ``Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature.
+" ``It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is."
+
+% To quote a participant, ``GeckoGraph is quite intuitive to see the permutations of the arguments. Also, to see how to produce and consume arguments."
 
@@
 They are designed to be different from well-known Haskell functions to minimize the familiarity variable.
-They are also designed with an interest of being puzzling and fun.
+They are also designed with an interest in being challenging and fun.
 These functions may not be the most representative of real-world Haskell programming.
@@
 \subsubsection{Identify the Most Important Features}
-One trend that we saw from the qualitative feedback is that programmers find GeckoGraph helpful for finding patterns and important features of the types.
+One trend we saw from the qualitative feedback is that programmers find GeckoGraph helpful for finding patterns and important features of the types.
 Programmers are very positive about GeckoGraph's ability to reveal the most helpful features of a type in distinctive visual elements such as color, length, and height.
+To quote some feedback from participants: ``{\it Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature.} " ``{\it It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is.}"
 
-\textbf{The colors of GeckoGraph} help programmers to see the permutation of type variables in the input and output of a function.
-A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
-One example of GeckoGraph's effective use of color is the ``rotation" function in the user study (Fig.
-\ref{fig:rotate}).
-With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions.
-For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game is less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.
-To quote a participant, ``GeckoGraph is quite intuitive to see the permutations of the arguments.
-Also, to see how to produce and consume arguments."
+\subsubsection{Generalize patterns in type classes}
+A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generalized types and those that work with concrete types \cite{Tirronen2015-nr}.
+For instance, the type constraints {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, as well as {\tt Foldable} instances versus {\tt []} can be particularly challenging.
+Many Haskell learners may find the differences between concrete types and their generic counterparts irrelevant to their programming tasks, thereby complicating their quest for solutions.
+In the example shown in Fig.
+\ref{fig:example-foldable}, a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String} literal.
+The compiler responds with an error message that compares the provided \texttt{Char} type to a confusingly named type \texttt{t0 a0}.
+In this case, the information has been generalized too much and the programmer can hardly find their bearing to the original problem.
+This level of generalization can undermine a programmer's confidence in understanding the language and navigating type errors.
+In GeckoGraph, both list types and types with a Foldable instance share a similar structure.
+Therefore, the type \texttt{t0 a0} in the error message matches the shape of \texttt{[a]}.
+Bring the programmer closer to the real issue, trying to match a complex type with a concrete simple type.
+This feature lowers the entry barrier for languages like Haskell, where even basic arithmetic functions {\tt +} and {\tt -} are defined on generic types.
+This approach lowers the barrier to introducing foundational language concepts without sacrificing generality.
 
-\begin{figure}[]
-  \includegraphics[width=0.6\linewidth]{figures/rotate}
-  \caption{\label{fig:rotate} The `rotate' function in level 8 of the user study. The name given in the game is `f2'. It shuffles the type arguments of a Zero type}
-\end{figure}
-
-\textbf{The horizontal axis of GeckoGraph} often becomes intuitive when identifying differences in function arities.
-For example, in Fig.
-\ref{fig:add3}, the programmer intended to implement a function that sums 3 integers.
-In the implementation, the programmer missed a \texttt{(+)} function at the end; the resulting function type is largely different in length.
-It is also clear that the function needs to apply to one more binary function to satisfy the length requirement.
-
-\begin{figure}[]
-  \includegraphics[width=0.6\linewidth]{figures/Add3}
-  \caption{\label{fig:add3} An implementation of function \texttt{add3} but the author missed an (+) from the correct implementation (.) ((+) .) (+).
-GeckGraph highlights the difference in arity, and reveals that a binary function is needed on the right-hand side for the arity to match.
-}
-\end{figure}
-
-\textbf{The vertical axis of GeckoGraph} often sheds light on the most complex structure of this type.
-This can often be very useful when inspecting mismatching type errors where data types are nested.
-Common examples include when programmers forget to apply the value to ``return" in a monadic block or to use \texttt{liftIO} to cast an \texttt{IO} effect.
-For example, in Fig.
-\ref{fig:maybe}, the uses of \texttt{return} are excessive.
-It can be easily identified by examining the difference in the vertical layers of the two types.
-In text-based type notation, this is distinguished by different pairs of parenthesis.
-However, parenthesis is an overloaded syntax in type notation.
-In Haskell, parentheses are used to enclose tuples \texttt{(a, b)}, specify the fixity \texttt{ (a -> b) -> c}, or have no effect \texttt{a -> (b -> c)}.
-
-To quote some feedback from participants: ``Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature.
-" ``It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is."
-
-\begin{figure}[]
-  \includegraphics[width=0.6\linewidth]{figures/Maybe}
-  \caption{\label{fig:maybe} The function \texttt{f} is planned to have the type \texttt{Maybe a -> a -> Maybe a}.
-The programmer mistakenly applied the result to the \texttt{return} function, making the result inside a Monad instance.
-GeckoGraph reveals the difference in the ``layers" of types.
-}
-\end{figure}
+\subsubsection{Consistent color scheme}\label{par:color-scheme}
+A common task in programming is to scan for a desired function from a sea of potentially useful functions, such as library documentation.
+During scanning, programmers often have partial knowledge of the desired function, e.g., the arity, one of the argument types, or the type class it must fulfill.
+A typical example is conversion: using a known \texttt{String} type to produce a desired \texttt{Data.Text}  type.
+Another example is the `lookup' function: using a known \texttt{Data.Map a b} to produce a desired \texttt{b} type.
+GeckoGraph supports this task by using consistent colors for the same type identifier.
+Programmers can rely on the color grouping to scan for the desired type in their project or in third-party library documentation.
 
@@
 GeckoGraph simulates the prefix notation in type constructors and the infix notation in type operators.
-With these considerations, we ensured that programmers who were familiar with the Haskell language had no problem understanding and using GeckoGraph fluently with a minimal amount of training.
+With these considerations, we ensured that programmers familiar with the Haskell language had no problem understanding and using GeckoGraph fluently with minimal training.
 
-One important class of feedback from the open section is that many programmers mentioned that they did not have any prior knowledge of Haskell but were able to solve the puzzles with the help of GeckoGraph.
+This is reflected in our user study feedback. Many participants mentioned that they did not have any prior knowledge of Haskell but were able to solve the puzzles with the help of GeckoGraph.
 ``It is similar enough to traditional types that it is intuitive." ``This was how I parse the textual representation of types" was pointed out by multiple participants.
+
+\subsubsection{Language Agnostic}
+GeckoGraph can be implemented in any language that exercises static typing.
+We identified applications of this feature in programming and education.
+Programming projects often use more than one language.
+Situations include projects using foreign function interfaces or multiple languages to accommodate a client-server architecture.
+In these cases, GeckoGraph provides a shared notation to describe the functionality and features of programming constructs despite their different text-based notations.
+In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages.
+For example, when describing the product type in algebraic data types, different language communities use various names: record, named tuple, struct, etc.
+GechoGraph helps to unify the same programming concept and ignore the minute nomenclature differences.
 
@@
 \subsubsection{Color Encoding}
-GeckoGraph highly relies on color hue as a main encoding.
+GeckoGraph relies highly on color as the primary encoding.
 It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify patterns in types, such as the order and placement of substructures.
@@
 Although GeckoGraph uses color-blind friendly schemes, it is only a method to avoid indistinguishable types and is not a strong guarantee of effectiveness.
-For this, we are exploring different visual encoding mehods, such as patterns and shapes, to maximize the accessibility of GeckoGraph.
+For this, we are exploring different visual encoding methods, such as patterns and shapes, to maximize the accessibility of GeckoGraph.
 
@@
 We practiced multiple gamification techniques: levels, story/theme, and goals/rewards.
-\cite{Hamari2014-mc} This not only allowed us to have confidence that participants are motivated to complete the tasks, it also lent us popularity in the Haskell community and led to a historically high participation rate.
+\cite{Hamari2014-mc} This not only allowed us to have confidence that participants are motivated to complete the tasks, but it also lent us popularity in the Haskell community and led to a historically high participation rate.
 Gamification has been shown to improve engagement and motivation.
 This has been harnessed by many research projects to improve participation in human studies \cite{He2014-vp}.
-We identify that studies on functional programming are often technical and intimidating; our use of gamification not only attracted historically high participation, but also attracted a wide distribution of experience levels.
+We identify that studies on functional programming are often technical and intimidating; our use of gamification attracted historically high participation with a wide distribution of experience levels.
 
@@
 We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of.
-The results of the study show that GeckoGraph helps improve programmers' ability to succeed in programming tasks we designed, especially for novice programmers.
+The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.
 

### 2025-04-08 main.tex
Resturcture intro

@@
 In programming languages, a polymorphic type \cite{Cardelli1987-fp} can represent values of different types while providing an interface to describe common behaviors for those values.
-Polymorphic types are central to the succinctness of contemporary statically-typed languages, enabling a considerable degree of type-safe abstraction and reusability.
-Parametric types are available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}.
-Polymorphism allows programs to be written in a way that is more generic and adaptable to different data types, enabling greater flexibility and code reuse.
+Polymorphic types are one of the oldest topics in programming language theory \cite{Cardelli1987-fp}.
+They are central to the succinctness of statically-typed languages, while also enabling a high degree of type-safe abstraction and reusability.
+Polymorphic types are available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}.
+
+% Polymorphism allows programs to be written in a way that is more generic and adaptable to different data types, enabling greater flexibility and code reuse.
 Polymorphic types are ideal for modeling abstractions, such as properties of mathematical objects and laws that hold on these objects.
 
-While promoting robustness and code reusability, polymorphic types present challenges in comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}.
+While promoting robustness and code reusability, polymorphic types present challenges in learning and comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}; as it is often argued that the expressiveness power of type systems often come at the cost of its usability~\cite{Hage2020-hg}.
 Research indicates that humans tend to focus on concrete types and rely on polymorphic type checking only as a last resort.
@@
 \ref{fig:example-foldable}).
+Unfortunately, not many research focuses on the \textit{usability} of polymorphic types.
 
@@
 
-Polymorphic type is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}.
-However, little research focuses on the \textit{usability} of polymorphic types.
-It is often argued that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}.
-We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI techniques.
+ We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI (Human-Computer Interaction) techniques.
 In this paper, we propose GeckoGraph, a graphical notation for types.
@@
 
-\section{Evaluation}
+\section{Evaluation} \label{sec:evaluation}
 To evaluate the usefulness of GeckoGraph, we designed and conducted a controlled experiment in the form of a game called "Zero to Hero".

### 2025-04-09 main.tex
Image float
Game levels

@@
 
-\begin{figure}[]
+\begin{figure}[h]
   \includegraphics[width=\linewidth]{figures/Indicator}
@@
 
-\subsection{An example level}
-We illustrate the task of the user study using level 4 of the game.
-At this level (Fig.
-\ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a b -> Hero b b}.
-The available functions are \texttt{f1::Zero a b -> Hero b a}, \texttt{f2::Zero a a -> Hero a a}, \texttt{f3::Zero a b -> Hero b a}, and \texttt{f4::Zero a b -> Hero b b}.
-Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions.
-The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
-
+\subsection{Example Levels}
 \begin{figure}[h]
-  \includegraphics[width=\linewidth]{figures/GamePlay}
+  \includegraphics[width=\linewidth]{figures/Level2}
   \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero.
-On this level -- level 4 (Shown at F) -- the players need to implement the function \texttt{ zeroToHero :: Zero a b -> Hero b b} (B).
+On this level -- level 2 (Shown at A) -- the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero a} (B).
 They write their own definitions in the code editor (D) using a set of provided functions (G).
 The inferred type of their current definition is shown in (C).
-When ready, they can test their solution by clicking on the \textit{Attempt} button (A).
-They can also skip a level by clicking on the \textit{Bypass} button next to it.
-The output from the compiler, if there is any, is shown in a window below (E).
-The GeckoGraph in the screenshot uses a different color scheme that is optimized for computer screens.}
+When ready, they can test their solution by clicking on the \textit{Attempt} button (E).
+They can also skip a level by clicking on the \textit{Skip (n left)} button next to it (F).}
+
 \end{figure}
 
-At each level of the game, the programmer must select the right set of functions and compose them correctly to achieve the target result.
-Not all provided functions are useful.
-In particular, for this level, participants must discover that only \texttt{f4} and \texttt{f2} are necessary to produce the desired results.
-One implementation that satisfies the target type is \texttt{zeroTohero z = f2 (f4  z)}.
+We illustrate the task of the user study using level 2 of the game.
+At this level (Fig.
+\ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a -> Hero b}.
+The available functions are \texttt{runZero::Zero a -> a}, \texttt{makeHero:: a -> Hero a}.
+A generic function \texttt{(\$)} is provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions.
+The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
+To complete this level, player only needs to use each of the provided function once. The implementation {\tt zeroToHero z = makeHero (runZero z)} is one possible solution.
+
+\begin{figure}[!h]
+  \includegraphics[width=\linewidth]{figures/Level3}
+  \caption{\label{fig:level3} A screenshot showing the level 3 of the game ZeroToHero.
+In this level, the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero (a a)}.
+Players are helped by the functions: {\tt f1 :: Zero a -> Hero a}, {\tt f2 :: Zero a -> (a, a)}, and {\tt f3 :: Hero a -> Hero (a, a)}.}
+\end{figure}
+
+In level 3, the player need to implement the function {\tt zeroToHero :: Zero a -> Hero (a, a)}.
+In this level, not all provided functions are used in the final result.
+In particular,  players must discover that only \texttt{f1} and \texttt{f3} are necessary to produce the desired results.
+One implementation that satisfies the target type is \texttt{zeroTohero z = f3 (f1 z)}.
+Another possible implementation is {\tt zeroToHero z = f3 .
+f1 \$ z}.
 
@@
 Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience.
-Level 3 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple.
+Level 5 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple.
 Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class.

### 2025-04-09 main.tex
More changes

@@
 A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
+Goanna uses one of the qualitative color scheme from ColorBrewer \cite{Harrower2003-vu}.
 One example of GeckoGraph's use of color is the ``rotation" function  (Fig.
@@
 With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions.
-For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game is less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.
+For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game.
+It shuffles the type arguments one step to the left.
+The purpose of this function becomes less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.
 
@@
   \includegraphics[width=0.6\linewidth]{figures/rotate}
-  \caption{\label{fig:rotate} The `rotate' function in level 8 of the user study. The name given in the game is `f2'. It shuffles the type arguments of a Zero type}
+  \caption{\label{fig:rotate} The `rotate' function in the user study. The name given in the game is `f2'. It shuffles the type arguments one step to the left.}
 \end{figure}
@@
 
-We illustrate the task of the user study using level 2 of the game.
-At this level (Fig.
-\ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a -> Hero b}.
+We illustrate the task of the user study using levels 2 and 3of the game.
+At level 2 (Fig.
+\ref{fig:level-example}), the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero b}.
 The available functions are \texttt{runZero::Zero a -> a}, \texttt{makeHero:: a -> Hero a}.
@@
 The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
-To complete this level, player only needs to use each of the provided function once. The implementation {\tt zeroToHero z = makeHero (runZero z)} is one possible solution.
+To complete this level, players only need to use each of the provided function once.
+The implementation {\tt zeroToHero z = makeHero (runZero z)} is one possible solution.
 
@@
 
-In level 3, the player need to implement the function {\tt zeroToHero :: Zero a -> Hero (a, a)}.
+In level 3 (Fig.
+\ref{fig:level3}),  players need to implement the function {\tt zeroToHero :: Zero a -> Hero (a, a)}.
 In this level, not all provided functions are used in the final result.
@@
 Another possible implementation is {\tt zeroToHero z = f3 .
-f1 \$ z}.
+f1 \$ z}, exercising the Haskell function composition idioms.
 

### 2025-04-09 main.tex
Updates from Overleaf
typo

@@
 
-We illustrate the task of the user study using levels 2 and 3of the game.
+We illustrate the task of the user study using levels 2 and 3 of the game.
 At level 2 (Fig.
@@
 We provide all the level settings we used in our user study.
-The online game is still open source and available for evaluation \cite{Anonymous_undated-ne}.
+The online game is open source and available for evaluation \cite{Anonymous_undated-ne}.
 However, this can be attempted locally with a Haskell interpreter or even with a pen and paper.

### 2025-05-02 main.tex
overleaf

@@
 
-While promoting robustness and code reusability, polymorphic types present challenges in learning and comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}; as it is often argued that the expressiveness power of type systems often come at the cost of its usability~\cite{Hage2020-hg}.
+While promoting robustness and code reusability, polymorphic types present challenges in learning and comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}; as it is often argued that the expressiveness power of type systems often comes at the cost of their usability~\cite{Hage2020-hg}.
 Research indicates that humans tend to focus on concrete types and rely on polymorphic type checking only as a last resort.
@@
 Our study evaluating 714 participants' ability to solve a series of programming challenges using GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
-The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the programming challenges, especially for novice programmers.
+The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving programming challenges, especially for novice programmers.
 
@@
 However, GeckoGraph could also be used in imperative and multiparadigm languages such as TypeScript and Rust.
-We provide a GeckoGraph visualization tool avaiable online for multiple languages \cite{Fu2025-ka}.
+We provide a GeckoGraph visualization tool available online for multiple languages \cite{Fu2025-ka}.
 
@@
 A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
-Goanna uses one of the qualitative color scheme from ColorBrewer \cite{Harrower2003-vu}.
+Goanna a qualitative color scheme from ColorBrewer \cite{Harrower2003-vu}.
 One example of GeckoGraph's use of color is the ``rotation" function  (Fig.
@@
   \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph.} Overall, programmers consider GeckoGraph to be intuitive and not distracting.
-However, opinions are split on its helpfulness.
+However, beginners are more likely than experienced functional programmers to find it helpful.
 }
@@
 \subsection{Strengths}
-From the results of our experiment, we see that using GeckoGraph has a significant effect on the success rate of our participants, especially on less experienced programmers.
+From the results of our experiment, we see that using GeckoGraph has a significant effect on the success rate of our participants, especially for less experienced programmers.
 We also see that with the data we collected, we did not find a significant time difference between programming with and without GeckoGraph.
@@
 In Jung's design, a function \texttt{a -> b} looks sufficiently distinct from a data type \texttt{f a b}.
-This is important because, in functional languages, it is very common for abstraction to be drawn from function and normal data types.
+This is important because, in functional languages, it is very common for abstraction to be drawn from functions and normal data types.
 For example,  functions and lists both have a functor instance, and their inner values can be altered using a \texttt{fmap} function.
@@
 However, GeckoGraph focuses on type-level information, which no other research projects do.
-In addition, GeckoGraph is evaluated in a much larger study than the other projects, and GeckoGraph is evaluated with a wider range of experience levels.
+In addition, GeckoGraph is evaluated in a much larger study than the other projects, and it is also evaluated with a wider range of experience levels.
 
@@
 Although our experiment scenarios are drawn from real-world programming tasks, a certain level of variable control is still applied to remove the effect of familiarity with the tools and libraries.
-However, it is necessary to assess the usefulness of tools such as GeckoGraph by their real-life usage.
-To study this, different research methods should be used to study realistic usage and human experience.
+However, it is necessary to assess the usefulness of tools such as GeckoGraph in terms of their real-life usage.
+Different research methods should be used to study the effects of GeckoGraph on realistic programming tasks.
 This may include field deployments or case studies.
@@
 
-\subsection{Level 2: Assemble required}
+\subsection{Level 2: Assembly required}
 
@@
 \begin{itemize}
-    \item \texttt{f1 :: Zero a b c -> Zero c a b}
-    \item \texttt{f2 :: Zero a b c -> Hero a b c}
-    \item \texttt{fmap :: (c -> d) -> Zero a b c -> Zero a b d}
+    \item \texttt{f1 :: Zero a b c d -> Zero a a b b}
+    \item \texttt{f2 :: Zero a b c d -> Hero c c d d}
+    \item \texttt{f3 :: Zero a b c d -> Zero d c b a}
     \item \texttt{(\$) :: (a -> b) -> a -> b}

### 2025-05-02 main.tex
More from overleaf

@@
   \includegraphics[width=\linewidth]{figures/SuccessfulRate}
-  \caption{\label{fig:success-rate}{\bf Success rate of each task with and without GeckoGraph, grouped by experience level.} The figure is cropped from 70\% to 100\% for readability.
+  \caption{\label{fig:success-rate}{\bf Success rate of each task, with 95\% confidence interval, grouped by experience level.} The figure is cropped from 50\% to 100\% for readability.
 In most tasks, GeckoGraph provides a small edge.
@@
   \includegraphics[width=\linewidth]{figures/Qualitative}
-  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph.} Overall, programmers consider GeckoGraph to be intuitive and not distracting.
+  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph, with 95\% confidence interval.} Overall, programmers consider GeckoGraph to be intuitive and not distracting.
 However, beginners are more likely than experienced functional programmers to find it helpful.

### 2025-10-02 main.tex
updates

@@
 
- We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI (Human-Computer Interaction) techniques.
+ We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern Human-Computer Interaction (HCI) techniques.
 In this paper, we propose GeckoGraph, a graphical notation for types.
@@
 GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
-Our study evaluating 714 participants' ability to solve a series of programming challenges using GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
+We designed an online puzzle game \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming challenges using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
 The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving programming challenges, especially for novice programmers.
@@
   \caption{
-  \label{fig:advanced}
   Advanced features of GeckoGraph.
@@
 During the study, the time spent by participants on each task is recorded.
-We also record the resulting status of each level, whether it is a success or failure.
+We also record the resulting status of each level, whether it is a success or a failure.
 Before each run, participants nominate their level of Haskell experience on a four-level scale: beginner, familiar,  knowledgeable, and expert.
@@
 Data collection from the study was stopped after the planned cut-off period of 14 days.
-After the cut-off date, the ZeroToHero game is open source and available for free evaluation \cite{Anonymous_undated-ne} and repeating our experiment, but no further data was collected.
+After the cut-off date, the ZeroToHero game is open source and available for free evaluation \footnote{https://zerotohero.fly.dev} and repeating our experiment, but no further data was collected.
 
@@
 
-\begin{figure}[]
+\begin{figure}[ht]
   \includegraphics[width=\linewidth]{figures/SuccessfulRate}
   \caption{\label{fig:success-rate}{\bf Success rate of each task, with 95\% confidence interval, grouped by experience level.} The figure is cropped from 50\% to 100\% for readability.
+Bootstrapping was used for computing the confidence interval because of the non-normal distribution.
 In most tasks, GeckoGraph provides a small edge.
@@
 \subsubsection{Identify the Most Important Features}
-One trend we saw from the qualitative feedback is that programmers find GeckoGraph helpful for finding patterns and important features of the types.
+One trend we saw from the qualitative feedback is that programmers find GeckoGraph helpful for spotting patterns and important features from the types.
 Programmers are very positive about GeckoGraph's ability to reveal the most helpful features of a type in distinctive visual elements such as color, length, and height.
@@
 We provide all the level settings we used in our user study.
-The online game is open source and available for evaluation \cite{Anonymous_undated-ne}.
+The online game is open source and available for evaluation \footnote{https://zerotohero.fly.dev}.
 However, this can be attempted locally with a Haskell interpreter or even with a pen and paper.

### 2025-10-20 main.tex
Updates

@@
 To our knowledge, this is the largest controlled user study on functional programming ever conducted.
-The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.
+Overall, no significant difference were found  in time and success rate between the programmers use GeckoGraph and the ones who do not.
+However, we find that in the most difficult task of the user study, GeckoGraph improve the success rate for novice users (14.5\%) and less experienced users (11.2\%), suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
 \end{abstract}
@@
   \caption{\label{fig:example-foldable} An example type error where a programmer mistakenly provided a \texttt{Char} instead of a \texttt{String} literal.
-\textbf{Left --} The compiler shows an error message comparing the provided Char type to a confusingly named type \texttt{t0 a0}.
-\textbf{Right --} GeckoGraph shows the exact message with the two types in graphic notation, highlighting the structural difference rather than identifier names.}
+{\bf Left:} Original type error.
+{\bf Right:} Type error with GeckoGraph}
 \end{figure}
@@
 GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
-We designed an online puzzle game \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming challenges using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
-The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving programming challenges, especially for novice programmers.
+We designed a gamified online study  \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming puzzles using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
+Overall, no significant difference were found  in time and success rate between the programmers use GeckoGraph and the ones who do not.
+However, we find that in the hardest task of the user study, GeckoGraph improve the success rate for novice users (14.5\%) and less experienced users (11.2\%), suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
+
+\section{Type Systems in Functional Programming}
+
+Static type systems are widely used in large and complex software projects because they help detect errors early and make refactoring safer.
+In functional programming languages, type systems are often more expressive and provide stronger safety guarantees.
+They enable programmers to model program behavior rigorously and enforce correctness at compile time.
+Common characteristics of type systems in functional programming include structural typing, algebraic data types, parametric polymorphism, and type inference.
+
+\subsection{Structural Type Systems}
+
+Unlike the nominal type systems found predominately in languages such as Java and C \cite{AbdelGawad2013-hh}, many functional languages heavily rely on structural typing.
+In nominal systems, two types are considered distinct even if they have identical structure but different names.
+In contrast, structural type systems consider types equivalent as long as their structures match, regardless of their names.
+This approach promotes flexibility, as types can be defined inline without introducing new names, and values are automatically valid for a type when they conform to the structure.
+However, in functional languages such as Haskell, and OCaml, the lack of nominality can also make complex types harder to read and reason about.
+For instance, when encounter type errors, programmers can face two unmatching types with complex structure, understanding where exactly the root error is becomes a tediotasksks of string comparison.
+
+\subsection{Algebraic Data Types}
+
+Algebraic data types (ADTs) \cite{Milner1978-qz} allow developers to construct complex types from simpler ones using sum and product operations.
+A sum types, often called tagged union, $A + B$ represent a type that can be a value of A or a value of B.
+Conversely, a product type $A \times B$ represent a tuple of A and B.
+This precise and composable type definitions for common data structures such as lists, tuples, and result types.
+For example, in Haskell, types like \texttt{Either a b} and \texttt{Maybe a} provide basic building blocks for error handling, and the type \texttt{data List a = [] | a : List a } provides a succinct recursive definitions for lists.
+This illustrates both the expressive power and the potential complexity of ADTs.
+In practice, simple type definitions can turn into an \textit{essay of types} \cite{Ferdowsi2023-cv} as program grows, causing difficulties to interpret and maintain.
+
+\subsection{Parametric Polymorphism}
+
+Parametric polymorphism (often called generics) allows functions and data structures to operate uniformly across different types.
+For instance in Haskell, the\texttt{head :: [a] -> a}, taking a list as input and return the first element, is a polymorphic functions that works on lists of any type, integers, strings, any type is valid.
+It provides a strong and principled form of abstraction, enabling code reuse without sacrificing type safety.
+
+However, Parametric polymorphism also increases cognitive load: programmers must mentally track type variables and reason about how concrete types are instantiated during function application.
+Overuse can make code difficult to follow, especially for readers unfamiliar with the underlying type relationships.
+
+\subsection{Type Inference}
+
+Type inference, or implicit typing, reduces the need for explicit type annotations by allowing the compiler to deduce types automatically.
+Most modern languages employ some level of inference.
+For instance, Java’s var keyword (introduced in version 10) infers variable types at compile time based on their assigned values \cite{Java_Developers2023-an}, saving programmers from writing redundenies like \textit {Car car = new Car()}.
+
+Functional languages such as Haskell and ML go further, using global type inference \cite{Damas1982-zw} to check entire programs without requiring explicit annotations.
+While this improves conciseness, it can reduce code readability and produce complex error messages.
+Many developers therefore choose to write explicit type signatures for clarity, even when inference is available.
+
+To summarize, the type systems of functional programming languages provide rigorous tools for specifying and enforcing program behavior.
+By encoding system constraints directly in types, they prevent many classes of runtime errors.
+However, this rigor introduces trade-offs: a steeper learning curve and less flexibility for exploratory or prototype development.
+In the following sections, we demonstrate how GeckoGraph supports these advanced type-system features by visualizing type level information that is often implicit or difficult for humans to interpret.
 
@@
 In this paper, we use Haskell as an example.
-As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, and many advanced type-level features.
-However, GeckoGraph could also be used in imperative and multiparadigm languages such as TypeScript and Rust.
+As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, type classes, and a few advanced type-level features.
+However, GeckoGraph can also be used in imperative and multiparadigm languages such as TypeScript and Rust.
 We provide a GeckoGraph visualization tool available online for multiple languages \cite{Fu2025-ka}.
@@
 For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity of a function, containing a known type, or type class requirements).
+
+Motivated by the design goals, we designed GeckoGraph with the following feature features and construction rules (Fig. \ref{fig:design}):
 
@@
 It appears as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
-Regular high-order functions (e.g., \texttt{(a -> b) -> c} ) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
+Regular high-order functions (e.g., \texttt{(a -> b) -> c}) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
 The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and high-order functions (design goal \ref{goal3}).
 It is easy to tell high-order functions from the vertical position of its function indicator.
-Similarly, it is easy to count the arity of a function by counting the number of horizontally connected function indicators (Fig.
-\ref{fig:indicator}).
+Similarly, it is easy to count the arity of the outer function by counting the number of function indicators at the highest layer.
+For example, in Fig.
+\ref{fig:indicator}, the higest layer function indicator (the ones colored in the lightest shade of gray) suggest a ternary function.
+The two arguments types and the return type are all functions.
 
@@
 When reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable.
-GeckoGraph uses approximity to help minimize the effort to associate each type variable with all its type classes (design goal \ref{goal2}).
+In GeckoGraph, type classes are explicitly displayed under the type variable cell,  minimizing programmers' effort to associate each type variable with all its type classes (design goal \ref{goal2}).
 
@@
 
-\begin{figure}[]
-  \includegraphics[width=\linewidth]{figures/PreviousVersions}
-  \caption{\label{fig:previous}Previous Versions of GeckoGraph. Different encodings represent named types, type variables, type constructors, and high-order functions.}
-\end{figure}
+\paragraph{Visual Dimensions of GeckoGraph}
 
-\paragraph{Visual Dimensions of GeckoGraph}
 \textbf{The colors of GeckoGraph} help programmers to see the permutation of type variables in the input and output of a function.
 A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
-Goanna a qualitative color scheme from ColorBrewer \cite{Harrower2003-vu}.
+GeckoGraph uses qualitative color scheme from ColorBrewer \cite{Harrower2003-vu} to encode symbolic values of types.
 One example of GeckoGraph's use of color is the ``rotation" function  (Fig.
@@
 With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions.
-For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game.
-It shuffles the type arguments one step to the left.
+For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} appeared in the evaluation (Section \ref{sec:evaluation}), shuffles the type arguments one step to the left.
 The purpose of this function becomes less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.
@@
   \includegraphics[width=0.6\linewidth]{figures/rotate}
-  \caption{\label{fig:rotate} The `rotate' function in the user study. The name given in the game is `f2'. It shuffles the type arguments one step to the left.}
+  \caption{\label{fig:rotate} The `f2' function in the user study. It shuffles the type arguments one step to the left and the leftmost argument wraparound to be the rightmost.}
 \end{figure}
@@
 
-\begin{figure}[]
+\begin{figure}
   \includegraphics[width=0.6\linewidth]{figures/Maybe}
@@
 The design of GeckoGraph enables the visualization of many advanced type-level features.
-\textbf{Kind visualization}: Kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is type constructor (e.g.
+\textbf{Kind visualization}: The kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is a type constructor (e.g.
 [], Maybe) or a type constant (e.g.
 Bool, Char), because of this, kind is often described as the type of types.
-GeckoGraph infers the {\it kind} of type variables and consistently displays this information in GeckoGraph.
-For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} is inferred to have the kind \texttt{* -> *} because of its use on the right-hand side.
+GeckoGraph infers the {\it kind} of type variables and consistently displays this information.
+For example, in Fig.
+\ref{fig:advanced} (A), the variable \texttt{a} is inferred to have the kind \texttt{* -> *} because of its use on the right-hand side.
 GeckoGraph respects this {\it kind} information and displays the same kind level on the first occurrence of \texttt{a}, but over an empty structure, indicated using a dotted outline.
@@
 Because of its graphical layout, GeckoGraph allows programmers to precisely select which part of a type signature they intend to query, that is, in Fig.
-\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a sepecific query that is hard to hard to execute with point and click over text based type annotation.
+\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a sepecific query that is hard to execute with point and click over text based type annotation.
 
-\begin{figure}[h]
-\label{fig:advanced}
+\begin{figure}
   \includegraphics[width=\linewidth]{figures/Advanced}
-  \caption{
+  \caption{\label{fig:advanced}
   Advanced features of GeckoGraph.
@@
 \section{Evaluation} \label{sec:evaluation}
-To evaluate the usefulness of GeckoGraph, we designed and conducted a controlled experiment in the form of a game called "Zero to Hero".
-The game contains 10 levels of varying difficulty.
+To evaluate the usefulness of GeckoGraph, we designed and conducted a gamified controlled experiment.
+The study, "Zero to Hero", contains 10 tasks (levels) of varying difficulty.
 At each level, participants are asked to implement a function named "\texttt{zeroToHero}" with the help of a limited list of provided functions.
 These functions are different at each level, and the target types of Zero and Hero vary at each level as well.
-The details of each level are provided in the Appendix (Appendix \ref{levels}).
+The details of each level are provided in the Appendix \ref{levels}.
 
 The experiment aims to study how programmers use and reason about polymorphic type when performing programming tasks.
-In particular, we studied how programmers scan and select potentially useful functions from a library and compare intended types and actual types during type errors.
+In particular, we studied how programmers scan and select potentially useful functions from a library and how they compare intended types and actual types when type errors are encountered.
 
@@
   \includegraphics[width=\linewidth]{figures/Level2}
-  \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero.
-On this level -- level 2 (Shown at A) -- the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero a} (B).
+  \caption{\label{fig:level-example} A screenshot from the user study.
+For this task -- level 2 (Shown at A) -- the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero a} (B).
 They write their own definitions in the code editor (D) using a set of provided functions (G).
@@
 
-We illustrate the task of the user study using levels 2 and 3 of the game.
-At level 2 (Fig.
+We illustrate the tasks of the user study using levels 2 and 3 as examples.
+For level 2 (Fig.
 \ref{fig:level-example}), the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero b}.
@@
 A generic function \texttt{(\$)} is provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions.
-The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
+The possible solution and other details of the level can be found in Appendix \ref{levels}.
 To complete this level, players only need to use each of the provided function once.
-The implementation {\tt zeroToHero z = makeHero (runZero z)} is one possible solution.
+One possible solution to this task would be {\tt zeroToHero z = makeHero (runZero z)}.
+The user illustrated in Fig.
+\ref{fig:level-example} did not reach a correct solution, as the type signature and GeckoGraphs in {\it To Implement} and {\it Your Result} do not match.
 
@@
   \includegraphics[width=\linewidth]{figures/Level3}
-  \caption{\label{fig:level3} A screenshot showing the level 3 of the game ZeroToHero.
+  \caption{\label{fig:level3} A screenshot showing the level 3 of the user study.
 In this level, the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero (a a)}.
@@
 f1 \$ z}, exercising the Haskell function composition idioms.
+Again, the user illustrated in Fig.
+\ref{fig:level3} did not reach a correct solution, as the type signature and GeckoGraphs in {\it To Implement} and {\it Your Result} do not match.
 
@@
 \subsection{Task Design} \label{subsection:task}
-During the study, participants receive 10 tasks, represented by the 10 levels of the game {\it zeroToHero}.
+During the study, participants receive 10 tasks, represented by the 10 levels of the user study.
 The tasks start off easy but gradually increase in difficulty.
@@
 
-Participants can skip a level during the game if they are stuck.
+Participants can skip a level if they are stuck.
 We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their feedback of the tool.
-We limit the number of skips that a participant can use during the game to four so that submitting qualitative feedback without completing at least some levels is impossible.
+The number of skips a participant can use is limited to four so that it is not possible to submit qualitative feedback without completing at least 6 tasks out of 10.
 
@@
 We also record the resulting status of each level, whether it is a success or a failure.
-Before each run, participants nominate their level of Haskell experience on a four-level scale: beginner, familiar,  knowledgeable, and expert.
+Before starting the study, participants nominate their level of Haskell experience on a four-level scale: beginner, familiar,  knowledgeable, and expert.
 If a participant has completed all 10 levels (with the help of skipping), we invite the participant to complete a post-study survey.
-In it, we ask for their opinion on how intuitive the GeckoGraph design is, how distracting they find GeckoGraph, and how helpful GeckoGraph is during the game, using a seven-point scale.
+In it, we ask for their opinion on how intuitive the GeckoGraph design is, how distracting they find GeckoGraph, and how helpful GeckoGraph is, using a seven-point scale.
 In the end, we ask a few open-ended questions, inviting participants to provide their experience using GeckoGraph and their expectations about the potential applications of GeckoGraph.
@@
 Data collection from the study was stopped after the planned cut-off period of 14 days.
-After the cut-off date, the ZeroToHero game is open source and available for free evaluation \footnote{https://zerotohero.fly.dev} and repeating our experiment, but no further data was collected.
+After the cut-off date, the ZeroToHero game is open source and available for free evaluation\footnote{https://zerotohero.fly.dev} and repeating our experiment, but no further data was collected.
 
@@
 \ref{fig:level-time}.
-Level 7 (mean = 334 seconds) is the hardest task in the game in terms of time, followed by level 8 (mean = 228 seconds) and level 5 (mean = 224 seconds).
+Level 7 (mean = 334 seconds) is the hardest task in terms of time, followed by level 8 (mean = 228 seconds) and level 5 (mean = 224 seconds).
 To complete an average level, the beginner group uses an average of 100 seconds, the familiar group uses 90 seconds, the knowledgeable group uses 80 seconds, and the expert group uses 70 seconds.
@@
   \caption{\label{fig:level-time} Time spent on each level, with 95\% confidence interval.
-We show that the difficulty steadily increases across the game, but levels 5, 7, and 8 are significantly harder than the authors intended.
+We show that the difficulty steadily increases across the user study, but levels 5, 7, and 8 are significantly harder than the authors intended.
 The overall task time of each group roughly matches experience level.}
@@
 We saw that, overall, GeckoGraph provides a higher success rate (96.88\%) than text-based type notation (94.62\%).
-This trend can be seen in every experienced group: beginner group (95.
-12\% vs.
-92.
-68\%), familiar group (97.
-39\% vs.
-93.
-34\%), knowledgeable group (96.
-82\% vs.
-96.
-06\%), and expert group (98.
-2\% vs.
-96.
-40\%).
+This trend can be seen in every experienced group: beginner group (95.12\% vs.
+92.68\%), familiar group (97.39\% vs.
+93.34\%), knowledgeable group (96.82\% vs.96.06\%), and expert group (98.2\% vs.96.40\%).
 We saw the significance decrease as the user's experience increased.
@@
 When breaking down the result in each task (Fig.
-\ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group and task 10 of the familiar group \ref{fig:success-rate}.
+\ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group and task 10 of the familiar group (Fig.
+\ref{fig:success-rate}).
 We will address this correlation in Section~\ref{sec:discussion}.
@@
 The beginner group found it to be the least intuitive but still rated a positive score of 4.71.
-When answering the question ``How helpful do you find GeckoGraph in finding the solution during the game?", the answer is more divided into different experience groups (Fig.
+When answering the question ``How helpful do you find GeckoGraph in finding the solution?", the answer is more divided into different experience groups (Fig.
 \ref{fig:qualitative} right).
@@
   \includegraphics[width=\linewidth]{figures/Qualitative}
-  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph, with 95\% confidence interval.} Overall, programmers consider GeckoGraph to be intuitive and not distracting.
+  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph, with 95\% confidence interval.} Overall, programmers are neutural to GeckoGraph's design.
+Some find GeckoGraph to be intuitive and not distracting.
 However, beginners are more likely than experienced functional programmers to find it helpful.
@@
 
+% We start this section with a strength and weakness analysis of GeckoGraph.
+The disuccion is based on our results of user study.
+Then, In the end, we speculate the applications of GeckoGraph.
+The envisioned applications are grounded in the strengths and weaknesses of GeckoGraph.
 \section{Discussion} \label{sec:discussion}
-\subsection{Strengths}
-From the results of our experiment, we see that using GeckoGraph has a significant effect on the success rate of our participants, especially for less experienced programmers.
-We also see that with the data we collected, we did not find a significant time difference between programming with and without GeckoGraph.
-To extrapolate the observed expressiveness, we speculate on the practical benefits of programming with GeckoGraph.
 
-\subsubsection{Identify the Most Important Features}
-One trend we saw from the qualitative feedback is that programmers find GeckoGraph helpful for spotting patterns and important features from the types.
-Programmers are very positive about GeckoGraph's ability to reveal the most helpful features of a type in distinctive visual elements such as color, length, and height.
-To quote some feedback from participants: ``{\it Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature.} " ``{\it It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is.}"
+This section discusses the main findings and implications of our study. 
+We begin by outlining the strengths of GeckoGraph observed from both quantitative and qualitative results (Section~\ref{sec:discussion:strengths}). 
+We then discuss the limitations of the current design (Section~\ref{sec:discussion:weaknesses}) and reflect on the gamified nature of our user study (Section~\ref{sec:discussion:gamification}). 
+Finally, we identify potential applications of GeckoGraph in programming tools, documentation, and education (Section~\ref{sec:discussion:applications}).
 
-\subsubsection{Generalize patterns in type classes}
-A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generalized types and those that work with concrete types \cite{Tirronen2015-nr}.
-For instance, the type constraints {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, as well as {\tt Foldable} instances versus {\tt []} can be particularly challenging.
-Many Haskell learners may find the differences between concrete types and their generic counterparts irrelevant to their programming tasks, thereby complicating their quest for solutions.
-In the example shown in Fig.
-\ref{fig:example-foldable}, a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String} literal.
-The compiler responds with an error message that compares the provided \texttt{Char} type to a confusingly named type \texttt{t0 a0}.
-In this case, the information has been generalized too much and the programmer can hardly find their bearing to the original problem.
-This level of generalization can undermine a programmer's confidence in understanding the language and navigating type errors.
-In GeckoGraph, both list types and types with a Foldable instance share a similar structure.
-Therefore, the type \texttt{t0 a0} in the error message matches the shape of \texttt{[a]}.
-Bring the programmer closer to the real issue, trying to match a complex type with a concrete simple type.
-This feature lowers the entry barrier for languages like Haskell, where even basic arithmetic functions {\tt +} and {\tt -} are defined on generic types.
-This approach lowers the barrier to introducing foundational language concepts without sacrificing generality.
+\subsection{Strengths}\label{sec:discussion:strengths}
+Overall, our experimental results show statistically significant difference in task completion time between conditions with and without GeckoGraph.
+However, we find using GeckoGraph significantly improves task success rates in the hardest task (task 10) across the two less experienced user groups by 14.5\% and 11.2\% respectively.
+To interpret the observed advantages, we discuss the practical benefits that GeckoGraph offers to programmers.
 
-\subsubsection{Consistent color scheme}\label{par:color-scheme}
-A common task in programming is to scan for a desired function from a sea of potentially useful functions, such as library documentation.
-During scanning, programmers often have partial knowledge of the desired function, e.g., the arity, one of the argument types, or the type class it must fulfill.
-A typical example is conversion: using a known \texttt{String} type to produce a desired \texttt{Data.Text}  type.
-Another example is the `lookup' function: using a known \texttt{Data.Map a b} to produce a desired \texttt{b} type.
-GeckoGraph supports this task by using consistent colors for the same type identifier.
-Programmers can rely on the color grouping to scan for the desired type in their project or in third-party library documentation.
+\subsubsection{Identifying the Most Important Features}
+Qualitative feedback revealed that programmers found GeckoGraph particularly helpful for spotting patterns and identifying important features in type signatures. 
+Participants appreciated how GeckoGraph highlights key aspects of a type using distinctive visual properties such as color, length, and height. 
+Representative comments include: 
+``{\it Types are much easier to understand with GeckoGraph than by trying to parse all parentheses and understand the types from the signature.}'' and 
+``{\it It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is.}''
 
-\subsubsection{Low barrier to learn and understand}
-GeckoGraph has some key similarities to traditional text-based type notation.
-GeckoGraph respects the left-to-right reading order.
-GeckoGraph uses the familiar symbolic name as the secondary encoding.
-GeckoGraph simulates the prefix notation in type constructors and the infix notation in type operators.
-With these considerations, we ensured that programmers familiar with the Haskell language had no problem understanding and using GeckoGraph fluently with minimal training.
+\subsubsection{Generalizing Patterns in Type Classes}
+A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generic types and those that work on concrete types \cite{Tirronen2015-nr}. 
+For example, differentiating between {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, or {\tt Foldable} and {\tt []} can be difficult. 
+Learners often find such distinctions irrelevant to their immediate tasks, which complicates problem-solving and error interpretation.
 
-This is reflected in our user study feedback. Many participants mentioned that they did not have any prior knowledge of Haskell but were able to solve the puzzles with the help of GeckoGraph.
-``It is similar enough to traditional types that it is intuitive." ``This was how I parse the textual representation of types" was pointed out by multiple participants.
+As shown in Fig.~\ref{fig:example-foldable}, when a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String}, the compiler reports a confusing mismatch involving a generalized type \texttt{t0 a0}.
+In GeckoGraph, both list types and types with a \texttt{Foldable} instance share similar visual structures. 
+This correspondence helps programmers relate generalized types back to familiar concrete ones, lowering the cognitive barrier to understanding polymorphism. 
+Such visual consistency is especially valuable in Haskell, where even simple arithmetic operators like {\tt +} and {\tt -} are defined generically.
+
+\subsubsection{Consistent Color Scheme}\label{par:color-scheme}
+A frequent task in programming is scanning through documentation or libraries to find a desired function, often based on partial knowledge such as argument types, arity, or expected output. 
+For example, one might look for a conversion from \texttt{String} to \texttt{Data.Text}, or a function like \texttt{lookup} that operates on \texttt{Data.Map a b} and returns a value of type \texttt{b}.
+GeckoGraph supports this by assigning consistent colors to identical type identifiers, allowing programmers to visually group related types and quickly locate relevant functions.
+
+\subsubsection{Low Barrier to Learning and Understanding}
+GeckoGraph retains key conventions from traditional textual type notation: it follows left-to-right reading order, preserves symbolic names as secondary encodings, and mimics prefix and infix structures.
+These design choices ensure that programmers familiar with Haskell can intuitively understand and use GeckoGraph with minimal training.
+
+This ease of adoption was evident in user feedback. Several participants noted that they had no prior Haskell experience but were able to solve the puzzles successfully with GeckoGraph’s help. 
+Comments included: 
+``{\it It is similar enough to traditional types that it feels intuitive.}'' and 
+``{\it This was how I parse the textual representation of types.}''
 
 \subsubsection{Language Agnostic}
-GeckoGraph can be implemented in any language that exercises static typing.
-We identified applications of this feature in programming and education.
-Programming projects often use more than one language.
-Situations include projects using foreign function interfaces or multiple languages to accommodate a client-server architecture.
-In these cases, GeckoGraph provides a shared notation to describe the functionality and features of programming constructs despite their different text-based notations.
-In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages.
-For example, when describing the product type in algebraic data types, different language communities use various names: record, named tuple, struct, etc.
-GechoGraph helps to unify the same programming concept and ignore the minute nomenclature differences.
+GeckoGraph can be implemented in any statically typed language. 
+This makes it useful in multi-language programming environments—such as projects combining client-server architectures or foreign function interfaces—where it can provide a unified visual notation across different syntaxes.
+In education, GeckoGraph abstracts away naming differences between languages. 
+For example, the product type in algebraic data types is known variously as a \textit{record}, \textit{struct}, or \textit{named tuple}. 
+GeckoGraph helps unify these equivalent concepts visually, promoting language-independent understanding.
 
-\subsection{Weaknesses}
+\subsection{Weaknesses}\label{sec:discussion:weaknesses}
 \subsubsection{Space Usage}\label{subsec:space}
-GeckoGraph uses horizontal space in proportion to the size of the type signature syntax tree, and GeckoGraph uses vertical space in proportion to the depth of the syntax tree.
-Compared to the traditional text-based language, GeckoGraph has the limitation of requiring vertical space.
-We have identified some approaches to minimize space usage while retaining most of the advantages of using GeckoGraph, such as displaying only the color blocks without the secondary encoding of identifier names.
+GeckoGraph’s layout scales horizontally with the size of a type’s syntax tree and vertically with its depth. 
+Compared to traditional text-based notation, this results in higher vertical space consumption. 
+We have identified possible optimizations, such as displaying only color blocks while omitting textual identifiers, to reduce space usage while preserving the main advantages of the design.
 
 \subsubsection{Color Encoding}
-GeckoGraph relies highly on color as the primary encoding.
-It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify patterns in types, such as the order and placement of substructures.
-However, the perception of color is different from person to person.
-This becomes an even bigger issue for visually impaired programmers.
-Although GeckoGraph uses color-blind friendly schemes, it is only a method to avoid indistinguishable types and is not a strong guarantee of effectiveness.
-For this, we are exploring different visual encoding methods, such as patterns and shapes, to maximize the accessibility of GeckoGraph.
+GeckoGraph relies heavily on color as its primary visual channel. 
+While this offers strong perceptual grouping \cite{Zeng2023-jz}, color perception varies among individuals, posing accessibility challenges for visually impaired programmers. 
+Although GeckoGraph adopts color-blind–friendly palettes, this only mitigates some issues. 
+To improve accessibility, future work will explore alternative encodings such as patterns and shapes in addition to color.
 
-\subsection{Gamified Human Study}
+\subsection{Gamified Human Study}\label{sec:discussion:gamification}
+Our user study was intentionally designed as a game-like experience, incorporating multiple gamification techniques such as levels, story elements, and goals/rewards \cite{Hamari2014-mc}. 
+This approach ensured participant motivation and contributed to a historically high participation rate within the Haskell community. 
+Gamification has been widely shown to enhance engagement and motivation in human–computer interaction research \cite{He2014-vp}. 
+In our context, it helped attract a diverse set of participants across a wide range of experience levels, addressing a common challenge in studies of functional programming.
 
-It is important to recognize that the human study is designed to be a series of puzzles.
+\subsection{Potential Applications}\label{sec:discussion:applications}
+
+\subsubsection{Programming Assistance}
+Post-study feedback highlighted strong interest in integrating GeckoGraph into programming environments. 
+Modern IDEs already allow users to inspect expression types via hover tooltips; GeckoGraph could enhance this feature by visualizing type structures directly in the tooltip. 
+Similarly, GeckoGraph could be used to visualize mismatched parts of conflicting types during compilation errors, as illustrated in Fig.~\ref{fig:example-foldable}.
+
+\subsubsection{Documentation Assistance}
+Our findings suggest GeckoGraph would be highly effective for visualizing function types in API documentation and programming libraries. 
+Because GeckoGraph representations can be generated automatically, integration into documentation systems like Hoogle \cite{Mitchell_undated-fb} would require minimal effort. 
+A quick visual scan on the page can quickly review the theme of a library or package, i.e., the domains and codomains most functions operate on.
+For instance, a library mostly support string operation should predominately show the color of {\tt String} type, while a low-level graphic related library may show mainly the colors for {\tt Vector} and {\tt Float}.
+
+\subsubsection{Pedagogical Applications}
+GeckoGraph can serve as a valuable teaching aid for introducing programming language concepts that are abstract or difficult to explain textually. 
+Several study participants with no prior Haskell experience reported that GeckoGraph helped them understand type-related puzzles and complete all tasks successfully.
+
+Beyond basic learning, GeckoGraph’s advanced features (Section~\ref{sec:benefits}) also make it suitable for teaching higher-level functional programming concepts. 
+For instance, consider the \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension (Fig.~\ref{fig:assoc}). 
+Although the type signature is concise, it is cognitively difficult to parse due to multiple variables and their hidden kinds. 
+GeckoGraph reveals these higher-kinded types visually, exposing partially applied data types and making the semantics easier to understand.
+
+\begin{figure}[]
+  \includegraphics[width=\linewidth]{figures/assoc}
+  \caption{\label{fig:assoc} 
+  The \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension. 
+  Even without knowing the exact definitions, viewers can easily see that \texttt{f}, \texttt{g}, and \texttt{h} are all higher-kinded types—information that is obscured in traditional textual notation.}
+\end{figure}
+
+\section{Related work}
+This section reviews prior work in three areas relevant to this research. First, we examine approaches to the visualization of types in programming languages. 
+We than broaden our view to general visualization techniques applied to programming environments, including topics like memory allocation and change history.
+Finally, we reviews studies comparing visual and textual representations in programming education and environments.
+
+\subsection{Visualization of Types}
+A similar half-enclosing structure was proposed in Jung’s visualization of types~\cite{Jung2000-oc}.
+In Jung’s notation, the type constructor partially encloses its arguments, but its figure is positioned at the bottom right (Fig.~\ref{fig:jung}).
+In contrast, GeckoGraph follows a natural left-to-right reading order, allowing larger structures in a type signature to take visual precedence over smaller ones.
+
+Compared to Jung’s approach, GeckoGraph offers two major advantages. 
+First, it naturally represents the general notion of a curried function. Partial application can be intuitively read as removing the first argument, which is not as clear in Jung’s notation. 
+Second, the shape of a function type remains visually consistent with that of normal data types.
+In Jung’s design, a function type such as \texttt{a -> b} looks markedly different from a data type \texttt{f a b}.
+This distinction is important because, in functional languages, abstractions often generalize over both functions and data types. 
+For example, both functions and lists have \texttt{Functor} instances, allowing their inner values to be transformed using \texttt{fmap}. 
+GeckoGraph’s consistent visual shape helps make such generalizations more apparent.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/Jung}
+  \caption{
+        \label{fig:jung}
+        Comparison of the \texttt{map} function (\texttt{(a -> b) -> [a] -> [b]}) in text notation, GeckoGraph, and Jung’s notation.
+  }
+\end{figure}
+
+More recent work, TYPEical~\cite{Moy2020-tr}, provides an interactive visualization of function types for exploring and comparing functions in R packages. 
+This project is innovative in that it uses a Sankey diagram~\cite{Keiderling2012-ag} to illustrate the flow between collections of functions and the data types they operate on, as well as the frequency of their use, providing a high level overview of the usage statistics.
+GeckoGraph shares TYPEical’s vision that visualization should support three key tasks: \textbf{finding functions}, \textbf{determining types}, and \textbf{comparing types}. 
+However, unlike GeckoGraph, TYPEical’s visualization is generated ahead of time and is aimed primarily at language designers and core library developers rather than students or everyday programmers.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/Typical.png}
+  \caption{
+        \label{fig:typical}
+        TYPEical’s type flow visualization. 
+        The figure displays the type signatures of functions in R’s \texttt{base} library (top axis) and their first and second arguments (middle and bottom axes).
+  }
+\end{figure}
+
+\subsection{Visualization in Programming}
+\label{sec:viz-programming}
+
+The use of visualization to improve the comprehension of polymorphic types is not new. 
+It has been applied to represent document structures, runtime behavior, and static analysis results. 
+For example, FluidEdt~\cite{Ou2015-vr} displays heap graphs in the editor margin, while I3~\cite{Beck2015-my} presents search similarity and change history through compact block-based diagrams. 
+Almeida et~al.~\cite{Almeida2022-bv} introduced a visualization technique to aid understanding of ownership and borrowing in Rust. 
+Although graphical representations of types remain a relatively narrow field, it has been explored in research such as Clerici et~al.~\cite{Clerici2013-ru}, who proposed a graph-based type inference system that visualizes unification states.
+
+GeckoGraph positions itself similarly to these projects, using color, shapes, and icons to provide easily interpretable visual cues. 
+However, GeckoGraph focuses specifically on type-level information—an area largely unexplored by previous work. 
+In addition, GeckoGraph was evaluated through a larger and more diverse study, including participants with a wider range of programming experience.
+
+\subsection{Visual vs. Textual Representation}
+\label{sec:visual-vs-textual}
+
+Many studies have compared visual programming environments with textual ones. 
+Research~\cite{Noone2018-wl, Da_Silva_Ribeiro2014-tm, Cliburn2008-jo, Daly2011-is} consistently shows that students taught with visual programming languages demonstrate greater confidence, retention, and enjoyment compared to those using purely textual languages.
+While following a similar trend, GeckoGraph is designed as a complement to text-based notation rather than a replacement for it.
+
+Several tools have explored the effect of visual augmentation—adding visual representations alongside textual code. 
+For example, Greenfoot~\cite{Montero2010-uh} integrates visual and textual programming elements to support learning, while PILeT~\cite{Alshaigy2015-wy} adapts its presentation style to users’ preferences.
+Both systems report positive effects of visual augmentation. 
+Although GeckoGraph also combines visual and textual representations, it differs in that prior studies focused on imperative languages such as Java and Python, whereas our evaluation investigates its effect in a functional programming context (Haskell).
+
+% \subsection{Gamified Human Study}
+
+% It is important to recognize that the human study is designed to be a series of game-like experience.
 The tasks are meant to contain entertaining values.
-We practiced multiple gamification techniques: levels, story/theme, and goals/rewards.
-\cite{Hamari2014-mc} This not only allowed us to have confidence that participants are motivated to complete the tasks, but it also lent us popularity in the Haskell community and led to a historically high participation rate.
+We practiced multiple gamification techniques: levels, story/theme, and goals/rewards \cite{Hamari2014-mc}.
+This not only allowed us to have confidence that participants are motivated to complete the tasks, but it also lent us popularity in the Haskell community and led to a historically high participation rate.
 Gamification has been shown to improve engagement and motivation.
@@
 We identify that studies on functional programming are often technical and intimidating; our use of gamification attracted historically high participation with a wide distribution of experience levels.
-
-\subsection{Potential Applications}
-\subsubsection{Programming Assistance}
-We envision many ways GeckoGraph can be integrated into programming tools.
-GeckoGraph can visualize and inspect types in tooltips and pop-ups.
-It can be used to discover the mismatching parts of two conflicting types in type errors.
-It can be used to generate type expressions and edit type expressions structurally.
-In our post-study survey, the potential integration of text editors and programming assistance were the most requested use cases proposed by the participants.
-
-\subsubsection{Documentation Assistance}
-From what we have learned from our human study, GeckoGraph is well suited to support the documentation of the programming library and the API documentation.
-It works in tandem with the traditional text-based language and can be generated mechanically, making it possible to standardize with minimal effort.
-For documentation sites that allow searching by name (e.g., Hoogle \cite{Mitchell_undated-fb}), programmers often need to sieve through a list of identically named functions.
-For example, a simple Hoogle search for the name \texttt{make} shows a list of functions with vastly different usage and purpose.
-GeckoGraph can help speed up the selection process by providing a visual notation for each type, and programmers can use a visual grouping of colors, sizes, and positions to home in on the correct documentation page.
-
-\subsubsection{Pedagogical Applications}
-
-We believe that GeckoGraph can be a valuable tool in teaching techniques and theories in programming languages that are difficult to convey in plain language.
-In fact, many participants in our study reported that they had no prior knowledge of Haskell programming and that they could understand the programming concepts in the game and complete all the puzzles with the help of GeckoGraph.
-
-Furthermore, the advanced features of GeckoGraph (Section~\ref{sec:benefits}) are also suitable for teaching and learning high-level functional programming concepts.
-Consider the \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension (Fig.
-\ref{fig:assoc}).
-Although the type signature is short, it is very difficult to trace the semantics mentally due to the number of variables, and their kinds are not obvious from the text-based notation.
-GeckoGraph makes understanding the type easier by exposing the ``hidden" higher-kinded types, revealing all the partially applied data types in play.
-
-\begin{figure}[]
-  \includegraphics[width=\linewidth]{figures/assoc}
-  \caption{\label{fig:assoc} The \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension.
-Even for people who are not familiar with the exact definitions, it is easy to see that the variables \texttt{f},  \texttt{g}, and \texttt{h} are all high-kinded types, which is hard to tell from the text-based type annotation.}
-\end{figure}
-
-\section{Related work}
-\subsection{Visualizing polymorphic types}
-A similar half-enclosing structure was proposed in the visualization of types by Jung \cite{Jung2000-oc}.
-In Jung's notation, the type constructor half encloses its arguments, but the figure for the type constructor is placed on the bottom right (Fig.
-\ref{fig:jung}).
-In contrast, GeckoGraph follows the natural reading order, allowing larger structures in a type signature to take precedence over smaller ones.
-
-Compared to functions in Jung's notation,  GeckoGraph shows two major benefits.
-First, GeckoGraph naturally translates the general notion of a curried function.
-Partially, the application of a function can be read as removing the first one of its arguments.
-This is not the case with Jung's notation.
-Second, the shape of a function type remains consistent with the shape of normal data types.
-In Jung's design, a function \texttt{a -> b} looks sufficiently distinct from a data type \texttt{f a b}.
-This is important because, in functional languages, it is very common for abstraction to be drawn from functions and normal data types.
-For example,  functions and lists both have a functor instance, and their inner values can be altered using a \texttt{fmap} function.
-The consistent shape of GeckoGraph makes this generalization easier to see visually.
-
-\begin{figure}[]
-  \includegraphics[width=\linewidth]{figures/Jung}
-  \caption{
-        \label{fig:jung}
-        Comparing the map function (\texttt{(a -> b) -> [a] -> [b]}) in text notation, GeckoGraph, and Jung's notation.
-  }
-\end{figure}
-
-%Many studies \cite{Jun2000-ec, Jun2000-yu} explored the challenges of how human experts use polymorphic types.
-The authors then show a history-preserving type inference algorithm and explanation generator that can explain the types in human language.
-Similarly, \cite{Beaven1993-ay} showed a system that explains the infeasibility of types using human-like languages, explaining the type inference process using a series of "Why" questions and "How" questions.
-Both studies use natural language to reduce the challenge of understanding polymorphic types.
-While these studies are promising, explaining polymorphic types with natural language has limitations.
-They are generally verbose and have to use another visualization assist (text decoration, icons, or indentation) to clarify the naturally hierarchical information.
-A very similar graphical type representation \cite{Jung2000-oc} that utilizes blocks and colors was proposed.
-In the paper, the visualization uses 2D blocks to represent types and enclosures to represent type constructors and type arguments.
-We have discussed thoroughly the difference in design between Jung's notation and GeckoGraph.
-In addition, the evaluation of Jung's notation uses fixed questions and answers, while GeckoGraph is evaluated using real programming tasks.
-
-\subsection{Visualization in programming}
-Using visualization techniques to improve the comprehension of polymorphic types is not new.
-This is often practiced to represent document properties, runtime information, and static analysis results.
-FluidEdt \cite{Ou2015-vr} displays heap graphs in the left margin.
-I3 \cite{Beck2015-my} offers search similarity and change history in compact block-based diagrams.
-Almeida et al.
-~\cite{Almeida2022-bv} introduced a novel visualization to aid in understanding ownership and borrowing in the Rust language.
-While the field of graphical type representation is relatively narrow, it has been studied to some extent.
-Clerici et al.
-~\cite{Clerici2013-ru} proposed a graph-based type inference system that shows the visual representations of unification states.
-GeckoGraph positions itself similarly to these projects, using color, shapes, symbols, and icons to provide easy-to-glance information.
-However, GeckoGraph focuses on type-level information, which no other research projects do.
-In addition, GeckoGraph is evaluated in a much larger study than the other projects, and it is also evaluated with a wider range of experience levels.
-
-\subsection{Visual vs Textual representation}
-
-Many studies have compared the effectiveness of a visual-based programming environment with a textual-based one.
-Studies \cite{Noone2018-wl, Da_Silva_Ribeiro2014-tm, Cliburn2008-jo, Daly2011-is} found that compared to a purely textual programming language with similar positioning, students who were taught a visual programming language show greater confidence, better retention, and enjoyment in programming courses.
-While showing a similar trend, GeckoGraph experiments in the context of accompanying text-based notation rather than replacing it.
-
-Many have studied the effect of visual augmentation, providing a visual representation of programming objects without removing the text-based notation.
-Greenfoot \cite{Montero2010-uh} allows visual and textual representations of programming concepts to be accessible to the learner.
-PILeT \cite{Alshaigy2015-wy}, providing a programming environment that is an adaptive presentation based on the user's preference.
-Both tools show positive results in the use of visual augmentation.
-Although similar to GeckoGraph in combining visual language and text-based programming environment, both studies evaluated the effect based on imperative languages (Java and Python), while our evaluation focused on the effect on a functional language (Haskell).
 
@@
 We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of.
-The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.
+Although no significant difference were found in the overall time
+and success rate between the programmers use GeckoGraph and the ones
+who do not. We find that GeckoGraph improve the success rate for novice users in tackling more difficult tasks.
 
@@
 GeckoGraph focuses on a faithful view of the tree structure of programming types.
-However, many more areas and types demand a human-focused approach.
-For instance, visualizing the ordinal relationship of subsumption or visualizing the numeric changes in dependently typed ``type programs".
+However, many more areas and types demand a visual approach.
+For instance, visualizing the ordinal relationship of subsumption or visualizing the numeric values in dependently typed ``type programs".
 
@@
 \appendix
-\section{Game levels} \label{levels}
+\section{User study tasks} \label{levels}
 We provide all the level settings we used in our user study.
-The online game is open source and available for evaluation \footnote{https://zerotohero.fly.dev}.
+The tasks are open source and available for evaluation\footnote{https://zerotohero.fly.dev}.
 However, this can be attempted locally with a Haskell interpreter or even with a pen and paper.

### 2025-10-21 main.tex
New changes

@@
 They allow functions to be written in a way that can be used with different data types, while still enforcing the relationship and constraints between the values.
-However, programmers often find polymorphic types difficult to use and understand and tend to reason using concrete types.
+However, programmers often find polymorphic types difficult to use and understand, and tend to reason with a mindset of concrete types.
 We propose GeckoGraph, a graphical notation for types.
@@
 To our knowledge, this is the largest controlled user study on functional programming ever conducted.
-Overall, no significant difference were found  in time and success rate between the programmers use GeckoGraph and the ones who do not.
-However, we find that in the most difficult task of the user study, GeckoGraph improve the success rate for novice users (14.5\%) and less experienced users (11.2\%), suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
+Our results show that no significant difference was found in time and success rate between the programmers who use GeckoGraph and those who do not.
+However, we also find that for the most challenging task of the user study, GeckoGraph improves the success rate by 14.5\% for beginners and 11.2\% for less experienced users, suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
 \end{abstract}
@@
 \ref{fig:example-foldable}).
-Unfortunately, not many research focuses on the \textit{usability} of polymorphic types.
+Unfortunately, not many studies focus on the \textit{usability} of polymorphic types.
 
@@
 GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
-We designed a gamified online study  \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming puzzles using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
-Overall, no significant difference were found  in time and success rate between the programmers use GeckoGraph and the ones who do not.
-However, we find that in the hardest task of the user study, GeckoGraph improve the success rate for novice users (14.5\%) and less experienced users (11.2\%), suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
+We designed a gamified user study \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming puzzles using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
+Overall, no significant difference was found in either time or success rate between the programmers who use GeckoGraph and those who do not.
+However, we find that in the hardest task of the user study, GeckoGraph improves the success rate by 14.5\%  for novice users (less than 1 year of experience) and 11.2\% for less experienced users (1-3 years of experience), suggesting the current iteration of GeckoGraph may be useful as a teaching and learning tool.
 
@@
 
-Static type systems are widely used in large and complex software projects because they help detect errors early and make refactoring safer.
+Functional programming languages have a rich history of embracing strong type systems and pioneering advanced type system features \cite{Milner1978-ob, MacQueen2020-qn, Hudak2007-kn}.
 In functional programming languages, type systems are often more expressive and provide stronger safety guarantees.
@@
 
-Unlike the nominal type systems found predominately in languages such as Java and C \cite{AbdelGawad2013-hh}, many functional languages heavily rely on structural typing.
+Unlike the nominal type systems found predominantly in languages such as Java and C \cite{AbdelGawad2013-hh}, many functional languages heavily rely on structural typing.
 In nominal systems, two types are considered distinct even if they have identical structure but different names.
@@
 This approach promotes flexibility, as types can be defined inline without introducing new names, and values are automatically valid for a type when they conform to the structure.
-However, in functional languages such as Haskell, and OCaml, the lack of nominality can also make complex types harder to read and reason about.
-For instance, when encounter type errors, programmers can face two unmatching types with complex structure, understanding where exactly the root error is becomes a tediotasksks of string comparison.
+However, this lack of nominality can also make complex types harder to read and reason about.
+For instance, when encountering type errors, programmers can face two unmatching types with a complex structure, and understanding where exactly the root error is becomes a tedious task of string comparison.
 
@@
 Algebraic data types (ADTs) \cite{Milner1978-qz} allow developers to construct complex types from simpler ones using sum and product operations.
-A sum types, often called tagged union, $A + B$ represent a type that can be a value of A or a value of B.
-Conversely, a product type $A \times B$ represent a tuple of A and B.
-This precise and composable type definitions for common data structures such as lists, tuples, and result types.
-For example, in Haskell, types like \texttt{Either a b} and \texttt{Maybe a} provide basic building blocks for error handling, and the type \texttt{data List a = [] | a : List a } provides a succinct recursive definitions for lists.
-This illustrates both the expressive power and the potential complexity of ADTs.
-In practice, simple type definitions can turn into an \textit{essay of types} \cite{Ferdowsi2023-cv} as program grows, causing difficulties to interpret and maintain.
+A {\bf sum type}, often called a tagged union, $A + B$, represents a type that can be a value of A or a value of B.
+Conversely, a {\bf product type} $A \times B$ represents a tuple type composed of two values of both {\tt A} and {\tt B}.
+This formalization allows precise and composable type construction for common data structures such as lists, tuples, and result types.
+For example, in Haskell, types like \texttt{data Either a b = Left a | Right b} and \texttt{data Maybe a = Nothing | Just a} provide basic building blocks for error handling, and the type \texttt{data List a = [] | a : List a } provides a succinct recursive definition for lists.
+This illustrates the expressive power of ADTs.
+However, in practice, type definitions can grow to be complex structures, for instance \texttt {Maybe (Either [Maybe a] (Either a String))} as programs grow, causing difficulties to interpret and maintain.
 
@@
 Parametric polymorphism (often called generics) allows functions and data structures to operate uniformly across different types.
-For instance in Haskell, the\texttt{head :: [a] -> a}, taking a list as input and return the first element, is a polymorphic functions that works on lists of any type, integers, strings, any type is valid.
+For instance, in Haskell, the function \texttt{head :: [a] -> a}, taking a list as input and returning the first element, is a polymorphic function that works on lists of any type; integers, strings, and functions are all valid candidates for the generic variable \texttt{a}.
 It provides a strong and principled form of abstraction, enabling code reuse without sacrificing type safety.
 
-However, Parametric polymorphism also increases cognitive load: programmers must mentally track type variables and reason about how concrete types are instantiated during function application.
-Overuse can make code difficult to follow, especially for readers unfamiliar with the underlying type relationships.
+However, parametric polymorphism also increases cognitive load: programmers must mentally track type variables and reason about how concrete types are instantiated during function application.
+This task becomes more challenging when functions involve multiple type variables.
+An example would be the foldM function: \texttt{foldM :: (Monad m) => (a -> b -> m a) -> a -> [b] -> m a}.
+The challenge of reading and understanding polymorphic types is more serious for novice users \cite{Jun2000-ec, Jun2000-yu}.
 
@@
 Most modern languages employ some level of inference.
-For instance, Java’s var keyword (introduced in version 10) infers variable types at compile time based on their assigned values \cite{Java_Developers2023-an}, saving programmers from writing redundenies like \textit {Car car = new Car()}.
+For instance, Java’s var keyword (introduced in version 10) infers variable types at compile time based on their assigned values \cite{Java_Developers2023-an}, saving programmers from writing redundant type declarations, such as \textit {Car car = new Car()}.
 
@@
 While this improves conciseness, it can reduce code readability and produce complex error messages.
-Many developers therefore choose to write explicit type signatures for clarity, even when inference is available.
+Many programmers, therefore, choose to write explicit type signatures for clarity, even when inference is available.
 
@@
 However, this rigor introduces trade-offs: a steeper learning curve and less flexibility for exploratory or prototype development.
-In the following sections, we demonstrate how GeckoGraph supports these advanced type-system features by visualizing type level information that is often implicit or difficult for humans to interpret.
+In the following sections, we explore how GeckoGraph supports these type-system features by visualizing type-level information that is often implicit or difficult for humans to interpret.
 
@@
 
-Motivated by the design goals, we designed GeckoGraph with the following feature features and construction rules (Fig. \ref{fig:design}):
+Motivated by the design goals, we designed GeckoGraph with the following features and construction rules (Fig. \ref{fig:design}):
 
@@
 
-\paragraph{Complex types}
+\paragraph{Algebraic Data Types}
 GeckoGraph displays an algebraic data type as a larger cell, where the type constructor half encloses its arguments: \includegraphics[height=1em]{figures/DataType.png}.
@@
 Functions are the fundamental building blocks of functional programming languages, and function types are ubiquitous and the most important in type-level programming.
-In Haskell, \texttt{(->)} is defined as an infix type operator with the right associativity to provide succinct type annotation.
+In Haskell, \texttt{(->)} is defined as an infix type operator with right associativity to provide succinct type annotations.
 GeckoGraph preserves this syntax feature to make the notation more intuitive (design goal \ref{goal1}): the 2 arguments of a function type in GeckoGraph are placed on both sides of the cell \includegraphics[height=1em]{figures/Function.png}.
@@
 Regular high-order functions (e.g., \texttt{(a -> b) -> c}) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
-The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and high-order functions (design goal \ref{goal3}).
-It is easy to tell high-order functions from the vertical position of its function indicator.
-Similarly, it is easy to count the arity of the outer function by counting the number of function indicators at the highest layer.
+The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and higher-order functions (design goal \ref{goal3}).
+It is easy to tell high-order functions from the vertical position of the function indicators.
+Similarly, it is easy to count the arity of a function by counting the number of function indicators at the outer layer.
 For example, in Fig.
-\ref{fig:indicator}, the higest layer function indicator (the ones colored in the lightest shade of gray) suggest a ternary function.
-The two arguments types and the return type are all functions.
+\ref{fig:indicator}, the outer layer function indicators (the ones colored in the lightest shade of gray) suggest a ternary function.
+The two argument types and the return type are all functions.
 
@@
         \label{fig:indicator}
-        An example of visual hierarchy of the function indicator.
+        An example of the visual hierarchy of the function indicator.
 The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators.
@@
 The GeckoGraph type class's design promotes the type class placements rather than the type class names.
-Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or pop-up window) to identify the name of the type class.
+Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or a pop-up window) to identify the name of the type class.
 We believe that this is a well-justified trade-off.
-The strength of this approach lies in situations when type class context is complex.
+The strength of this approach lies in situations when the type class context is complex.
 When reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable.
@@
 A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers.
-GeckoGraph uses qualitative color scheme from ColorBrewer \cite{Harrower2003-vu} to encode symbolic values of types.
+GeckoGraph uses a qualitative color scheme from ColorBrewer \cite{Harrower2003-vu} to encode symbolic values of types.
 One example of GeckoGraph's use of color is the ``rotation" function  (Fig.
@@
   \includegraphics[width=0.6\linewidth]{figures/rotate}
-  \caption{\label{fig:rotate} The `f2' function in the user study. It shuffles the type arguments one step to the left and the leftmost argument wraparound to be the rightmost.}
+  \caption{\label{fig:rotate} The `f2' function in the user study. It shuffles the type arguments one step to the left, and the leftmost argument wraps around to be the rightmost.}
 \end{figure}
 
-\textbf{The horizontal axis of GeckoGraph} often are helpful in identifying differences in function arities.
+\textbf{The horizontal axis of GeckoGraph} is often helpful for identifying differences in function arities.
 For example, in Fig.
@@
 \textbf{The vertical axis of GeckoGraph} often helps identify the number of levels in a nested complex type.
-This can often be very useful when inspecting mismatching type errors where data types are nested in different levels.
+This can often be very useful when inspecting mismatched types where data types are nested at different levels.
 Common examples include when programmers forget to apply the value to ``return" in a monadic block or to use \texttt{liftIO} to cast an \texttt{IO} effect.
@@
 The design of GeckoGraph enables the visualization of many advanced type-level features.
-\textbf{Kind visualization}: The kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is a type constructor (e.g.
-[], Maybe) or a type constant (e.g.
-Bool, Char), because of this, kind is often described as the type of types.
+\textbf{Kind visualization}: The kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is a type constructor (e.g., [], Maybe) or a type constant (e.g., Bool, Char).
+Because of this, kind is often described as the type of types.
 GeckoGraph infers the {\it kind} of type variables and consistently displays this information.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \includegraphics[width=\linewidth]{figures/Advanced}
@@
 
-The experiment aims to study how programmers use and reason about polymorphic type when performing programming tasks.
+The experiment aims to study how programmers use and reason about polymorphic types when performing programming tasks.
 In particular, we studied how programmers scan and select potentially useful functions from a library and how they compare intended types and actual types when type errors are encountered.
@@
 The possible solution and other details of the level can be found in Appendix \ref{levels}.
-To complete this level, players only need to use each of the provided function once.
+To complete this level, players only need to use each of the provided functions once.
 One possible solution to this task would be {\tt zeroToHero z = makeHero (runZero z)}.
@@
   \includegraphics[width=\linewidth]{figures/Level3}
-  \caption{\label{fig:level3} A screenshot showing the level 3 of the user study.
+  \caption{\label{fig:level3} A screenshot showing level 3 of the user study.
 In this level, the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero (a a)}.
@@
 
-The experiment uses a between-subject design.
+The experiment uses a between-subjects design.
 Participants are assigned to one of two groups.
 The tasks assigned to the two groups are in the same order.
-Participants receive both treatments (with and without GeckoGraph) during their runs but in a different order.
+Participants receive both treatments (with and without GeckoGraph) during their runs, but in a different order.
 Group one is assisted by GeckoGraph on even levels and text-only type annotation on odd levels.
-Group two participants are the same but with the order flipped.
+Group two participants are the same, but with the order flipped.
 The number of participants in the two groups is counterbalanced.
@@
 Participants can skip a level if they are stuck.
-We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their feedback of the tool.
-The number of skips a participant can use is limited to four so that it is not possible to submit qualitative feedback without completing at least 6 tasks out of 10.
+We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their feedback on the tool.
+The number of skips a participant can use is limited to four, so that it is not possible to submit qualitative feedback without completing at least 6 tasks out of 10.
 
@@
 Data collection from the study was stopped after the planned cut-off period of 14 days.
-After the cut-off date, the ZeroToHero game is open source and available for free evaluation\footnote{https://zerotohero.fly.dev} and repeating our experiment, but no further data was collected.
+After the cut-off date, the ZeroToHero game is open source and available for free evaluation\footnote{https://zerotohero.fly.dev}, and repeating our experiment, but no further data was collected.
 
@@
 This roughly aligns with self-reported expertise.
-We show that the task time on each level follows normal distributions using a Shapiro-Wilk test \cite{Shaphiro1965-dx} (p-value  $ \leq 1.018 \times 10^-16$, for an alpha value of 0.05, p less than 0.05 is considered normal distribution).
+We show that the task time on each level follows normal distributions using a Shapiro-Wilk test \cite{Shaphiro1965-dx} (p-value  $ \leq 1.018 \times 10^-16$, for an alpha value of 0.05, p less than 0.05 is considered a normal distribution).
 
-Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience.
-Level 5 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple.
+Levels 5, 7, and 8 are the only three levels that include functions from the standard Haskell library other than the generic \texttt{(.)} and \texttt{(\$)}, which are provided for pure convenience.
+Level 5 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract a value from a tuple.
 Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class.
@@
 However, when comparing the task time between the two treatments, we were unable to reject the null hypothesis.
-In a two-sample T-test, we could not find any significant difference between the two groups overall (p-value = 0.457), nor does there differ between the two groups in any of the four levels of experience (beginner: p-value = 0.845, familiar: p-value = 0.524, Knowledgeable: p-value = 0.712, expert p-value = 0.771).
+In a two-sample T-test, we could not find any significant difference between the two groups overall (p-value = 0.457), nor did there differ between the two groups in any of the four levels of experience (beginner: p-value = 0.845, familiar: p-value = 0.524, Knowledgeable: p-value = 0.712, expert: p-value = 0.771).
 
@@
 We show that the difficulty steadily increases across the user study, but levels 5, 7, and 8 are significantly harder than the authors intended.
-The overall task time of each group roughly matches experience level.}
+The overall task time of each group roughly matches the experience level.}
 \end{figure}
@@
 \ref{fig:qualitative} right).
-It is slightly positive for beginners (4.25) and slightly negative for the other groups, the familiar group (3.86) and the knowledgeable group (3.32).
-The expert group finds that GeckoGraph is relatively unhelpful, with an average score of 2.95.
+It is slightly positive for beginners (mean=4.25) and slightly negative for the other groups, the familiar group (mean=3.86) and the knowledgeable group (mean=3.32).
+The expert group finds that GeckoGraph is the least helpful (mean=2.95)
 
@@
   \includegraphics[width=\linewidth]{figures/Qualitative}
-  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph, with 95\% confidence interval.} Overall, programmers are neutural to GeckoGraph's design.
+  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph, with 95\% confidence interval.} Overall, programmers are neutral to GeckoGraph's design.
 Some find GeckoGraph to be intuitive and not distracting.
@@
 
-% We start this section with a strength and weakness analysis of GeckoGraph.
-The disuccion is based on our results of user study.
-Then, In the end, we speculate the applications of GeckoGraph.
-The envisioned applications are grounded in the strengths and weaknesses of GeckoGraph.
 \section{Discussion} \label{sec:discussion}
@@
 \subsection{Strengths}\label{sec:discussion:strengths}
-Overall, our experimental results show statistically significant difference in task completion time between conditions with and without GeckoGraph.
-However, we find using GeckoGraph significantly improves task success rates in the hardest task (task 10) across the two less experienced user groups by 14.5\% and 11.2\% respectively.
+Overall, our experimental results show a statistically significant difference in task completion time between conditions with and without GeckoGraph.
+However, we find using GeckoGraph significantly improves task success rates in the hardest task (task 10) across the two less experienced user groups by 14.5\% and 11.2\%, respectively.
 To interpret the observed advantages, we discuss the practical benefits that GeckoGraph offers to programmers.
@@
 \subsubsection{Programming Assistance}
-Post-study feedback highlighted strong interest in integrating GeckoGraph into programming environments. 
+Post-study feedback highlighted a strong interest in integrating GeckoGraph into programming environments. 
 Modern IDEs already allow users to inspect expression types via hover tooltips; GeckoGraph could enhance this feature by visualizing type structures directly in the tooltip. 
@@
 Because GeckoGraph representations can be generated automatically, integration into documentation systems like Hoogle \cite{Mitchell_undated-fb} would require minimal effort. 
-A quick visual scan on the page can quickly review the theme of a library or package, i.e., the domains and codomains most functions operate on.
-For instance, a library mostly support string operation should predominately show the color of {\tt String} type, while a low-level graphic related library may show mainly the colors for {\tt Vector} and {\tt Float}.
+A quick visual scan of the page can quickly review the theme of a library or package, i.e., the domains and codomains most functions operate on.
+For instance, a library that mostly supports string operations should predominantly show the color of {\tt String} type, while a low-level graphics-related library may show mainly the colors for {\tt Vector} type and {\tt Float} type.
 
@@
 This section reviews prior work in three areas relevant to this research. First, we examine approaches to the visualization of types in programming languages. 
-We than broaden our view to general visualization techniques applied to programming environments, including topics like memory allocation and change history.
-Finally, we reviews studies comparing visual and textual representations in programming education and environments.
+We then broaden our view to general visualization techniques applied to programming environments, including topics like memory allocation and change history.
+Finally, we review studies comparing visual and textual representations in programming education and environments.
 
@@
 More recent work, TYPEical~\cite{Moy2020-tr}, provides an interactive visualization of function types for exploring and comparing functions in R packages. 
-This project is innovative in that it uses a Sankey diagram~\cite{Keiderling2012-ag} to illustrate the flow between collections of functions and the data types they operate on, as well as the frequency of their use, providing a high level overview of the usage statistics.
+This project is innovative in that it uses a Sankey diagram~\cite{Keiderling2012-ag} to illustrate the flow between collections of functions and the data types they operate on, as well as the frequency of their use, providing a high-level overview of the usage statistics.
 GeckoGraph shares TYPEical’s vision that visualization should support three key tasks: \textbf{finding functions}, \textbf{determining types}, and \textbf{comparing types}. 
@@
 However, GeckoGraph focuses specifically on type-level information—an area largely unexplored by previous work. 
-In addition, GeckoGraph was evaluated through a larger and more diverse study, including participants with a wider range of programming experience.
+In addition, GeckoGraph was evaluated through a larger and more diverse study, including participants with a broader range of programming experience.
 
@@
 
-% \subsection{Gamified Human Study}
-
-% It is important to recognize that the human study is designed to be a series of game-like experience.
-The tasks are meant to contain entertaining values.
-We practiced multiple gamification techniques: levels, story/theme, and goals/rewards \cite{Hamari2014-mc}.
-This not only allowed us to have confidence that participants are motivated to complete the tasks, but it also lent us popularity in the Haskell community and led to a historically high participation rate.
-Gamification has been shown to improve engagement and motivation.
-This has been harnessed by many research projects to improve participation in human studies \cite{He2014-vp}.
-We identify that studies on functional programming are often technical and intimidating; our use of gamification attracted historically high participation with a wide distribution of experience levels.
-
 \section{Conclusion and Future Work}
@@
 We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of.
-Although no significant difference were found in the overall time
-and success rate between the programmers use GeckoGraph and the ones
-who do not. We find that GeckoGraph improve the success rate for novice users in tackling more difficult tasks.
+Although no significant effect of GeckoGraph was found on the overall time and success rate, we find that GeckoGraph significantly improves the success rate for less experienced users in tackling more difficult tasks.
 

### 2025-10-24 main.tex
Updates

@@
 \begin{abstract}
-Polymorphic types are an important feature in most strongly typed programming languages.
+Polymorphic type is an important feature in most strongly typed programming languages.
 They allow functions to be written in a way that can be used with different data types, while still enforcing the relationship and constraints between the values.
@@
 We propose GeckoGraph, a graphical notation for types.
-GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier.
-We conducted a large-scale human study using GeckoGraph compared to text-based type notation.
+GeckoGraph aims to complement traditional text-based type notation by making reading, understanding, and comparing types easier.
+We conducted a large-scale user study using GeckoGraph, comparing it to text-based type notation.
 To our knowledge, this is the largest controlled user study on functional programming ever conducted.
-Our results show that no significant difference was found in time and success rate between the programmers who use GeckoGraph and those who do not.
-However, we also find that for the most challenging task of the user study, GeckoGraph improves the success rate by 14.5\% for beginners and 11.2\% for less experienced users, suggesting the current iteration of GeckoGraph may be more useful as a teaching and learning tool.
+The results show no overall significant differences in task completion time or success rate.
+However, we observed a modest improvement in success rate among beginner participants.
+For the most challenging task, GeckoGraph improved success rates by 14.5\% for beginners and 11.2\% for less experienced users.
+These findings suggest that GeckoGraph has potential as a teaching and learning aid for polymorphic type systems.
 \end{abstract}
@@
 GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
-We designed a gamified user study \textit{ZeroToHero}\footnote{Available at https://zerotohero.fly.dev}, evaluating 714 participants' performance to solve a series of programming puzzles using GeckoGraph versus text-based type annotation, which is, to our knowledge, the largest controlled user study of a functional programming tool or concept.
-Overall, no significant difference was found in either time or success rate between the programmers who use GeckoGraph and those who do not.
-However, we find that in the hardest task of the user study, GeckoGraph improves the success rate by 14.5\%  for novice users (less than 1 year of experience) and 11.2\% for less experienced users (1-3 years of experience), suggesting the current iteration of GeckoGraph may be useful as a teaching and learning tool.
+ To evaluate its effectiveness, we conducted a large-scale user study comparing GeckoGraph with standard text-based type notation.
+To our knowledge, this represents the largest controlled user study on functional programming to date.
+The results show no overall significant differences in task completion time or success rate.
+However, we observed a modest improvement in success rate among beginner participants.
+For the most challenging task, GeckoGraph improved success rates by 14.5\% for beginners and 11.2\% for less experienced users.
+These findings suggest that GeckoGraph has potential as a teaching and learning aid for polymorphic type systems.
 
@@
 
-\subsection{Structural Type Systems}
-
-Unlike the nominal type systems found predominantly in languages such as Java and C \cite{AbdelGawad2013-hh}, many functional languages heavily rely on structural typing.
+{\bf Structural typing} is often the primary approach for determining type equivalence and compatibility in functional programming languages (e.g., Elm and OCaml) and multi-paradigm languages (e.g., TypeScript).
+In contrast, nominal type systems are often predominant in Object-Oriented languages such as Java and C\# \cite{AbdelGawad2013-hh}.
 In nominal systems, two types are considered distinct even if they have identical structure but different names.
@@
 However, this lack of nominality can also make complex types harder to read and reason about.
-For instance, when encountering type errors, programmers can face two unmatching types with a complex structure, and understanding where exactly the root error is becomes a tedious task of string comparison.
+For instance, when encountering type errors, programmers can face two mismatched types with a complex structure, and understanding where exactly the root error is becomes a tedious task of string comparison (Fig.
+\ref{fig:structural}).
 
-\subsection{Algebraic Data Types}
+\begin{figure}[]
+  \includegraphics[width=\linewidth]{figures/structural}
+  \caption{\label{fig:structural} Example of a TypeScript structural type error.
+(Top) The function \texttt{f} expects a parameter with a string field \texttt{d}, but the argument {\tt x} provides a numeric {\tt d}.
+(Bottom) The error message is dreadful to read.}
+\end{figure}
 
-Algebraic data types (ADTs) \cite{Milner1978-qz} allow developers to construct complex types from simpler ones using sum and product operations.
+{\bf Algebraic data types} (ADTs) \cite{Milner1978-qz} allow developers to construct complex types from simpler ones using sum and product operations.
 A {\bf sum type}, often called a tagged union, $A + B$, represents a type that can be a value of A or a value of B.
@@
 
-\subsection{Parametric Polymorphism}
-
-Parametric polymorphism (often called generics) allows functions and data structures to operate uniformly across different types.
+{\bf Parametric polymorphism} (often called generics) allows functions and data structures to operate uniformly across different types.
 For instance, in Haskell, the function \texttt{head :: [a] -> a}, taking a list as input and returning the first element, is a polymorphic function that works on lists of any type; integers, strings, and functions are all valid candidates for the generic variable \texttt{a}.
@@
 
-\subsection{Type Inference}
-
-Type inference, or implicit typing, reduces the need for explicit type annotations by allowing the compiler to deduce types automatically.
+{\bf Type inference}, or implicit typing, reduces the need for explicit type annotations by allowing the compiler to deduce types automatically.
 Most modern languages employ some level of inference.
@@
 
-To summarize, the type systems of functional programming languages provide rigorous tools for specifying and enforcing program behavior.
+{\it To summarize}, the type systems of functional programming languages provide rigorous tools for specifying and enforcing program behavior.
 By encoding system constraints directly in types, they prevent many classes of runtime errors.
@@
 Each type class required on a type variable displays as a square indicator aligned on the right of the extended area.
-In GeckoGraph, a type-class gliph is embedded in every type variable that requires this class.
+In GeckoGraph, a type-class glyph is embedded in every type variable that requires this class.
 This means when displaying the type \texttt{(==) :: Eq a =>
@@
 
-\begin{figure}[]
+\begin{figure}[h]
   \includegraphics[width=0.6\linewidth]{figures/Add3}
@@
 Because of its graphical layout, GeckoGraph allows programmers to precisely select which part of a type signature they intend to query, that is, in Fig.
-\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a sepecific query that is hard to execute with point and click over text based type annotation.
+\ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a specific query that is hard to execute with point and click over text based type annotation.
 
@@
   \caption{\label{fig:advanced}
-  Advanced features of GeckoGraph.
-(A) GeckoGraph supports Kind Visualization if the inferred kind is greater than \texttt{*}.
-(B) GeckoGraph supports qualified constraints by extending the extended area across multiple type variables.
-(C) GeckoGraph supports Multiple Parameter Type Classes, using different shapes of the same color to indicate that multiple variables must satisfy certain type classes collectively.
-(D) GeckoGraph supports the precise selection of its sub-structures.
-}
+  Advanced features of GeckoGraph. (A) Kind visualization. (B)  Qualified constraints. (C)  Multiple Parameter Type Classes. (D) Precise selection of sub-structures. }
 \end{figure}
@@
 \section{Evaluation} \label{sec:evaluation}
-To evaluate the usefulness of GeckoGraph, we designed and conducted a gamified controlled experiment.
-The study, "Zero to Hero", contains 10 tasks (levels) of varying difficulty.
+To evaluate the effectiveness of GeckoGraph, we designed and conducted a gamified controlled experiment.
+The study, ``Zero to Hero", adopted a game-like experience.
+It contains 10 tasks (levels) of varying difficulty.
 At each level, participants are asked to implement a function named "\texttt{zeroToHero}" with the help of a limited list of provided functions.
@@
 A generic function \texttt{(\$)} is provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions.
-The possible solution and other details of the level can be found in Appendix \ref{levels}.
+The possible solutions and other details of the level can be found in Appendix \ref{levels}.
 To complete this level, players only need to use each of the provided functions once.
@@
 The user illustrated in Fig.
-\ref{fig:level-example} did not reach a correct solution, as the type signature and GeckoGraphs in {\it To Implement} and {\it Your Result} do not match.
+\ref{fig:level-example} did not reach a correct solution, as the type signature and GeckoGraph in {\it To Implement} panel do not match the one in {\it Your Result} panel.
 
@@
 \ref{fig:level3}),  players need to implement the function {\tt zeroToHero :: Zero a -> Hero (a, a)}.
-In this level, not all provided functions are used in the final result.
+In this level, not all provided functions are useful.
 In particular,  players must discover that only \texttt{f1} and \texttt{f3} are necessary to produce the desired results.
@@
 
+Participants are assigned to one of two groups (A and B).
 The experiment uses a between-subjects design.
-Participants are assigned to one of two groups.
-The tasks assigned to the two groups are in the same order.
-Participants receive both treatments (with and without GeckoGraph) during their runs, but in a different order.
-Group one is assisted by GeckoGraph on even levels and text-only type annotation on odd levels.
-Group two participants are the same, but with the order flipped.
+The tasks assigned to the two groups are the same.
+All participants receive both treatments (with and without GeckoGraph) during their runs, but in a different order.
+{\bf Group A} is assisted by GeckoGraph on even levels and text-based type annotation on odd levels.
+{\it Group B} is assisted by text-based type annotation on even levels and GeckoGraph on odd levels.
 The number of participants in the two groups is counterbalanced.
@@
 \subsection{Task Design} \label{subsection:task}
-During the study, participants receive 10 tasks, represented by the 10 levels of the user study.
+During the study, participants received 10 tasks.
 The tasks start off easy but gradually increase in difficulty.
@@
 If a participant has completed all 10 levels (with the help of skipping), we invite the participant to complete a post-study survey.
-In it, we ask for their opinion on how intuitive the GeckoGraph design is, how distracting they find GeckoGraph, and how helpful GeckoGraph is, using a seven-point scale.
-In the end, we ask a few open-ended questions, inviting participants to provide their experience using GeckoGraph and their expectations about the potential applications of GeckoGraph.
+In it, we ask for them to rate how {\bf intuitive} the GeckoGraph design is, how {\bf distracting} they find GeckoGraph, and how {\bf helpful} GeckoGraph is, using a seven-point scale.
+In the end, we ask 2 open-ended questions: {\it What is your overall experience using GeckoGraph} and {\it What are your expectations on the potential applications of GeckoGraph}.
 
-Data collection from the study was stopped after the planned cut-off period of 14 days.
-After the cut-off date, the ZeroToHero game is open source and available for free evaluation\footnote{https://zerotohero.fly.dev}, and repeating our experiment, but no further data was collected.
+Data collection of the study was stopped after the planned cut-off period of 14 days.
+After the cut-off date, the ZeroToHero website is open-sourced and available for free evaluation\footnote{https://zerotohero.fly.dev}, and repeating our experiment.
+However, no further data was collected.
 
@@
 \subsection{Success rate}
-We saw that, overall, GeckoGraph provides a higher success rate (96.88\%) than text-based type notation (94.62\%).
+We saw that, overall, GeckoGraph provides a slightly higher success rate (96.88\%) than text-based type notation (94.62\%).
 This trend can be seen in every experienced group: beginner group (95.12\% vs.
@@
 When breaking down the result in each task (Fig.
-\ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group and task 10 of the familiar group (Fig.
-\ref{fig:success-rate}).
-We will address this correlation in Section~\ref{sec:discussion}.
+\ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group (94.9\% with GecoGraph, and 80.4\% without GeckoGraph, p-value=0.0452) and task 10 of the familiar group (98.5\% with GecoGraph, and 87.3\% without GeckoGraph, p-value = 0.136).
 
@@
 We speculate that this level demands the mental flexibility typical of a Haskell expert to internalize these functions quickly.
-GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding structural clarity.
+GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding visual clarity.
 
@@
 \subsection{Qualitative Feedback}
-In their responses to the post-study survey, most programmers believe that the design of GeckoGraph is intuitive and that its appearance in the interface does not cause distraction.
-For the question ``Do you find the GeckoGraph distracting", most of the participants rated a negative score, with an average of 2.88 (Fig.
+In their responses to the post-study survey, participants showed mostly neutral sentiments towards GeckoGraph's design and appeal.
+They find GeckoGraph moderately intuitive to use and that its appearance in the interface does not cause distraction.
+
+For the question ``Do you find the GeckoGraph distracting", most of the participants rated a low score, with an average of 2.88 (Fig.
 \ref{fig:qualitative} left).
@@
 \ref{fig:qualitative} Middle): experts find the GeckoGraph most intuitive (5.07), followed by the knowledgeable group (4.87), and the familiar group (4.80).
-The beginner group found it to be the least intuitive but still rated a positive score of 4.71.
+The beginner group found it to be the least intuitive but still rated it positively, albeit with a low score of 4.71.
 When answering the question ``How helpful do you find GeckoGraph in finding the solution?", the answer is more divided into different experience groups (Fig.
@@
 It is slightly positive for beginners (mean=4.25) and slightly negative for the other groups, the familiar group (mean=3.86) and the knowledgeable group (mean=3.32).
-The expert group finds that GeckoGraph is the least helpful (mean=2.95)
+The expert group finds that GeckoGraph is moderately unhelpful (mean=2.95)
 
 Two interesting trends emerge.
-First, the more experienced the users are, the more likely they are to find GeckoGraph intuitive and easy to use.
+First, experienced users are more likely to find GeckoGraph intuitive and easy to use.
 This aligns with our design decision to make GeckoGraph resemble text-based type annotation.
@@
 \paragraph{Task design}
-In our human study, most of the provided functions are very abstract.
+In our user study, most of the provided functions are very abstract.
 These functions are created by the authors solely for the gamified study.
@@
 \subsection{Strengths}\label{sec:discussion:strengths}
-Overall, our experimental results show a statistically significant difference in task completion time between conditions with and without GeckoGraph.
-However, we find using GeckoGraph significantly improves task success rates in the hardest task (task 10) across the two less experienced user groups by 14.5\% and 11.2\%, respectively.
+Overall, our experimental results show no statistically significant difference in task completion time between conditions with and without GeckoGraph.
+However, we find that using GeckoGraph shows a significant albeit moderate effect on the task success rates for beginners.
+Particularly, in the hardest task (task 10), GeckoGraph improved the success rates across the two less experienced user groups by 14.5\% and 11.2\%, respectively.
 To interpret the observed advantages, we discuss the practical benefits that GeckoGraph offers to programmers.
@@
 In this paper, we propose GeckoGraph, a graphical notation for type annotations in functional programming languages.
-GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier.
-We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of.
-Although no significant effect of GeckoGraph was found on the overall time and success rate, we find that GeckoGraph significantly improves the success rate for less experienced users in tackling more difficult tasks.
+GeckoGraph aims to complement traditional text-based type notation by making reading, understanding, and comparing types easier.
+We conducted a large-scale human study using GeckoGraph, comparing it to text-based type notation.
+This is the largest user study on functional programming we are aware of.
+Although no significant effect was found on the overall time to complete each programming task, we found a modest effect on the success rate in the beginner group.
+In one of the more difficult programming tasks, we find that GeckoGraph significantly improves the success rate by 14.5\% for beginners and 11.2\% for less experienced programmers.
 

### 2025-10-24 response.tex
Squashed changes

\documentclass{article}

\newcommand{\pointRaised}[2]{\medskip \noindent 
               \paragraph{#1} \textsl{#2}} 
\newcommand{\reply}{\noindent \textbf{Response}:\ }

\begin{document}

Dear Editor,

Thank you and the reviewers for the opportunity to revise our manuscript “GeckoGraph: A Visual Language for Polymorphic Types.”
We acknowledge the constructive and insightful feedback provided. We think they help improve the clarity, structure, and balance of our paper.
In this revised version, we have addressed all reviewer comments and made corresponding improvements throughout the manuscript. Below is a summary of major changes, followed by detailed point-by-point responses to each reviewer.

We hope our revisions satisfactorily address all concerns and that the paper is now suitable for publication in the Journal of Computer Languages.

Sincerely,

Authors of "GeckoGraph: A Visual Language for Polymorphic Types"

\section{Summary of Major Revisions}

\paragraph{Balanced claims} – We have now moderated the tone in the Abstract, Introduction, and Conclusion to be less ambitious and more balanced, explicitly acknowledging both the strengths and limitations of our findings.

\paragraph{Terminology refinement} – We have now replaced references to the user study as a “game” with more precise wording, such as “gamified user study” or simply “user study.”

\paragraph{New background section} – We have now added a new section, Type Systems in Functional Languages, after the Introduction to provide theoretical background and motivate the design of GeckoGraph.

\paragraph{Section structure overviews} – We have now added section structure overviews at the beginning of the Related Work and Discussion sections to better guide the reader.

\paragraph{Expanded related work} – We have now discussed the relevant study “Just TYPEical: Visualizing Common Function Type Signatures in R” to strengthen the discussion of related approaches and compare to our own contributions.

\paragraph{Enhanced discussion} – We have now improved the Discussion section with more concrete examples demonstrating the potential applications of GeckoGraph.

\paragraph{Concise figure captions} – We have shortened our figure captions throughout the manuscript for readability and stylistic consistency.

\paragraph{Improved flow between sections} – We have added a transitional paragraph linking the Design Goals and Detailed GeckoGraph Features subsections in Section 3.1.

\paragraph{Clarified explanation of type classes} – We have now refined the text in Section 3.1 to improve clarity when explaining type classes.

\paragraph{Removed unused figure} – We removed Figure 4 from the previous version, as it did not add new information to the paper.

\paragraph{Cosmetic and consistency edits} – We improved our grammar, spelling, LaTeX syntax, and overall logical coherence across the manuscript.

\section{Detailed Responses to Reviewers
Response to Reviewer 1}

\pointRaised{Comment 1:}{ The manuscript’s tone in the abstract and conclusion appears overly strong given the presented evidence.}

\reply We appreciate this observation. We have moderated the language in the Abstract, Introduction, and Conclusion to ensure our claims are proportionate to the results, and to present both positive and negative aspects of our findings.

\pointRaised{Comment 2:}{The term “game” used to describe the user study could be misleading.}

\reply We agree that we don't meet all the expectations for a game, and it's more appropriate to frame it as "an entertaining coding challenge". We changed our word of choice for “game” to “gamified user study” or simply, “user study”.

\pointRaised{Comment 3:}{The paper lacks sufficient background on type systems in functional languages.}

\reply We have added a new section titled Type Systems in Functional Languages immediately following the Introduction. This section introduces relevant concepts and better motivates the design of GeckoGraph.

\pointRaised{Comment 4:}{Please clarify the explanation of type classes.}

\reply The subsection explaining type classes (Section 3.1) has been rewritten for clarity, with simpler language and improved examples.

\section{Response to Reviewer 2}

\pointRaised{Comment 1:}{The related work section could benefit from additional references and a clearer structure.}

\reply We have expanded the Related Work section to include “Just TYPEical: Visualizing Common Function Type Signatures in R” and added an introductory overview to outline its structure.

\pointRaised{Comment 2:}{The discussion section is abstract and lacks concrete examples.}

\reply We have revised the Discussion section to include several new examples illustrating potential applications of GeckoGraph.

\pointRaised{Comment 3:}{Some figures and captions are overly detailed.}

\reply We have shortened all figure captions for conciseness and removed Figure 4 from the earlier version, as it did not contribute additional insight.

\pointRaised{omment 4:}{The paper’s flow between sections feels abrupt.}

\reply We added a transitional paragraph between the Design Goals and Detailed GeckoGraph Features sections to improve narrative flow.

\pointRaised{Comment 5:}{Minor issues with grammar, formatting, and coherence.}

\reply We have now carefully proofread the entire manuscript, correcting grammatical errors, LaTeX syntax issues, and improving overall readability.

\end{document}

### 2025-10-25 response.tex
More changes

@@
 Authors of "GeckoGraph: A Visual Language for Polymorphic Types"
-
 \section{Summary of Major Revisions}
 
-\paragraph{Balanced claims} – We have now moderated the tone in the Abstract, Introduction, and Conclusion to be less ambitious and more balanced, explicitly acknowledging both the strengths and limitations of our findings.
+\paragraph{Balanced claims} We have moderated the tone in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} to present a more balanced and evidence-based discussion, explicitly acknowledging both the strengths and limitations of our findings.
 
-\paragraph{Terminology refinement} – We have now replaced references to the user study as a “game” with more precise wording, such as “gamified user study” or simply “user study.”
+\paragraph{Terminology refinement} We have replaced references to the user study as a “game” with more accurate terms such as “gamified user study” or simply “user study,” to avoid overstating the level of interactivity.
 
-\paragraph{New background section} – We have now added a new section, Type Systems in Functional Languages, after the Introduction to provide theoretical background and motivate the design of GeckoGraph.
+\paragraph{New background section} We have added a new section, \textit{Type Systems in Functional Languages}, following the \textit{Introduction} to provide theoretical background and better motivate the design of GeckoGraph.
 
-\paragraph{Section structure overviews} – We have now added section structure overviews at the beginning of the Related Work and Discussion sections to better guide the reader.
+\paragraph{Section structure overviews} We have introduced short structural overviews at the beginning of the \textit{Related Work} and \textit{Discussion} sections to help guide readers through their organization.
 
-\paragraph{Expanded related work} – We have now discussed the relevant study “Just TYPEical: Visualizing Common Function Type Signatures in R” to strengthen the discussion of related approaches and compare to our own contributions.
+\paragraph{Expanded related work} We have extended the \textit{Related Work} section to include additional references, notably the paper “Just TYPEical: Visualizing Common Function Type Signatures in R,” and used it to better position GeckoGraph among related visualization approaches.
 
-\paragraph{Enhanced discussion} – We have now improved the Discussion section with more concrete examples demonstrating the potential applications of GeckoGraph.
+\paragraph{Enhanced discussion} The \textit{Discussion} section has been revised to include more concrete examples demonstrating potential applications of GeckoGraph.
 
-\paragraph{Concise figure captions} – We have shortened our figure captions throughout the manuscript for readability and stylistic consistency.
+\paragraph{Concise figure captions} All figure captions have been shortened for clarity, readability, and stylistic consistency.
 
-\paragraph{Improved flow between sections} – We have added a transitional paragraph linking the Design Goals and Detailed GeckoGraph Features subsections in Section 3.1.
+\paragraph{Improved flow between sections} A transitional paragraph has been added to connect the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3.1, improving the logical flow of ideas.
 
-\paragraph{Clarified explanation of type classes} – We have now refined the text in Section 3.1 to improve clarity when explaining type classes.
+\paragraph{Clarified explanation of type classes} The explanation of type classes in Section~3.1 has been rewritten in simpler language with improved examples.
+Figure~3 now includes concrete illustrations to support the description.
 
-\paragraph{Removed unused figure} – We removed Figure 4 from the previous version, as it did not add new information to the paper.
+\paragraph{Removed unused figure} We removed the previous Figure~4, as it did not contribute additional information to the paper.
 
-\paragraph{Cosmetic and consistency edits} – We improved our grammar, spelling, LaTeX syntax, and overall logical coherence across the manuscript.
+\paragraph{Cosmetic and consistency edits} Grammar, spelling, LaTeX syntax, and stylistic consistency have been carefully reviewed and corrected throughout the manuscript.
 
-\section{Detailed Responses to Reviewers
-Response to Reviewer 1}
+\section{Detailed Responses to Reviewers}
 
-\pointRaised{Comment 1:}{ The manuscript’s tone in the abstract and conclusion appears overly strong given the presented evidence.}
+\subsection*{Response to Reviewer 1}
+
+\pointRaised{Comment 1:}{The manuscript’s tone in the abstract and conclusion appears overly strong given the presented evidence.}
 
 \reply We appreciate this observation.
-We have moderated the language in the Abstract, Introduction, and Conclusion to ensure our claims are proportionate to the results, and to present both positive and negative aspects of our findings.
+The language in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} has been moderated to ensure that claims are proportionate to the results and that both positive and negative aspects of our findings are clearly presented.
 
@@
 
-\reply We agree that we don't meet all the expectations for a game, and it's more appropriate to frame it as "an entertaining coding challenge".
-We changed our word of choice for “game” to “gamified user study” or simply, “user study”.
+\reply We agree.
+The tasks in our study do not fully meet the criteria for a “game.” We now describe it more accurately as a “gamified user study” or simply a “user study,” emphasizing its educational and experimental nature.
 
-\pointRaised{Comment 3:}{The paper lacks sufficient background on type systems in functional languages.}
+\pointRaised{Comment 3:}{Please clarify the explanation of type classes.}
 
-\reply We have added a new section titled Type Systems in Functional Languages immediately following the Introduction.
-This section introduces relevant concepts and better motivates the design of GeckoGraph.
+\reply The subsection on type classes (Section~3.1) has been rewritten for clarity and supported by clearer examples.
+Figure~3 has also been improved to include illustrative examples that reinforce our description of GeckoGraph.
 
-\pointRaised{Comment 4:}{Please clarify the explanation of type classes.}
+\pointRaised{Comment 4:}{Issues with figures and references.}
 
-\reply The subsection explaining type classes (Section 3.1) has been rewritten for clarity, with simpler language and improved examples.
+\reply The previous Figure~4 has been removed. Incorrect cross-references such as “Fig.~2.2” have been corrected (now “Fig.~8”). Repetitions such as “Appendix Appendix~A” have been fixed.
 
-\section{Response to Reviewer 2}
+\pointRaised{Comment 5:}{Temper the narrative of qualitative results—the data show neutral user opinions.}
 
-\pointRaised{Comment 1:}{The related work section could benefit from additional references and a clearer structure.}
+\reply We have clarified the neutrality of participants’ overall feedback and added clearer indicators of sentiment (e.g., “slightly positive,” “moderately negative”) to reflect the scale values more accurately.
 
-\reply We have expanded the Related Work section to include “Just TYPEical: Visualizing Common Function Type Signatures in R” and added an introductory overview to outline its structure.
+\pointRaised{Comment 6:}{Clarify speculative claims in the Discussion.}
 
-\pointRaised{Comment 2:}{The discussion section is abstract and lacks concrete examples.}
+\reply We now explicitly mark speculative interpretations in Section~6 (\textit{Discussion}) with leading phrases such as “We hypothesize that...”.
 
-\reply We have revised the Discussion section to include several new examples illustrating potential applications of GeckoGraph.
+\pointRaised{Comment 7:}{Wording and grammar issues.}
 
-\pointRaised{Comment 3:}{Some figures and captions are overly detailed.}
+\reply We have corrected numerous stylistic and grammatical errors, including:
+“expressiveness power” → “expressiveness”,
+“high-order function” → “higher-order function”,
+and “GechoGraph” → “GeckoGraph.”  
+Unclear terms such as “tangibly” and “regular higher-order function” have been clarified or replaced with precise expressions and supported with examples.  
+We also removed unnecessary spaces before footnote labels.
 
-\reply We have shortened all figure captions for conciseness and removed Figure 4 from the earlier version, as it did not contribute additional insight.
+\subsection*{Response to Reviewer 2}
 
-\pointRaised{omment 4:}{The paper’s flow between sections feels abrupt.}
+\pointRaised{Comment 1:}{The paper lacks sufficient background on type systems in functional languages.}
 
-\reply We added a transitional paragraph between the Design Goals and Detailed GeckoGraph Features sections to improve narrative flow.
+\reply We have added a new section titled \textit{Type Systems in Functional Languages} immediately after the \textit{Introduction}.
+This section introduces key theoretical concepts and motivates the design rationale for GeckoGraph.
 
-\pointRaised{Comment 5:}{Minor issues with grammar, formatting, and coherence.}
+\pointRaised{Comment 2:}{The related work section could benefit from additional references and a clearer structure.}
 
-\reply We have now carefully proofread the entire manuscript, correcting grammatical errors, LaTeX syntax issues, and improving overall readability.
+\reply We have expanded and reorganized the \textit{Related Work} section, including new references such as “Just TYPEical: Visualizing Common Function Type Signatures in R.” A short overview paragraph now outlines the section’s structure and focus.
+
+\pointRaised{Comment 3:}{The discussion section is abstract and lacks concrete examples.}
+
+\reply We have revised the \textit{Discussion} section to include additional examples illustrating practical applications of GeckoGraph. Figure~14 has been added to support this discussion.
+
+\pointRaised{Comment 4:}{Begin Sec.~5 (“Discussion”) and Sec.~6 (“Related Work”) with overview paragraphs explaining purpose and structure.}
+
+\reply Overview paragraphs have been added at the start of both sections to help orient readers and clarify how each section is organized.
+
+\pointRaised{Comment 5:}{Usage of figures.Some figures (1,3,10,14 previous version) captions are too long. Figure 4 is hard to understand and its purpose is unclear.}
+
+\reply Long figure captions have been shortened for conciseness. The previous Figure~4, which did not contribute new insights, has been removed.
+
+\pointRaised{Comment 6:}{The paper’s flow between sections feels abrupt.}
+
+\reply We have added a transitional paragraph between the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3 to improve the narrative flow.
+
+\pointRaised{Comment 7:}{Minor issues with grammar, formatting, and coherence.}
+
+\reply The manuscript has been carefully proofread to correct grammatical and LaTeX syntax errors and improve readability.  
+Examples include correcting figure references (e.g., “Fig.~2.2” → “Fig.~8”), fixing duplicate words (“Appendix Appendix~A” → “Appendix~A”), and removing extra spaces before footnote labels.
+
+\pointRaised{Comment 8:}{Clarify user tasks and results in Figures 9 and 10.}
+
+\reply We have clarified that the user in Figures~9 and~10 did not reach the correct solution and explained how GeckoGraph’s visual feedback can help identify and correct such mistakes.
+We also fixed a typo in Figure~9, changing “runHero z” to “runZero z.”
+
+\pointRaised{Comment 9:}{Clarify whether participants indicate their experience once or before each run.}
+
+\reply We clarified that participants report their programming experience once, before beginning the tasks.
+
+\pointRaised{Comment 10:}{Distinguish between empirical findings and authors’ interpretations in the Discussion.}
+
+\reply We now explicitly distinguish between observed results and speculative interpretations in the \textit{Discussion}, using clear phrases such as “We hypothesize that...” to signal interpretive commentary.
+
+\pointRaised{Comment 10:}{Add reference for “between-subject design.”}
+
+\reply We now correctly cite Fisher, R.A. {\it The Design of Experiements} for the phase "between-subject design."
 

### 2025-10-25 main.tex
More changes

@@
 They allow functions to be written in a way that can be used with different data types, while still enforcing the relationship and constraints between the values.
-However, programmers often find polymorphic types difficult to use and understand, and tend to reason with a mindset of concrete types.
+However, programmers often find polymorphic types difficult to use and understand, and tend to reason in terms of concrete types.
 We propose GeckoGraph, a graphical notation for types.
 GeckoGraph aims to complement traditional text-based type notation by making reading, understanding, and comparing types easier.
-We conducted a large-scale user study using GeckoGraph, comparing it to text-based type notation.
+We conducted a large-scale user study comparing GeckoGraph to text-based type notation.
 To our knowledge, this is the largest controlled user study on functional programming ever conducted.
@@
 Polymorphic types are one of the oldest topics in programming language theory \cite{Cardelli1987-fp}.
-They are central to the succinctness of statically-typed languages, while also enabling a high degree of type-safe abstraction and reusability.
+They are central to the succinctness of statically typed languages while also enabling a high degree of type-safe abstraction and reusability.
 Polymorphic types are available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}.
@@
 
-While promoting robustness and code reusability, polymorphic types present challenges in learning and comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}; as it is often argued that the expressiveness power of type systems often comes at the cost of their usability~\cite{Hage2020-hg}.
+While promoting robustness and code reusability, polymorphic types present challenges in learning and comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}; as it is often argued that the expressiveness of type systems often comes at the cost of their usability~\cite{Hage2020-hg}.
 Research indicates that humans tend to focus on concrete types and rely on polymorphic type checking only as a last resort.
-Moreover, polymorphic types often complicate type error usability.
+Moreover, polymorphic types often complicate the usability of type errors.
 During type checking, compilers frequently generate new polymorphic type variables temporarily.
@@
 GeckoGraph aims to complement traditional text-based type notation and make reading, understanding, and comparing types easier.
-GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts.
+GeckoGraph is prototyped and iteratively verified, resulting in a design with visual clarity applicable across many programming contexts.
  To evaluate its effectiveness, we conducted a large-scale user study comparing GeckoGraph with standard text-based type notation.
@@
 
-\section{Type Systems in Functional Programming}
+% In Section~\ref{sec:type-systems}, we review key type system features in functional programming languages, many of which have recently been adopted by object-oriented and multi-paradigm languages.
+We highlight how these features enable flexible abstraction and expressive type modeling, while sometimes hindering usability for beginners.
+Section~\ref{sec:gecko-graph} presents the design of GeckoGraph, illustrating how it visualizes complex type-level information that is often implicit or hard to interpret textually.
+Section~\ref{sec:evaluation} describes our user study, ZeroToHero, a series of 10 programming challenges, along with the study design, recruitment, and hypotheses.
+Section~\ref{sec:results} reports quantitative and qualitative findings, including completion time, success rates, and participant feedback.
+In Section~\ref{sec:discussion}, we interpret these results, discuss GeckoGraph’s benefits and usability challenges, and explore potential applications.
+Section~\ref{sec:related-work} reviews related visualization research, and Section~\ref{sec:conclusion} summarizes our contributions and outlines future directions.
+
+\section{Type Systems in Functional Programming} \label{sec:type-systems}
 
@@
 
-\section{GeckoGraph}
+\section{GeckoGraph} \label{sec:gecko-graph}
 
@@
 Participants are assigned to one of two groups (A and B).
-The experiment uses a between-subjects design.
+The experiment uses between-subjects design \cite{Fisher1949-fn}.
 The tasks assigned to the two groups are the same.
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 
@@
 This section discusses the main findings and implications of our study. 
-We begin by outlining the strengths of GeckoGraph observed from both quantitative and qualitative results (Section~\ref{sec:discussion:strengths}). 
+We speculate the strengths of GeckoGraph observed from both quantitative and qualitative results (Section~\ref{sec:discussion:strengths}). 
 We then discuss the limitations of the current design (Section~\ref{sec:discussion:weaknesses}) and reflect on the gamified nature of our user study (Section~\ref{sec:discussion:gamification}). 
@@
 Overall, our experimental results show no statistically significant difference in task completion time between conditions with and without GeckoGraph.
-However, we find that using GeckoGraph shows a significant albeit moderate effect on the task success rates for beginners.
-Particularly, in the hardest task (task 10), GeckoGraph improved the success rates across the two less experienced user groups by 14.5\% and 11.2\%, respectively.
-To interpret the observed advantages, we discuss the practical benefits that GeckoGraph offers to programmers.
+However, we find that using GeckoGraph shows a significant, albeit moderate, effect on task success rates for beginners.
+In particular, on the hardest task (task 10), GeckoGraph improved the success rates across the two less experienced user groups by 14.5\% and 11.2\%, respectively.
+To interpret the observed advantages, we hypothesize the practical benefits that GeckoGraph offers to programmers.
 
@@
 Post-study feedback highlighted a strong interest in integrating GeckoGraph into programming environments. 
-Modern IDEs already allow users to inspect expression types via hover tooltips; GeckoGraph could enhance this feature by visualizing type structures directly in the tooltip. 
-Similarly, GeckoGraph could be used to visualize mismatched parts of conflicting types during compilation errors, as illustrated in Fig.~\ref{fig:example-foldable}.
+Modern IDEs already allow users to inspect expression types via hover tooltips; GeckoGraph could enhance this feature by visualizing type structures directly in the tooltip, as envisioned in Fig.~\ref{fig:application} (Left).
+Similarly, GeckoGraph could be used to visualize mismatched parts of conflicting types during compilation errors, as illustrated in Fig.~\ref{fig:application} (Right).
+
+\begin{figure}[]
+  \includegraphics[width=\linewidth]{figures/Application.png}
+  \caption{\label{fig:application} 
+ {\bf Envisioning GeckoGraph as programming assistance}. (Left) Display GeckoGraph as type hints; (Right) Visualize mismatched types with GeckoGraph. }
+\end{figure}
 
@@
 
-\section{Related work}
+\section{Related work} \label{sec:related-work}
 This section reviews prior work in three areas relevant to this research. First, we examine approaches to the visualization of types in programming languages. 
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion and Future Work} \label{sec:conclusion}
 In this paper, we propose GeckoGraph, a graphical notation for type annotations in functional programming languages.

### 2025-10-26 response.tex
More changes

@@
 
+\pointRaised{Comment 8:}{Clairify the syntactic sugar of curried function (Section 3)}
+
+\reply We explicitly clarify that function notation \texttt{a -> b -> c } desugars to \texttt{a -> (b -> c) }, which has a consistent representation in GeckoGraph.
+
 \subsection*{Response to Reviewer 2}
@@
 
-\pointRaised{Comment 10:}{Add reference for “between-subject design.”}
+\pointRaised{Comment 11:}{Add reference for “between-subject design.”}
 
@@
 
+\pointRaised{Comment 12:}{Wording and grammar issues.}
+
+\reply We have corrected numerous stylistic and grammatical errors, including:
+Unclear terms such as “approximity” and “during type errors” have been clarified or replaced with precise expressions and supported with examples.
+
 \end{document}

### 2025-10-26 main.tex
More changes

@@
 
-% In Section~\ref{sec:type-systems}, we review key type system features in functional programming languages, many of which have recently been adopted by object-oriented and multi-paradigm languages.
+In Section~\ref{sec:type-systems}, we review key type system features in functional programming languages, many of which have recently been adopted by object-oriented and multi-paradigm languages.
 We highlight how these features enable flexible abstraction and expressive type modeling, while sometimes hindering usability for beginners.
-Section~\ref{sec:gecko-graph} presents the design of GeckoGraph, illustrating how it visualizes complex type-level information that is often implicit or hard to interpret textually.
+Section~\ref{sec:gecko-graph} presents the design of GeckoGraph, illustrating how it supports the type level features outlined in Section~\ref{sec:type-systems}.
 Section~\ref{sec:evaluation} describes our user study, ZeroToHero, a series of 10 programming challenges, along with the study design, recruitment, and hypotheses.
@@
 
-Curried functions (e.g., \texttt{ a -> b -> c}) are displayed as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}.
+Curried functions are functions that return other functions (e.g.
+\texttt{ a -> (b -> c)}).
+In functional languages, they are often written with a syntactic shorthand, such as \texttt{ a -> b -> c}).
+These two forms are uniquely represented in GeckoGraph as two functions cells merged together \includegraphics[height=1em]{figures/Curry.png}.
 It appears as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
@@
 Bootstrapping was used for computing the confidence interval because of the non-normal distribution.
-In most tasks, GeckoGraph provides a small edge.
-However, significant differences were found in task 10 of the beginner group and task 10 of the familiar group.
+In most tasks, GeckoGraph provides no significant differences.
+However, substantial improvements were observed in task 10 for the beginner group (14.5\%) and the familiar group (11.2\%).
 }

### 2025-10-26 response.tex
Final updates

--- 2025-10-25 response.tex
@@
 \paragraph{Cosmetic and consistency edits} Grammar, spelling, LaTeX syntax, and stylistic consistency have been carefully reviewed and corrected throughout the manuscript.
-
 \section{Detailed Responses to Reviewers}
@@
 
-\pointRaised{Comment 1:}{The manuscript’s tone in the abstract and conclusion appears overly strong given the presented evidence.}
+\pointRaised{Comment 1:}{(Abstract, Introduction, Conclusion – Major) The manuscript’s tone in the Abstract and Conclusion appears overly strong given the presented evidence.}
 
 \reply We appreciate this observation.
-The language in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} has been moderated to ensure that claims are proportionate to the results and that both positive and negative aspects of our findings are clearly presented.
+The language in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} has been moderated to ensure that claims are proportionate to the results.
+The revised version now presents both the positive and negative aspects of our findings.
 
-\pointRaised{Comment 2:}{The term “game” used to describe the user study could be misleading.}
+\pointRaised{Comment 2:}{(Overall) The term “game” used to describe the user study could be misleading.}
 
@@
 
-\pointRaised{Comment 3:}{Please clarify the explanation of type classes.}
+\pointRaised{Comment 3:}{(Section 3.1) Clarify the explanation of type classes.}
 
@@
 
-\pointRaised{Comment 4:}{Issues with figures and references.}
+\pointRaised{Comment 4:}{(Section 3) Clarify the syntactic sugar of curried functions.}
 
-\reply The previous Figure~4 has been removed. Incorrect cross-references such as “Fig.~2.2” have been corrected (now “Fig.~8”). Repetitions such as “Appendix Appendix~A” have been fixed.
+\reply We now explicitly clarify that the function notation \texttt{a -> b -> c} desugars to \texttt{a -> (b -> c)}, which is consistently represented in GeckoGraph.
 
-\pointRaised{Comment 5:}{Temper the narrative of qualitative results—the data show neutral user opinions.}
+\pointRaised{Comment 5:}{(Section 3) Figure 3: The second higher-order function type {\tt ((a -> b) -> b)} is tagged as binary but is unary.}
 
-\reply We have clarified the neutrality of participants’ overall feedback and added clearer indicators of sentiment (e.g., “slightly positive,” “moderately negative”) to reflect the scale values more accurately.
+\reply We agree that the previous figure was confusing. Figure~4 (previously Figure~3) has been simplified for readability.
 
-\pointRaised{Comment 6:}{Clarify speculative claims in the Discussion.}
+\pointRaised{Comment 6:}{(Section 3) Issues with figures and references.}
 
-\reply We now explicitly mark speculative interpretations in Section~6 (\textit{Discussion}) with leading phrases such as “We hypothesize that...”.
+\reply The previous Figure~4 has been removed. Incorrect cross-references such as “Fig.~2.2” have been corrected (now “Fig.~8”), and redundant phrases like “Appendix Appendix~A” have been fixed.
 
-\pointRaised{Comment 7:}{Wording and grammar issues.}
+\pointRaised{Comment 7:}{(Section 5) Temper the narrative of qualitative results—the data show neutral user opinions.}
 
-\reply We have corrected numerous stylistic and grammatical errors, including:
+\reply We have clarified that participants’ feedback was overall neutral. Descriptors such as “slightly positive” and “moderately negative” are now used to reflect sentiment values more precisely.
+
+\pointRaised{Comment 8:}{(Section 6) Clarify speculative claims in the Discussion.}
+
+\reply We now explicitly mark speculative interpretations in Section~6 (\textit{Discussion}) using leading phrases such as “We hypothesize that...”.
+
+\pointRaised{Comment 9:}{(Overall) Wording and grammar issues.}
+
+\reply We have corrected numerous stylistic and grammatical issues, including:
 “expressiveness power” → “expressiveness”,
-“high-order function” → “higher-order function”,
+“high-order function” → “higher-order function,”
 and “GechoGraph” → “GeckoGraph.”  
-Unclear terms such as “tangibly” and “regular higher-order function” have been clarified or replaced with precise expressions and supported with examples.  
-We also removed unnecessary spaces before footnote labels.
+Unclear expressions such as “tangibly” and “regular higher-order function” have been revised or replaced with precise terms, and unnecessary spaces before footnotes have been removed.
 
@@
 
-\pointRaised{Comment 1:}{The paper lacks sufficient background on type systems in functional languages.}
+\pointRaised{Comment 1:}{(Section 1) Add a concluding paragraph in the Introduction describing the article’s organization.}
+
+\reply We agree. A short paper-structure overview has been added at the end of Section~1 to guide the reader through the paper’s organization.
+
+\pointRaised{Comment 2:}{(Section 2) The paper lacks sufficient background on type systems in functional languages.}
 
@@
 
-\pointRaised{Comment 2:}{The related work section could benefit from additional references and a clearer structure.}
+\pointRaised{Comment 3:}{(Section 3) The paper’s flow between subsections feels abrupt.}
 
-\reply We have expanded and reorganized the \textit{Related Work} section, including new references such as “Just TYPEical: Visualizing Common Function Type Signatures in R.” A short overview paragraph now outlines the section’s structure and focus.
+\reply We have added a transitional paragraph between the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3 to improve narrative flow.
 
-\pointRaised{Comment 3:}{The discussion section is abstract and lacks concrete examples.}
-
-\reply We have revised the \textit{Discussion} section to include additional examples illustrating practical applications of GeckoGraph. Figure~14 has been added to support this discussion.
-
-\pointRaised{Comment 4:}{Begin Sec.~5 (“Discussion”) and Sec.~6 (“Related Work”) with overview paragraphs explaining purpose and structure.}
-
-\reply Overview paragraphs have been added at the start of both sections to help orient readers and clarify how each section is organized.
-
-\pointRaised{Comment 5:}{Usage of figures.Some figures (1,3,10,14 previous version) captions are too long. Figure 4 is hard to understand and its purpose is unclear.}
-
-\reply Long figure captions have been shortened for conciseness. The previous Figure~4, which did not contribute new insights, has been removed.
-
-\pointRaised{Comment 6:}{The paper’s flow between sections feels abrupt.}
-
-\reply We have added a transitional paragraph between the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3 to improve the narrative flow.
-
-\pointRaised{Comment 7:}{Minor issues with grammar, formatting, and coherence.}
-
-\reply The manuscript has been carefully proofread to correct grammatical and LaTeX syntax errors and improve readability.  
-Examples include correcting figure references (e.g., “Fig.~2.2” → “Fig.~8”), fixing duplicate words (“Appendix Appendix~A” → “Appendix~A”), and removing extra spaces before footnote labels.
-
-\pointRaised{Comment 8:}{Clarify user tasks and results in Figures 9 and 10.}
+\pointRaised{Comment 4:}{(Section 4) Clarify user tasks and results in Figures 9 and 10.}
 
@@
 
-\pointRaised{Comment 9:}{Clarify whether participants indicate their experience once or before each run.}
+\pointRaised{Comment 5:}{(Section 4) Clarify whether participants indicate their experience once or before each run.}
 
@@
 
-\pointRaised{Comment 10:}{Distinguish between empirical findings and authors’ interpretations in the Discussion.}
+\pointRaised{Comment 6:}{(Section 4) Add reference for “between-subject design.”}
 
-\reply We now explicitly distinguish between observed results and speculative interpretations in the \textit{Discussion}, using clear phrases such as “We hypothesize that...” to signal interpretive commentary.
+\reply We now correctly cite Fisher, R.A. \textit{The Design of Experiments} for the phrase “between-subject design.”
 
-\pointRaised{Comment 10:}{Add reference for “between-subject design.”}
+\pointRaised{Comment 7:}{(Section 6) The Discussion section is abstract and lacks concrete examples.}
 
-\reply We now correctly cite Fisher, R.A. {\it The Design of Experiements} for the phase "between-subject design."
+\reply We have revised the \textit{Discussion} section to include additional examples illustrating practical applications of GeckoGraph. Figure~14 has been added to support this discussion.
+
+\pointRaised{Comment 8:}{(Section 6 and 7) Begin Sections~6 (“Discussion”) and~7 (“Related Work”) with overview paragraphs explaining purpose and structure.}
+
+\reply Overview paragraphs have been added at the start of both sections to help orient readers and clarify their organization.
+
+\pointRaised{Comment 9:}{(Section 7) The Related Work section could benefit from additional references and a clearer structure.}
+
+\reply We have expanded and reorganized the \textit{Related Work} section, including new references such as “Just TYPEical: Visualizing Common Function Type Signatures in R.” A short overview paragraph now outlines the section’s focus and structure.
+
+\pointRaised{Comment 10:}{(Section 6) Distinguish between empirical findings and interpretations in the Discussion.}
+
+\reply We now explicitly distinguish between observed results and speculative interpretations in the \textit{Discussion}, using phrases such as “We hypothesize that...” to signal interpretive commentary.
+
+\pointRaised{Comment 11:}{(Overall) Usage of figures—some captions too long, Figure~4 unclear.}
+
+\reply Figure captions throughout the paper have been shortened for conciseness. The previous Figure~4, which did not add new insight, has been removed.
+
+\pointRaised{Comment 12:}{(Overall) Minor wording and grammar issues.}
+
+\reply The manuscript has been thoroughly proofread for grammar, formatting, and coherence.  
+Examples include corrected figure references (e.g., “Fig.~2.2” → “Fig.~8”), removal of redundant text (“Appendix Appendix~A” → “Appendix~A”), and clarification of unclear terms such as “approximity” and “during type errors.”
 

### 2025-10-26 main.tex
Final updates

@@
 Most modern languages employ some level of inference.
-For instance, Java’s var keyword (introduced in version 10) infers variable types at compile time based on their assigned values \cite{Java_Developers2023-an}, saving programmers from writing redundant type declarations, such as \textit {Car car = new Car()}.
+For instance, Java’s var keyword (introduced in version 10) infers variable types at compile time based on their assigned values \cite{Java-Developers2023-sp}, saving programmers from writing redundant type declarations, such as \textit {Car car = new Car()}.
 
@@
 It appears as if the second function covers the right half of the first function, indicating that the second function is the return type of the first.
-Regular high-order functions (e.g., \texttt{(a -> b) -> c}) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
+Higher-order functions that take other functions as arguments  (e.g., \texttt{(a -> b) -> c}) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}.
 The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and higher-order functions (design goal \ref{goal3}).
@@
 \begin{figure}[h]
-  \includegraphics[width=\linewidth]{figures/Indicator}
+  \includegraphics[width=0.6\linewidth]{figures/Indicator.png}
   \caption{
         \label{fig:indicator}
-        An example of the visual hierarchy of the function indicator.
-The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators.
-For higher-order functions where functions are arguments or return values of other functions, it is very easy to see the ``level" of functions.
+        An example of the visual hierarchy of the function indicator. The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators.
   }
